### 2015-08-21 MScWriteup.tex
Initial commit of files for the writeup

\documentclass[11pt,journal]{./IEEE_latex_class/IEEEtran}

% In order to reference go to http://truben.no/latex/bibtex/

% *** CITATION PACKAGES ***
\usepackage{cite}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{./Figures/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx} 
  % declare the path(s) where your graphic files are 
  % \graphicspath{{../eps/}}   
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

% *** MATH PACKAGES ***
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%  OTHER PACKAGES

\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes}

\usepackage[caption=false]{subfig}

\begin{document}

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Computational design of RNA-based oscillatory circuits}

\author{J.~Binysh
        \\ \IEEEmembership{* University of Warwick, Complexity Department}
        \vspace{-3em}}

% The paper headers
\markboth{hi}%
{Optimisation of Car Electronics}

% make the title area
\maketitle

% This defines the title + page out of # of pages at the heading of the pages
\thispagestyle{empty}

\newcommand{\MYheader}{\smash{\scriptsize

\hfil\parbox[t][\height][t]{\textwidth}{\centering {\normalsize
Place conference title here}}\hfil\hbox{}}}
\makeatletter

\if@twoside
  \def\ps@headings{%
      \let\@oddfoot\@empty\let\@evenfoot\@empty
      \def\@evenhead{\small\thepage\hfil\leftmark\strut\vadjust{\vskip .1ex\hrule}}%
      \def\@oddhead{\small\rightmark\hfil\thepage\strut\vadjust{\vskip .1ex\hrule}}%
      \let\@mkboth\markboth
    \def\chaptermark##1{%
      \markboth{\scshape%
        \ifnum \c@secnumdepth >\m@ne
            \@chapapp\ \thechapter. \ %
        \fi
        ##1}{}}%
    \def\sectionmark##1{%
      \markright{\scshape%
        \ifnum \c@secnumdepth >\z@
          \thesection. \ %
        \fi
        ##1}}}
\else
  \def\ps@headings{%
    \let\@oddfoot\@empty
    \def\@oddhead{{\slshape\rightmark}\hfil\thepage\ of\ \pageref{LastPage} \strut\vadjust{\vskip .1ex\hrule}}%
    \let\@mkboth\markboth
    \def\chaptermark##1{%
      \markright{\scshape%
        \ifnum \c@secnumdepth >\m@ne
            \@chapapp\ \thechapter. \ %
        \fi
        ##1}}}
\fi
\makeatother

\makeatother

% make changes take effect
\pagestyle{headings}
% adjust as needed
%\addtolength{\footskip}{0\baselineskip}
%\addtolength{\textheight}{-0.1\baselineskip}

\begin{abstract}
the abstract
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec: Intro}
The process of gene expression can be briefly summarised as follows: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}). This RNA molecule (known as messenger RNA, or mRNA) makes its way to a piece of cellular machinery called the Ribosome, which reads it, and makes a protein - which protein is made depends on the DNA sequence originally read (\textit{translation}).

 The path from genetic transcription to protein expression is naturally regulated in many ways \cite{MolecularBiology}. This regulation allows a cell to control protein expression, and so cell behaviour, in response to various environmental cues. The natural cell machinery in place to perform this regulation offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.

As well as acting as the intermediate between DNA and protein, RNA's play direct and important roles in regulating cell behaviour \cite{Isaacs2006}

 \begin{itemize}
 \item goal of the project - emph on quantitative modelling of biological system and data, getting a quantitative model, finding its parameters.
 \item Biological background, taken from those 3 papers.
 \item give structure of the report.
 \end{itemize}

%%%
%NEW SECTION%
%%%
\section{Results and Discussion}
\label{Results and Discussion}

%NEW SUBSECTION%

\subsection{Model Derivation}

 \begin{itemize}
 \item introduce the proposed ode model, explain its biology.
 \item Biological background, taken from those 3 papers.
 \end{itemize}

%NEW SUBSECTION%

\subsection{Parameter Estimation}
 \begin{itemize}
\item basically the whole project.
\cite{Isaacs2006}
 \end{itemize}

\section{Conclusions and Further work}

% references section
\bibliographystyle{./IEEE_bibtex_class/IEEEtran}
\bibliography{./IEEE_bibtex_class/IEEEabrv,library}

\end{document}

### 2015-08-21 MScWriteup.tex
removing old images

@@
-\documentclass[11pt,journal]{./IEEE_latex_class/IEEEtran}
+\documentclass[10pt,journal]{./IEEE_latex_class/IEEEtran}
 
@@
  The path from genetic transcription to protein expression is naturally regulated in many ways \cite{MolecularBiology}.
-This regulation allows a cell to control protein expression, and so cell behaviour, in response to various environmental cues.
+This regulation allows the cell to control protein expression, and so cell behaviour, in response to various environmental cues.
 The natural cell machinery in place to perform this regulation offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.
 
-As well as acting as the intermediate between DNA and protein, RNA's play direct and important roles in regulating cell behaviour \cite{Isaacs2006}
+As well as acting as the intermediate between DNA and protein, RNA molecules play direct and important roles in regulating cell behaviour \cite{Isaacs2006}.
+For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA's offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription.
+In comparison to proteins, it is relatively straightforward to predict the structure and function of an RNA from its sequence using physiochemical models.
+Recently, this has been exploited to computationally design DNA sequences encoding synthetic sRNA's, with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.
+
+This report will focus on one such system, introduced in \cite{Rodrigo2012}.
+It will extend existing understanding of the system beyond the qualitative, by proposing a quantitative model of gene expression, in the form of a set of ODE's, and fitting it to available time series data to estimate the model's unknown parameters.
+
+The report is structured as follows. In section \ref{The sRNA regulatory system} we review the sRNA regulatory system we will consider. In section \ref{Results and Discussion} 
+we propose a set of ODE's to model the system, and estimate its unknown parameters by fitting to time series data.
+Finally, in section \ref{Conclusions and Further work}, we conclude, and suggest directions for further work.
 
@@
  \end{itemize}
+
+\section{The sRNA regulatory system}
+\label{The sRNA regulatory system}
+
+ \begin{itemize}
+ \item  
+ \item
+
+ \end{itemize}
+
+\begin{figure}[H]
+\centering
+\includegraphics[trim = 60 630 300 30,page=2,clip = true]{pnas1203831109.pdf}
+\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{community_figure}}
+\label{community}
+\end{figure}
 
@@
 \section{Conclusions and Further work}
+\label{Conclusions and Further work}
 

### 2015-08-23 MScWriteup.tex
writing the background section

@@
 \author{J.~Binysh
-        \\ \IEEEmembership{* University of Warwick, Complexity Department}
-        \vspace{-3em}}
+        \\ \IEEEmembership{* University of Warwick, Complexity Department}}
 
@@
 \begin{abstract}
-the abstract
+genetic circuitry, RNA's offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription.
+In comparison to proteins, it is relatively straightfor
 \end{abstract}
@@
 \label{sec: Intro}
-The process of gene expression can be briefly summarised as follows: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}).
+The process of gene expression can be briefly summarised as: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}).
 This RNA molecule (known as messenger RNA, or mRNA) makes its way to a piece of cellular machinery called the Ribosome, which reads it, and makes a protein - which protein is made depends on the DNA sequence originally read (\textit{translation}).
@@
 This regulation allows the cell to control protein expression, and so cell behaviour, in response to various environmental cues.
-The natural cell machinery in place to perform this regulation offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.
+The natural cell machinery which performs this regulation offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.
 
@@
 
- \begin{itemize}
- \item goal of the project - emph on quantitative modelling of biological system and data, getting a quantitative model, finding its parameters.
- \item Biological background, taken from those 3 papers.
- \item give structure of the report.
- \end{itemize}
-
 \section{The sRNA regulatory system}
 \label{The sRNA regulatory system}
+In bacteria, one mechanism by which gene expression is regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA.
+This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA.
+Translation may be self repressed by the tail of the mRNA folding over and binding across the RBS, forming a stem loop in the mRNA and preventing the Ribosome from binding (Fig.
+\ref{RBS}).
+This self repression may be released with an sRNA which binds to the same region on the mRNA - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind (Fig.
+\ref{RBS}}).
+In summary, the presence of the sRNA positively regulates gene expresssion.
+\todo[inline]{define sRNA}
 
- \begin{itemize}
- \item  
- \item
+\begin{figure}[H]
+\centering
+\includegraphics[trim = 100 170 100 400,page=10,clip = true,scale = 0.6]{Appendix.pdf}
+\caption{A mechanism by which sRNA's can regulate gene expression.
+Initially, the mRNA is folded over on itself in a loop, covering the RBS and blocking Ribosome binding.
+The sRNA binds to this mRNA, causing a conformational change which uncovers the RBS, and allows translation to occur.
+}
+\label{RBS}
+\end{figure}
 
- \end{itemize}
+\cite{Rodrigo2012} proposed a computational methodology to design general genetic circuits based on RNA interactions, and as a case study of the methodology chose to design a synthetic sRNA- mRNA pair capable of acting in the manner described above.
+The algorithm assumed an interaction scheme between the RNA's as shown in Fig.
+\ref{reactionscheme}.
+The two RNA's, originally in their own individually folded states, would initially interact via a small 'toehold' sequence of unpaired nucleotides to form an unstable transition state.
+This intermediate complex would then rapidly form a final, stable complex with the desired conformation.
+By suggesting sRNA and mRNA sequences which optimised this energy landscape, \cite{Rodrigo2012} suggested several devices which would form a stable hybrid with the RBS free, and experimentally validated their function in \textit{E.
+coli}.
 
@@
 \caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{community_figure}}
-\label{community}
+\label{reactionscheme}
+\end{figure}
+
+Further, by placing the concentrations of the sRNA and mRNA under the control of tuneable promoters, \cite{Rodrigo2012} constructs a logical AND gate from one of the proposed devices (RAJ11) \textit{in vivo} (Fig.
+\ref{ANDGate}).
+
+\begin{figure*}[b!]
+\centering
+\includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
+\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{community_figure}}
+\label{ANDGate}
 \end{figure}

### 2015-08-24 MScWriteup.tex
still writing background section

@@
 In comparison to proteins, it is relatively straightforward to predict the structure and function of an RNA from its sequence using physiochemical models.
-Recently, this has been exploited to computationally design DNA sequences encoding synthetic sRNA's, with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.
+Recently, this has been exploited to computationally design DNA sequences encoding synthetic sRNA's - small RNA's which do not code for a protein, but rather have some direct regulatory function -  with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.
 
@@
 This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA.
-Translation may be self repressed by the tail of the mRNA folding over and binding across the RBS, forming a stem loop in the mRNA and preventing the Ribosome from binding (Fig.
+In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the UTR), upstream of the RBS.
+Translation may be self repressed by this `tail' of the mRNA folding over and binding across the RBS, forming a stem loop in the mRNA and preventing the Ribosome from binding (Fig.
 \ref{RBS}).
 This self repression may be released with an sRNA which binds to the same region on the mRNA - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind (Fig.
-\ref{RBS}}).
+\ref{RBS}).
 In summary, the presence of the sRNA positively regulates gene expresssion.
-\todo[inline]{define sRNA}
 
@@
 \caption{A mechanism by which sRNA's can regulate gene expression.
-Initially, the mRNA is folded over on itself in a loop, covering the RBS and blocking Ribosome binding.
-The sRNA binds to this mRNA, causing a conformational change which uncovers the RBS, and allows translation to occur.
-}
+Initially, the 5' UTR of the mRNA is folded over the RBS, forming a loop and blocking Ribosome binding.
+The sRNA binds to this mRNA, causing a conformational change which uncovers the RBS, and allows translation to occur.Image reproduced from \cite{Rodrigo2012} }
 \label{RBS}
@@
 By suggesting sRNA and mRNA sequences which optimised this energy landscape, \cite{Rodrigo2012} suggested several devices which would form a stable hybrid with the RBS free, and experimentally validated their function in \textit{E.
-coli}.
+coli}, using an mRNA which codes for GFP for experimental ease (note the algorithm only optimises the 5' UTR of the mRNA, so the actual protein being coded for is unimportant).
 
@@
 \includegraphics[trim = 60 630 300 30,page=2,clip = true]{pnas1203831109.pdf}
-\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{community_figure}}
+\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{Rodrigo2012}}
 \label{reactionscheme}
@@
 \ref{ANDGate}).
+In this system, transcription of the designed sRNA and mRNA are placed under the control of promoter regions, $P_{LtetO-1}$ and $P_{LlacO-1}$ \cite{Lutz1997}.
+These are in turn controlled by two transcriptional repressors, TetR and LacI, which are naturally present in the strain of \text{E.
+coli} considered.
+These repressors disable the promoter regions, and so by default transcription of the RNA's is turned off, and no protein is produced.
+These repressors can themselves be disabled by the presence of two chemicals, aTC and IPTG, which can be introduced externally into the cell (Fig.
+\ref{ANDGate}).
+So transcription of the two RNA's is indirectly controlled by the presence of two chemicals - if neither is present, sRNA and mRNA transcription is repressed, and no protein is produced.
+If only one is present, the AND gate remains off, either because there is no mRNA to be translated into protein, or because the mRNA is self repressed.
+But when both are present, the conformational change discussed above occurs, and protein is produced.
+\todo[inline]{promoter?}
 
@@
 \includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
-\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{community_figure}}
+\caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS.
+In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{LtetO-1}$ and $P_{LlacO-1}$.
+These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E.
+coli} discussed.
+These chemical repressors are themselves disabled by two chemicals, aTC and IPTG.
+In the notation of the diagram, a barred line indicates repression, and an arrowed line indicates production.
+We see a 'double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA).
+Thus presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.
+Image reproduced from \cite{Rodrigo2012}.}
 \label{ANDGate}
-\end{figure}
+\end{figure*}
 

### 2015-08-24 MScWriteup.tex
writing up model section

@@
 
-\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes}
-
+\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e}
 \usepackage[caption=false]{subfig}
@@
 \markboth{hi}%
-{Optimisation of Car Electronics}
+{Computational design of RNA-based oscillatory circuits}
 
@@
 
+%%%
+%%%
+%%%
+%%%
+%%%
+%NEW SECTION%
+%%%
+%%%
+%%%
+%%%
+
 \section{Introduction}
@@
 
-The report is structured as follows. In section \ref{The sRNA regulatory system} we review the sRNA regulatory system we will consider. In section \ref{Results and Discussion} 
-we propose a set of ODE's to model the system, and estimate its unknown parameters by fitting to time series data.
+The report is structured as follows.
+In the remainder of this section we review the sRNA regulatory system we will consider.
+In section \ref{Results and Discussion} we propose a set of ODE's to model the system, and estimate its unknown parameters by fitting to time series data.
 Finally, in section \ref{Conclusions and Further work}, we conclude, and suggest directions for further work.
 
-\section{The sRNA regulatory system}
+%NEW SUBSECTION%
+
+\subsection{The sRNA regulatory system}
 \label{The sRNA regulatory system}
-In bacteria, one mechanism by which gene expression is regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA.
+In bacteria, one mechanism by which gene expression is regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA (Fig.
+\ref{RBS}).
 This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA.
@@
 \ref{RBS}).
-This self repression may be released with an sRNA which binds to the same region on the mRNA - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind (Fig.
-\ref{RBS}).
+This self repression may be released with an sRNA which binds to the same region on the mRNA - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind.
 In summary, the presence of the sRNA positively regulates gene expresssion.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
+\caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS.
+In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{LtetO-1}$ and $P_{LlacO-1}$.
+These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E.
+coli} discussed.
+These chemical repressors are themselves disabled by two chemicals, aTC and IPTG.
+In the notation of the diagram, a barred line indicates repression, and an arrowed line indicates production.
+We see a 'double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA).
+Thus presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.
+Image reproduced from \cite{Rodrigo2012}.}
+\label{ANDGate}
+\end{figure*}
 
@@
 
-\begin{figure*}[b!]
-\centering
-\includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
-\caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS.
-In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{LtetO-1}$ and $P_{LlacO-1}$.
-These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E.
-coli} discussed.
-These chemical repressors are themselves disabled by two chemicals, aTC and IPTG.
-In the notation of the diagram, a barred line indicates repression, and an arrowed line indicates production.
-We see a 'double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA).
-Thus presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.
-Image reproduced from \cite{Rodrigo2012}.}
-\label{ANDGate}
-\end{figure*}
+Although a qualitative understanding of this system exists \cite{Rodrigo2012}, it is of interest to attempt a quantitative understanding of the genetic circuit involved.
+Such an understanding would allow, for example, tailoring of the system in reponse to design requirements, by altering the values of the important parameters of the model.
+By changing which sRNA-mRNA device is used in the system, it would also allow exploration of the relationship between the thermodynamic properties of each device, and the model's rate constants.
 
+%%%
+%%%
+%%%
+%%%
 %%%
@@
 %%%
+%%%
+%%%
+%%%
+
 \section{Results and Discussion}
@@
 \subsection{Model Derivation}
+We propose a simple model, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}, to describe the system.
+Tables \ref{ModelParameters},\ref{StateVariables} give complete descriptions of parameter and state variable meanings.
 
- \begin{itemize}
- \item introduce the proposed ode model, explain its biology.
- \item Biological background, taken from those 3 papers.
- \end{itemize}
+\begin{align}
+\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\mu + \delta_{s})s -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m \label{eq:s}\\
+\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\mu + \delta_{m})m -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m  \label{eq:m}\\
+\frac{ds:m}{dt} & = k_{\mathrm{on}}sm  - (k_{\mathrm{off}}+ k_{\mathrm{hyb}})s:m  -(\mu + \delta_{sm} )s:m \label{eq:sm}\\
+\frac{dc}{dt} & = k_{\mathrm{hyb}}s:m  -(\mu + \delta_{c})c  \label{eq:c} \\
+\frac{dp}{dt} & = \beta m +f_{s}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}  \label{eq:p} \\
+\frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \label{eq:g} \\
+z &= z_{0} +\frac{g}{\theta} \label{eq:z}
+\end{align}
+  Based on the reaction mechanism in Fig.
+\ref{reactionscheme}, the hybridization of the sRNA and mRNA first into an unstable complex, then a stable one, is modelled in eqs.\ref{eq:s} - \ref{eq:c}.
+The initial binding is modelled as a reversible reaction with forward and backward rates $k_{on}$ and $k_{off}$.
+\begin{align*}
+sRNA \chemsign+ mRNA\chemrel[$k_{on}$][$k_{off}$]{<>} sRNA:mRNA_{unstable} 
+\end{align*}\\
+\noindent
+after which the stabilization is modelled as an irreversible reaction with rate $k_{hyb}$. 
+ \begin{align*}
+sRNA:mRNA_{unstable} \chemrel[$k_{hyb}$]{->} sRNA:mRNA_{stable} 
+\end{align*}
+
+ In addition, these complexes are given degradation rates, $\delta_{s}$, $\delta_{m}$, $\delta_{sm}$, $\delta_{c}$, and dilution of chemical concentrations due to cell growth are modelled with a dilution rate $\mu$.
+
+ Control of the system by aTc and IPTG are modelled by $y(t)$ and $x(t)$ in eqs.\ref{eq:s}, \ref{eq:m}.
+$y(t)$ models the response of the sRNA transcription rate to a time varying aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}.
+$x(t)$ behaves analogously for the mRNA.
+
+\begin{table}[h]
+\renewcommand{\arraystretch}{1.3}
+\caption{Model Parameters}
+\label{ModelParameters}
+\centering
+\begin{tabular}{| l | l | l |}
+\hline Parameter &  Units & Definition  \\
+\hline \hline N & & plasmid copy number  \\
+\hline $z_{0}$ &  AFU & baseline fluorescence  \\
+\hline $\alpha_{L}$ & nM/min & Max translation rate of Pllac01 promoter  \\
+\hline $\alpha_{T}$  &  nM/min  & Max translation rate of Pltet01 promoter \\
+\hline $f_{L}$ &  & Max llacI fold  \\
+\hline $f_{T}$ &  & lltet fold\\
+\hline $\delta_{g}$  & /min  & gfp degradation rate  \\
+\hline $\gamma$ &  /min & gfp maturation rate  \\
+\hline $v_{z}$ & nM/min & degradation constant of clpx  \\
+\hline $K_{z}$   &   nM/min & Dissociation constant of clpx  \\
+\hline $\theta$  &   nM/AFU & ratio between gfp amount and fluorensence  \\
+\hline $\mu$ &  /min & dilution rate  \\
+\hline $\delta_{m}$ &  /min & mrna degredation rate  \\
+\hline $\delta_{s}$ &  /min & srna degredation rate  \\
+\hline $\delta_{sm}$ &  /min & unstable s:mrna degredation rate  \\
+\hline $\delta_{c}$ &  /min & stable s;m degredation rate  \\
+\hline $k_{on}$ &   /min & s+m $\rightarrow$ s:m rate \\
+\hline $k_{off}$ &  /min & s+m $\leftarrow$ s:m rate \\
+\hline $k_{hyb}$ &  /min & s:m $\rightarrow$ c rate \\
+\hline $\beta$ &   /min & basal translation rate \\
+\hline $f_{s}$ & & srna repression fold \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+\renewcommand{\arraystretch}{1.3}
+\caption{State Variables}
+\label{StateVariables}
+\centering
+\begin{tabular}{| l | l |}
+\hline State variable &  Definition  \\
+\hline\hline $s$ &  Definition  \\
+\hline $m$ &  Definition  \\
+\hline $s:m$ &  Definition  \\
+\hline $c$ &  Definition  \\
+\hline $p$ &  Definition  \\
+\hline $g$ &  Definition  \\
+\hline $z$ &  Definition  \\
+\hline $y(t)$ &  Definition  \\
+\hline $x(t)$ &  Definition  \\
+\hline
+\end{tabular}
+\end{table}
 

### 2015-08-25 MScWriteup.tex
still writing introductory sections

@@
 The report is structured as follows.
-In the remainder of this section we review the sRNA regulatory system we will consider.
+In the remainder of this section we review the sRNA regulatory system we will consider, and discuss recent single cell fluorescence experiments performed on this system.
 In section \ref{Results and Discussion} we propose a set of ODE's to model the system, and estimate its unknown parameters by fitting to time series data.
@@
 \caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS.
-In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{LtetO-1}$ and $P_{LlacO-1}$.
+In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$.
 These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E.
@@
 \ref{ANDGate}).
-In this system, transcription of the designed sRNA and mRNA are placed under the control of promoter regions, $P_{LtetO-1}$ and $P_{LlacO-1}$ \cite{Lutz1997}.
+In this system, transcription of the designed sRNA and mRNA are placed under the control of promoter regions, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}.
 These are in turn controlled by two transcriptional repressors, TetR and LacI, which are naturally present in the strain of \text{E.
@@
 By changing which sRNA-mRNA device is used in the system, it would also allow exploration of the relationship between the thermodynamic properties of each device, and the model's rate constants.
+
+\subsection{Single Cell Fluorescence Data}
+
+Recent experiments have used timelapse microscopy to observe bacteria as they are periodically forced with a varying aTc or IPTG concentration \cite{Jaramillo}.
+The bacteria are grown a single layer thick in rows of chambers.
+A medium constantly flows through these chambers, allowing normal feeding of the bacteria, and the introduction of aTc or IPTG.
+The chambers are monitored with software which traces each cell, allowing timeseries of individual cell fluorescences to be recorded.
+
+\begin{figure}[H]
+\centering
+\includegraphics[trim = 0 0 0  85 ,page=17 ,clip = true, scale=0.3]{WCPM_Warwick-talk45-220115.pdf}
+\centering
+\includegraphics[trim = 0 0 0  90 ,page=18 ,clip = true, scale=0.3]{WCPM_Warwick-talk45-220115.pdf}
+\caption{A schematic of an experimental setup which allows single cell fluorescences to be recorded over time.
+Bacteria are grown in chambers, with a constant supply of medium flowing through them from the media inlet.
+This inlet also allows the introduction of aTc IPTG.
+Software then traces these cells, allowing fluorescence timeseries to be built.
+Image reproduced from \cite{Jaramillo}.}
+\label{reactionscheme}
+\end{figure}
 
@@
 We propose a simple model, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}, to describe the system.
-Tables \ref{ModelParameters},\ref{StateVariables} give complete descriptions of parameter and state variable meanings.
+Tables \ref{ModelParameters}, \ref{StateVariables} give complete descriptions of the parameters and state variables.
 
@@
 The initial binding is modelled as a reversible reaction with forward and backward rates $k_{on}$ and $k_{off}$.
+
 \begin{align*}
-sRNA \chemsign+ mRNA\chemrel[$k_{on}$][$k_{off}$]{<>} sRNA:mRNA_{unstable} 
-\end{align*}\\
-\noindent
+\mathrm{
+sRNA \chemsign+ mRNA\chemrel[$k_{on}$][$k_{off}$]{<>} sRNA:mRNA\textsubscript{unstable}
+}
+\end{align*}
 after which the stabilization is modelled as an irreversible reaction with rate $k_{hyb}$.
+
 \begin{align*}
-sRNA:mRNA_{unstable} \chemrel[$k_{hyb}$]{->} sRNA:mRNA_{stable} 
+\mathrm{
+sRNA:mRNA\textsubscript{unstable} \chemrel[$k_{hyb}$]{->} sRNA:mRNA\textsubscript{stable} 
+}
 \end{align*}
@@
 
- Control of the system by aTc and IPTG are modelled by $y(t)$ and $x(t)$ in eqs.\ref{eq:s}, \ref{eq:m}.
+ Control of the system by aTc and IPTG is modelled by $y(t)$ and $x(t)$ in eqs.\ref{eq:s}-\ref{eq:m}.
 $y(t)$ models the response of the sRNA transcription rate to a time varying aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}.
-$x(t)$ behaves analogously for the mRNA.
+Thus the forcing may vary between
+ $N\frac{1}{f_{T}}$ and $N\frac{\alpha_{T}}{f}$, where $\alpha_{T}$ is the maximal transcription rate of the $P_{\mathrm{LtetO-1}}$ promoter.
+When engineering the system, many copies of the $P_{\mathrm{LtetO-1}}$ promoter and section of DNA coding for sRNA may be placed in the plasmid DNA - this is modelled by the copy number, $N$.
+Identical considerations hold for $x(t)$ and IPTG concentrations.
+
+We explicitly model translation as a simple one step process in eqs.
+\ref{eq:p}-\ref{eq:z}.
+There is a small rate of translation of the self repressed mRNA \cite{Rodrigo2012}, which is modelled at rate $\beta$, and a larger one for translation of the stable complex, $\beta f_s$.
+Here $f_s$ represents the fractional change in translation rate between the repressed mRNA and the unrepressed complex.
+
+\begin{align*}
+\mathrm{
+sRNA:mRNA\textsubscript{Stable} \chemrel[$\beta f_{s}$]{->} GFP\textsubscript{Immature}
+}
+\end{align*}
+
+\begin{align*}
+\mathrm{
+mRNA \chemrel[$\beta$ ]{->} GFP\textsubscript{Immature}
+}
+\end{align*}
+
+Initially, the translated GFP is in an immature state, and will not fluoresce. To account for this, we include a maturation rate, $\gamma$
+
+\begin{align*}
+\mathrm{
+GFP\textsubscript{Immature}  \chemrel[$\gamma$]{->} GFP\textsubscript{Mature}
+}
+\end{align*}
+
+Degradation of the immature and mature GFP is modelled in two ways.
+Firstly a generic degradation rate $\delta_{g}$, assumed identical for the mature and immature species, with the dilution rate $\mu$ shared by all species.
+\todo[inline]{para about clpX}
+Finally, eq. \ref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence.
 
@@
 \centering
-\begin{tabular}{| l | l | l |}
-\hline Parameter &  Units & Definition  \\
-\hline \hline N & & plasmid copy number  \\
-\hline $z_{0}$ &  AFU & baseline fluorescence  \\
-\hline $\alpha_{L}$ & nM/min & Max translation rate of Pllac01 promoter  \\
-\hline $\alpha_{T}$  &  nM/min  & Max translation rate of Pltet01 promoter \\
-\hline $f_{L}$ &  & Max llacI fold  \\
-\hline $f_{T}$ &  & lltet fold\\
-\hline $\delta_{g}$  & /min  & gfp degradation rate  \\
-\hline $\gamma$ &  /min & gfp maturation rate  \\
+\begin{tabular}{| l | l | p{0.7\linewidth} |}
+\hline \textbf{Parameter} &  \textbf{Units} & \textbf{Definition}  \\
+\hline \hline N & & Number of copies of promoter existing on plasmid DNA  \\
+\hline $z_{0}$ &  AFU & Baseline experimental fluorescence  \\
+\hline $\alpha_{L}$ & nM/min & Maximal transcription rate of $P_\mathrm{LlacO-1}$ promoter\\
+\hline $\alpha_{T}$  &  nM/min  & Maximal transcription rate of $P_{\mathrm{LtetO-1}}$ promoter \\
+\hline $f_{L}$ &  & Unitless ratio between repressed and unrepressed $P_\mathrm{LlacO-1}$ transcription rate   \\ 
+\hline $f_{T}$ &  & Unitless ratio between repressed and unrepressed $P_{\mathrm{LtetO-1}}$ transcription rate  \\
+\hline $\delta_{g}$  & /min  & GFP degradation rate  \\
+\hline $\gamma$ &  /min & GFP maturation rate  \\
 \hline $v_{z}$ & nM/min & degradation constant of clpx  \\
 \hline $K_{z}$   &   nM/min & Dissociation constant of clpx  \\
-\hline $\theta$  &   nM/AFU & ratio between gfp amount and fluorensence  \\
-\hline $\mu$ &  /min & dilution rate  \\
-\hline $\delta_{m}$ &  /min & mrna degredation rate  \\
-\hline $\delta_{s}$ &  /min & srna degredation rate  \\
-\hline $\delta_{sm}$ &  /min & unstable s:mrna degredation rate  \\
-\hline $\delta_{c}$ &  /min & stable s;m degredation rate  \\
-\hline $k_{on}$ &   /min & s+m $\rightarrow$ s:m rate \\
-\hline $k_{off}$ &  /min & s+m $\leftarrow$ s:m rate \\
-\hline $k_{hyb}$ &  /min & s:m $\rightarrow$ c rate \\
-\hline $\beta$ &   /min & basal translation rate \\
-\hline $f_{s}$ & & srna repression fold \\
+\hline $\theta$  &   nM/AFU & Ratio between GFP concentration and observed fluoresence  \\
+\hline $\mu$ &  /min & Dilution rate  \\
+\hline $\delta_{m}$ &  /min & mRNA degredation rate  \\
+\hline $\delta_{s}$ &  /min & sRNA degredation rate  \\
+\hline $\delta_{sm}$ &  /min & Unstable sRNA:mRNA degradation rate  \\
+\hline $\delta_{c}$ &  /min & Stable sRNA:mRNA degradation rate  \\
+\hline $k_{on}$ &   /min & sRNA:mRNA binding rate \\
+\hline $k_{off}$ &  /min & sRNA:mRNA unbinding rate \\
+\hline $k_{hyb}$ &  /min & sRNA:mRNA hybridization rate \\
+\hline $\beta$ &   /min & Baseline translation rate of repressed mRNA \\
+\hline $f_{s}$ & & Ratio of repressed mRNA to unrepressed complex translation rate. \\
 \hline
@@
 
-\begin{table}
+\begin{table}[h]
 \renewcommand{\arraystretch}{1.3}
@@
 \centering
-\begin{tabular}{| l | l |}
-\hline State variable &  Definition  \\
-\hline\hline $s$ &  Definition  \\
-\hline $m$ &  Definition  \\
-\hline $s:m$ &  Definition  \\
-\hline $c$ &  Definition  \\
-\hline $p$ &  Definition  \\
-\hline $g$ &  Definition  \\
-\hline $z$ &  Definition  \\
-\hline $y(t)$ &  Definition  \\
-\hline $x(t)$ &  Definition  \\
+\begin{tabular}{| l | l | l|}
+\hline \textbf{State variable} & Units &  \textbf{Definition}  \\
+\hline\hline $s$  & nM & sRNA concentration \\
+\hline $m$ & nM & mRNA concentration  \\
+\hline $s:m$ &  nM & Unstable sRNA:mRNA complex concentration  \\
+\hline $c$ &  nM & Stable sRNA:mRNA complex concentration  \\
+\hline $p$ & nM & Immature GFP concentration  \\
+\hline $g$ &  nm & Mature GFP concentration  \\
+\hline $z$ & AFU & Observed fluorescence  \\
+\hline $y(t)$ & & Unitless aTc forcing function  \\
+\hline $x(t)$ &  & Unitless IPTG forcing function  \\
 \hline

### 2015-08-25 MScWriteup.tex
beginning the parameter estimation section

@@
 A medium constantly flows through these chambers, allowing normal feeding of the bacteria, and the introduction of aTc or IPTG.
-The chambers are monitored with software which traces each cell, allowing timeseries of individual cell fluorescences to be recorded.
+The chambers are monitored with software which traces the position of each cell over time, allowing timeseries of individual cell fluorescences to be recorded.
 
@@
 \centering
-\begin{tabular}{| l | l | p{0.7\linewidth} |}
+\begin{tabular}{| l | l | p{0.6\linewidth} |}
 \hline \textbf{Parameter} &  \textbf{Units} & \textbf{Definition}  \\
@@
 \subsection{Parameter Estimation}
- \begin{itemize}
-\item basically the whole project.
-\cite{Isaacs2006}
- \end{itemize}
+Our next goal is to estimate the unknown parameters of this model, given the available fluorescence time series data, by fitting predicted time series from the model to the data.
+Typically, this is done by minimising the least squares error between model prediction and the experimental data.
+Suppose we have some ODE model of our system
+
+\begin{align}
+\frac{\mathrm{d}\mathbf{y}}{\mathrm{d}t} = \mathbf{f}(\mathbf{y},\boldsymbol{\theta},t)
+\end{align}
+where $\mathbf{y}$ is our state vector, $\boldsymbol{\theta}$ is a vector of model parameters, and $t$ is time (representing the fact that the model may have some explicit time dependence).
+The model may be integrated numerically, giving a prediction $\mathbf{y}(t,\boldsymbol{\theta})$.
+An error between the model prediction and an experimental time series is defined as
+\begin{align}
+J(\boldsymbol{\theta}) = \sum_{i =1}^{N} (\mathbf{y}_{\mathrm{exp}}(t_{i}) - \mathbf{y}(t_{i},\boldsymbol{\theta}))^2
+\end{align}
+where the experimental timeseries, $\mathbf{y}_{\mathrm{exp}}(t_{i})$ is evaluated at timepoints $t_{i}$, $i = 1 \hdots N$.
+This error function defines a landscape in $\boldsymbol{\theta}$ space, and we seek to minimise it by varying $\boldsymbol{\theta}$.
+In our case, we do not have experimental data on the full state vector, but only one component of it - the observed fluorescence, $z(t)$.
+In addition, rather than a single experimental run, we have many, corresponding to a timeseries from each cell.
+We incorporate this by fitting to the experimental mean of the data, and only minimising over the observed component.
+Our minimisation problem is thus
+\begin{align}
+\min_{\boldsymbol{\theta}} \sum_{i =1}^{N} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2
+\end{align}
+The next step is performing the minimisation.
+In general, the landscape defined by the error function is multimodal, and may be very rugged.
+\footnote{this is true even if ODE model is linear in its parameters, as is almost the case for us - though the ODE model is linear, the resulting solutions are in general not.
+A counterexample is the harmonic oscillator.} A local optimisation algorithm will
 

### 2015-08-26 MScWriteup.tex
more modelling talk

@@
 
+The data we will consider consists of three sets of individual cell timeseries, labelled \textit{13\_9}, \textit{14\_7} and \textit{14\_9}.
+They correspond to three different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response, and aTc concentrations were changed with varying periods.
+Section \ref{Initial  Experimental Data} shows the full datasets, with their forcing functions.
+
 %%%
@@
 \frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \label{eq:g} \\
-z &= z_{0} +\frac{g}{\theta} \label{eq:z}
+z &= z_{0} +\frac{g}{\Theta} \label{eq:z}
 \end{align}
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Model Parameters}
+\caption{Model Parameters (those to be estimated shown in bold)}
 \label{ModelParameters}
@@
 \hline $K_{z}$   &   nM/min & Dissociation constant of clpx  \\
-\hline $\theta$  &   nM/AFU & Ratio between GFP concentration and observed fluoresence  \\
-\hline $\mu$ &  /min & Dilution rate  \\
-\hline $\delta_{m}$ &  /min & mRNA degredation rate  \\
-\hline $\delta_{s}$ &  /min & sRNA degredation rate  \\
+\hline $\boldsymbol{\Theta}$  &   nM/AFU & Ratio between GFP concentration and observed fluoresence  \\
+\hline $\boldsymbol{\mu}$ &  /min & Dilution rate  \\
+\hline $\boldsymbol{\delta_{m}}$ &  /min & mRNA degredation rate  \\
+\hline $\boldsymbol{\delta_{s}}$ &  /min & sRNA degredation rate  \\
 \hline $\delta_{sm}$ &  /min & Unstable sRNA:mRNA degradation rate  \\
 \hline $\delta_{c}$ &  /min & Stable sRNA:mRNA degradation rate  \\
-\hline $k_{on}$ &   /min & sRNA:mRNA binding rate \\
-\hline $k_{off}$ &  /min & sRNA:mRNA unbinding rate \\
-\hline $k_{hyb}$ &  /min & sRNA:mRNA hybridization rate \\
-\hline $\beta$ &   /min & Baseline translation rate of repressed mRNA \\
-\hline $f_{s}$ & & Ratio of repressed mRNA to unrepressed complex translation rate. \\
+\hline $\boldsymbol{k_{on}}$ &   /min & sRNA:mRNA binding rate \\
+\hline $\boldsymbol{k_{off}}$ &  /min & sRNA:mRNA unbinding rate \\
+\hline $\boldsymbol{k_{hyb}}$ &  /min & sRNA:mRNA hybridization rate \\
+\hline $\boldsymbol{\beta}$ &   /min & Baseline translation rate of repressed mRNA \\
+\hline $\boldsymbol{f_{s}}$ & & Ratio of repressed mRNA to unrepressed complex translation rate. \\
 \hline
@@
 Our next goal is to estimate the unknown parameters of this model, given the available fluorescence time series data, by fitting predicted time series from the model to the data.
-Typically, this is done by minimising the least squares error between model prediction and the experimental data.
+Typically, this is done by minimising the least squares error between model prediction and the experimental data \cite{Brewer2008,Algorithms2003, Hu2015}.
 Suppose we have some ODE model of our system
@@
 \footnote{this is true even if ODE model is linear in its parameters, as is almost the case for us - though the ODE model is linear, the resulting solutions are in general not.
-A counterexample is the harmonic oscillator.} A local optimisation algorithm will
+A counterexample is the harmonic oscillator.} A local optimisation algorithm will often get stuck in local minima.
+To try and surmount this problem, \cite{Algorithms2003} suggests the use of a global optimisation algorithm, and in particular recommends several Evolutionary Algorithms, of which we choose one, the CMA-ES \cite{Hansen2006,Hansen2011}.
+
+In order to reduce the dimensionality of our search space, we can perform a literature search for existing values of some of our parameters, simplify our model to remove others, and place bounds on those that remain.
+Section \ref{Parameter literature review} contains a list of parameter values found in the literature, where available, and their reference, as well as initial bounds placed on parameters not found in the literature.
+To further reduce the search space, we assume that $\delta_m$, $\delta_{sm}$ and $\delta_c$ all take similar values, and set them equal.
+
+ After this is done, we are left with a 9 dimensional search space, bounded by a hypercube (parameters to be estimated are shown in bold in \ref{ModelParameters}).
+We begin by fitting two of the datasets, \textit{13\_9} and \textit{14\_7}
 
@@
 
+\appendix
+
+\section{Initial Experimental Data}
+\label{Initial Experimental Data}
+
+\begin{figure}[h]
+\centering
+\includegraphics[trim = 0 320 0 300 , scale = 0.4, clip = true]{13_9}
+\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{Rodrigo2012}}
+\label{reactionscheme}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[trim = 0 300 0 250 , scale = 0.4, clip = true]{14_7}
+\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{Rodrigo2012}}
+\label{reactionscheme}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[trim = 0 200 0 270 , scale = 0.4, clip = true]{14_9}
+\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{Rodrigo2012}}
+\label{reactionscheme}
+\end{figure}
+
+\section{Parameter literature review}
+\label{Parameter literature review}
+
+\begin{table}[h]
+\renewcommand{\arraystretch}{1.3}
+\caption{Model Parameters}
+\centering
+\begin{tabular}{| l | l | p{0.3\linewidth} | p{0.1\linewidth} | p{0.05\linewidth}}
+\hline \textbf{Parameter} &  \textbf{Value} & \textbf{Definition} & \textbf{Reference} & \textbf{Initial Bounds}  \\
+\hline \hline N & 300 & Number of copies of promoter existing on plasmid DNA & Experimentally set  \\
+\hline $z_{0}$ &   9 AFU & Baseline experimental fluorescence & Experimentally determined  \\
+\hline $\alpha_{L}$ & 11 nM/min & Maximal transcription rate of $P_\mathrm{LlacO-1}$ promoter & \cite{Lutz1997}\\
+\hline $\alpha_{T}$  & 11 nM/min  & Maximal transcription rate of $P_{\mathrm{LtetO-1}}$ promoter  & \cite{Lutz1997} \\
+\hline $f_{L}$ & 620 & Unitless ratio between repressed and unrepressed $P_\mathrm{LlacO-1}$ transcription rate  & \cite{Lutz1997}  \\ 
+\hline $f_{T}$ & 2535 & Unitless ratio between repressed and unrepressed $P_{\mathrm{LtetO-1}}$ transcription rate   & \cite{Lutz1997} \\
+\hline $\delta_{g}$  & 0.0005 /min  & GFP degradation rate & \cite{Andersen1998} \\
+\hline $\gamma$ &  0.132 /min & GFP maturation rate & \cite{Iizuka2011}  \\
+\hline $v_{z}$ & 100 nM/min & degradation constant of clpx & \cite{Hersch2004} \\
+\hline $K_{z}$   &   75 nM/min & Dissociation constant of clpx & \cite{Hersch2004} \\
+\hline $\boldsymbol{\Theta}$  &   nM/AFU & Ratio between GFP concentration and observed fluoresence & & 300 - 1000  \\
+\hline $\boldsymbol{\mu}$ &  /min & Dilution rate & & 0.001-0.05\\
+\hline $\boldsymbol{\delta_{m}}$ &  /min & mRNA degredation rate & & 1 - $10^5$ \\
+\hline $\boldsymbol{\delta_{s}}$ &  /min & sRNA degredation rate & & 1 - $10^3$ \\
+\hline $\delta_{sm}$ &  /min & Unstable sRNA:mRNA degradation rate & &  Set to $\delta_m$\\
+\hline $\delta_{c}$ &  /min & Stable sRNA:mRNA degradation rate & & Set to $\delta_m$ \\
+\hline $\boldsymbol{k_{on}}$ &   /min & sRNA:mRNA binding rate & & 100 - $10^7$\\
+\hline $\boldsymbol{k_{off}}$ &  /min & sRNA:mRNA unbinding rate & & 1 -$10^8$\\
+\hline $\boldsymbol{k_{hyb}}$ &  /min & sRNA:mRNA hybridization rate & & 1 - $10^4$ \\
+\hline $\boldsymbol{\beta}$ &   /min & Baseline translation rate of repressed mRNA & & 0.0001 - 10\\
+\hline $\boldsymbol{f_{s}}$ & & Ratio of repressed mRNA to unrepressed complex translation rate. & & 0.1 - $10^4$\\
+\hline
+\end{tabular}
+\end{table}
+
 \end{document}

### 2015-08-27 MScWriteup.tex
one big figure made

@@
 
-\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e}
-\usepackage[caption=false]{subfig}
+\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e, epstopdf, caption, subcaption}
 
@@
 
-Recent experiments have used timelapse microscopy to observe bacteria as they are periodically forced with a varying aTc or IPTG concentration \cite{Jaramillo}.
+Recent experiments have used timelapse microscopy to observe the fluoresence of bacteria which contain the above device, as they are periodically forced with a varying aTc or IPTG concentration \cite{Jaramillo}.
+
 The bacteria are grown a single layer thick in rows of chambers.
@@
 The data we will consider consists of three sets of individual cell timeseries, labelled \textit{13\_9}, \textit{14\_7} and \textit{14\_9}.
-They correspond to three different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response, and aTc concentrations were changed with varying periods.
+They correspond to three different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response, and aTc concentrations were varied periodically.
 Section \ref{Initial  Experimental Data} shows the full datasets, with their forcing functions.
@@
  After this is done, we are left with a 9 dimensional search space, bounded by a hypercube (parameters to be estimated are shown in bold in \ref{ModelParameters}).
-We begin by fitting two of the datasets, \textit{13\_9} and \textit{14\_7}
+We begin by fitting two of the datasets, \textit{13\_9} and \textit{14\_7}, by choosing 200 points, uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
+Results are shown in Figs.
+\ref{InitialResults}, \ref{hist_f}.
+Fig.
+\ref{InitialResults} demonstrates that the model is capable of quantitatively capturing the data - however, the fitting also suggests that some parameters are not tightly constrained, taking values right across the initial bounding range specified.
+
+ \begin{itemize}
+ \item discuss estimatbility criteria- uniqeuess and sensitivity
+ \item put in some sensitivity analysis results
+ \item discuss model fixed points
+ \item discuss systematic paramater shift.
+ \end{itemize}
+
+\begin{figure*}
+    \begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.35, clip = true, trim = 80 0 80 0]{13_9_hist.eps}
+        \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset. }
+    \end{subfigure}
+    \begin{subfigure}[c]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.38,clip = true, trim = 60 12 80 0]{14_7_hist.eps}
+        \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{14\_7} dataset. }
+    \end{subfigure}
+
+        \begin{subfigure}[c]{0.49\textwidth}
+        \centering
+    \includegraphics[scale = 0.25, clip = true, trim = 100 0 100 0]{13_9_heatmap.eps}
+        \caption{Correlation matrix of estimated parameter sets, from the  \textit{13\_9} dataset. }
+    \end{subfigure}
+    \begin{subfigure}[c]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.25, clip = true, trim = 80 0 80 0]{14_7_heatmap.eps}
+        \caption{Correlation matrix of estimated parameter sets, from the  \textit{14\_7} dataset.}
+    \end{subfigure}
+
+    \begin{subfigure}[c]{0.49\textwidth}
+    \centering
+    \includegraphics[scale = 0.37]{13_9_bestPlot}
+        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset. A}
+    \end{subfigure}
+    \begin{subfigure}[c]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.37]{14_7_bestPlot}
+        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{14\_7} dataset.}
+    \end{subfigure}
+    \caption{Parameter estimates, inter-parameter correlation values, and model predictions for the \textit{13\_9} and \textit{14\_7} datasets.
+Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response.
+The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response}
+\label{InitialResults}
+\end{figure*}
+
+\clearpage
 
@@
 \label{Conclusions and Further work}
+
+\todo[inline]{some stuff}
 
@@
 
-\appendix
+\clearpage
+\onecolumn
+
+\appendices
+\renewcommand\thefigure{\thesection.\arabic{figure}}
 
@@
 \label{Initial Experimental Data}
+\setcounter{figure}{0}
 
@@
 \centering
-\includegraphics[trim = 0 320 0 300 , scale = 0.4, clip = true]{13_9}
-\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{Rodrigo2012}}
+\includegraphics[trim = 0 310 0 300 , scale = 0.5, clip = true]{13_9}
+\caption{The \textit{13\_9} dataset, with aTc forcing shown.
+IPTG concentration is constant at a level which saturates the cell's response.
+Note the forcing curve's height is schematic - aTc concentation is switched between off, and a level which saturates the cell's response}
+\label{reactionscheme}
+\end{figure}
+
+%trim=l b r t
+\begin{figure}[h]
+\centering
+\includegraphics[trim = 0 270 0 360 , scale = 0.5, clip = true]{14_7}
+\caption{The \textit{14\_7} dataset, with aTc forcing shown.
+IPTG concentration is constant at a level which saturates the cell's response.
+Note the forcing curve's height is schematic - aTc concentation is switched between off, and a level which saturates the cell's response}
 \label{reactionscheme}
@@
 \centering
-\includegraphics[trim = 0 300 0 250 , scale = 0.4, clip = true]{14_7}
-\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{Rodrigo2012}}
+\includegraphics[trim = 0 330 0 270 , scale = 0.5, clip = true]{14_9}
+\caption{The \textit{14\_9} dataset, with aTc forcing shown.
+IPTG concentration is constant at a level which saturates the cell's response.
+Note the forcing curve's height is schematic - aTc concentation is switched between off, and a level which saturates the cell's response}
 \label{reactionscheme}
@@
 
-\begin{figure}[h]
+\section{Histogram of Initial Fit Error values}
+\label{Initial fit Error}
+\setcounter{figure}{0}
+
+\begin{figure}[H]
+    \begin{subfigure}[c]{\textwidth}
     \centering
-\includegraphics[trim = 0 200 0 270 , scale = 0.4, clip = true]{14_9}
-\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{Rodrigo2012}}
-\label{reactionscheme}
+        \includegraphics[scale = 0.5]{13_9_f_hist}
+        \caption{Error values found from 200 initial parameter estimates, \textit{13\_9 dataset}.}
+    \end{subfigure}
+
+    \begin{subfigure}[c]{\textwidth}
+    \centering
+        \includegraphics[scale = 0.5]{14_7_f_hist}
+        \caption{Error values found from 200 initial parameter estimates, \textit{14\_7 dataset}}
+    \end{subfigure}
+    \caption{}
+\label{hist_f}
 \end{figure}
@@
 \label{Parameter literature review}
+\setcounter{figure}{0}
 
-\begin{table}[h]
+\begin{table}[H]
 \renewcommand{\arraystretch}{1.3}
-\caption{Model Parameters}
+\caption{Literature references, or initial rough bounds, on parameter values, with those to be estimated shown in bold}
 \centering
-\begin{tabular}{| l | l | p{0.3\linewidth} | p{0.1\linewidth} | p{0.05\linewidth}}
+\begin{tabular}{| l | l | p{0.3\linewidth} | p{0.1\linewidth} | p{0.1\linewidth} |}
 \hline \textbf{Parameter} &  \textbf{Value} & \textbf{Definition} & \textbf{Reference} & \textbf{Initial Bounds}  \\
-\hline \hline N & 300 & Number of copies of promoter existing on plasmid DNA & Experimentally set  \\
-\hline $z_{0}$ &   9 AFU & Baseline experimental fluorescence & Experimentally determined  \\
-\hline $\alpha_{L}$ & 11 nM/min & Maximal transcription rate of $P_\mathrm{LlacO-1}$ promoter & \cite{Lutz1997}\\
-\hline $\alpha_{T}$  & 11 nM/min  & Maximal transcription rate of $P_{\mathrm{LtetO-1}}$ promoter  & \cite{Lutz1997} \\
-\hline $f_{L}$ & 620 & Unitless ratio between repressed and unrepressed $P_\mathrm{LlacO-1}$ transcription rate  & \cite{Lutz1997}  \\ 
-\hline $f_{T}$ & 2535 & Unitless ratio between repressed and unrepressed $P_{\mathrm{LtetO-1}}$ transcription rate   & \cite{Lutz1997} \\
-\hline $\delta_{g}$  & 0.0005 /min  & GFP degradation rate & \cite{Andersen1998} \\
-\hline $\gamma$ &  0.132 /min & GFP maturation rate & \cite{Iizuka2011}  \\
-\hline $v_{z}$ & 100 nM/min & degradation constant of clpx & \cite{Hersch2004} \\
-\hline $K_{z}$   &   75 nM/min & Dissociation constant of clpx & \cite{Hersch2004} \\
+\hline \hline N & 300 & Number of copies of promoter existing on plasmid DNA & Experimentally set & \\
+\hline $z_{0}$ &   9 AFU & Baseline experimental fluorescence & Experimentally determined &  \\
+\hline $\alpha_{L}$ & 11 nM/min & Maximal transcription rate of $P_\mathrm{LlacO-1}$ promoter & \cite{Lutz1997} &\\
+\hline $\alpha_{T}$  & 11 nM/min  & Maximal transcription rate of $P_{\mathrm{LtetO-1}}$ promoter  & \cite{Lutz1997}& \\
+\hline $f_{L}$ & 620 & Unitless ratio between repressed and unrepressed $P_\mathrm{LlacO-1}$ transcription rate  & \cite{Lutz1997}  &  \\ 
+\hline $f_{T}$ & 2535 & Unitless ratio between repressed and unrepressed $P_{\mathrm{LtetO-1}}$ transcription rate   & \cite{Lutz1997} &\\
+\hline $\delta_{g}$  & 0.0005 /min  & GFP degradation rate & \cite{Andersen1998} &\\
+\hline $\gamma$ &  0.132 /min & GFP maturation rate & \cite{Iizuka2011} & \\
+\hline $v_{z}$ & 100 nM/min & degradation constant of clpx & \cite{Hersch2004} &\\
+\hline $K_{z}$   &   75 nM/min & Dissociation constant of clpx & \cite{Hersch2004} &\\
 \hline $\boldsymbol{\Theta}$  &   nM/AFU & Ratio between GFP concentration and observed fluoresence & & 300 - 1000  \\

### 2015-08-30 MScWriteup.tex
adding sensitivity analysis stuff

@@
 \end{align}
-where $\mathbf{y}$ is our state vector, $\boldsymbol{\theta}$ is a vector of model parameters, and $t$ is time (representing the fact that the model may have some explicit time dependence).
+where $\mathbf{y}$ is our state vector, $\boldsymbol{\theta}$ is a vector of model parameters, and $t$ is time.
 The model may be integrated numerically, giving a prediction $\mathbf{y}(t,\boldsymbol{\theta})$.
@@
 To further reduce the search space, we assume that $\delta_m$, $\delta_{sm}$ and $\delta_c$ all take similar values, and set them equal.
+After this is done, we are left with a 9 dimensional search space, bounded by a hypercube (parameters to be estimated are shown in bold in \ref{ModelParameters}).
 
- After this is done, we are left with a 9 dimensional search space, bounded by a hypercube (parameters to be estimated are shown in bold in \ref{ModelParameters}).
  We begin by fitting two of the datasets, \textit{13\_9} and \textit{14\_7}, by choosing 200 points, uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
@@
 \ref{InitialResults} demonstrates that the model is capable of quantitatively capturing the data - however, the fitting also suggests that some parameters are not tightly constrained, taking values right across the initial bounding range specified.
+Further, the correlation matrices in Fig.
+\ref{InitialResults} indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
+Referring to eq.
+\ref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
+The results suggest a fitness landscape relatively flat to perturbations in certain combinations of parameters, and indicate we may have difficulty obtaining unique estimates of the model parameters.
+
+ There are two main reasons why a parameter may not be estimable  \cite{Mclean2012,Yao2003, Beck}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others - changes in one parameter may be balanced by adjustments to the values of several others.
+
+ We may investigate whether our model suffers from such problems by performing a local sensitivity analysis about one of the solutions found in our initial parameter estimation.
+We numerically estimate the sensitivity matrix, $S$:
+
+ \begin{align}
+S_{ij} = \hat{\theta}_{j} \frac{\partial z}{ \partial \theta_{j}}\Bigr|_{t_{i}}
+\end{align}
+
+where $S_{ij}$ is the derivative of the observed fluorescence, $z$, with parameter $\theta_{j}$, evaluated at timepoint $t_{i}$.
+This matrix contains information about how sensitive $z$ is to perturbations in parameter values, and at what times it is most sensitive.
+$\hat{\theta}_{j}$ is the value of the parameter that the derivative is being evaluated at.
+It is included to set the scale that parameters may vary at, to ensure that apparently small sensitivity values do not result from a poor choice of units.
+
+Fig.
+\ref{SensitivityMatrix} shows plots of each column of $S$, evaluated at the parameter set giving the lowest error in the \textit{13\_9} dataset - these show the evolution of the sensitivity matrix over time.
+Fig.
+\ref{SensitivityMatrix_unscaled} shows them unscaled, Fig.
+\ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$, so that their shapes may be more easily compared.
+We see that many of the parameters give sensitivity curves of similar shapes - the effects of perturbing any one of these parameters all look similar in terms of model output.
+It is unsurprising that some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
+Note that the sensitivity curve that looks least similar to the others in Fig.
+\ref{SensitivityMatrix_scaled} - $\mu$ - corresponds to a relatively tightly estimated value in Fig.
+\ref{InitialResults}.
+
+\begin{figure}[h]	
+	\begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.3]{Sensitivity_unscaled}
+        \caption{Unscaled}
+        \label{SensitivityMatrix_unscaled} 
+    \end{subfigure}
+    \begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.3]{Sensitivity_scaled}
+        \caption{Scaled}
+                \label{SensitivityMatrix_scaled} 
+    \end{subfigure}
+    \caption{Sensitivity co-efficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset.
+\ref{SensitivityMatrix_unscaled} shows them unscaled, \ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$.}
+\label{SensitivityMatrix}   
+\end{figure}
 
@@
  \item discuss systematic paramater shift.
+ \item discuss cross - validation and predicitive ability of parameter sets found.
+ \item suggest two times, work out fixed point for all the solutions found, look at structure of soln,
+ \item re run from starting pararms.
  \end{itemize}

### 2015-08-30 MScWriteup.tex
rewrites to the intro

@@
 \label{sec: Intro}
-The process of gene expression can be briefly summarised as: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}).
+The process of gene expression can be summarised as follows: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}).
 This RNA molecule (known as messenger RNA, or mRNA) makes its way to a piece of cellular machinery called the Ribosome, which reads it, and makes a protein - which protein is made depends on the DNA sequence originally read (\textit{translation}).
@@
 This regulation allows the cell to control protein expression, and so cell behaviour, in response to various environmental cues.
-The natural cell machinery which performs this regulation offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.
+The natural cell machinery which performs it takes the form a genetic circuits - networks of interacting gene expression regulators.
+This genetic circuitry offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.
 
-As well as acting as the intermediate between DNA and protein, RNA molecules play direct and important roles in regulating cell behaviour \cite{Isaacs2006}.
-For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA's offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription.
+As well as acting as the intermediate between DNA and protein, RNA molecules play direct and important roles in regulating gene expression \cite{Isaacs2006}.
+For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription.
 In comparison to proteins, it is relatively straightforward to predict the structure and function of an RNA from its sequence using physiochemical models.
-Recently, this has been exploited to computationally design DNA sequences encoding synthetic sRNA's - small RNA's which do not code for a protein, but rather have some direct regulatory function -  with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.
+Recently, this has been exploited to computationally design synthetic sRNA's - small RNA's which do not code for a protein, but rather have some direct regulatory function -  with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.
 
-This report will focus on one such system, introduced in \cite{Rodrigo2012}.
-It will extend existing understanding of the system beyond the qualitative, by proposing a quantitative model of gene expression, in the form of a set of ODE's, and fitting it to available time series data to estimate the model's unknown parameters.
+This report will focus on one such sRNA system, introduced in \cite{Rodrigo2012}.
+It will extend existing understanding of it beyond the qualitative by proposing a quantitative model of gene expression, in the form of a set of ODE's, and fitting this model to available time series data to estimate its unknown parameters.
 

### 2015-09-02 MScWriteup.tex
more writing

@@
 
-\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e, epstopdf, caption, subcaption}
+\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e, epstopdf, caption, subcaption,centernot }
 
@@
 In the remainder of this section we review the sRNA regulatory system we will consider, and discuss recent single cell fluorescence experiments performed on this system.
-In section \ref{Results and Discussion} we propose a set of ODE's to model the system, and estimate its unknown parameters by fitting to time series data.
+In section \ref{Model Derivation} we propose a set of ODE's to model the system.
+In sect \ref{Parameter Estimation} and estimate its unknown parameters by fitting to time series data.
 Finally, in section \ref{Conclusions and Further work}, we conclude, and suggest directions for further work.
@@
 
-\section{Results and Discussion}
-\label{Results and Discussion}
+\section{Model Derivation}
+\label{Model Derivation}
 
-%NEW SUBSECTION%
-
-\subsection{Model Derivation}
 We propose a simple model, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}, to describe the system.
@@
 
-%NEW SUBSECTION%
+%%%
+%%%
+%%%
+%%%
+%%%
+%NEW SECTION%
+%%%
+%%%
+%%%
+%%%
 
-\subsection{Parameter Estimation}
+\section{Parameter Estimation}
+\label{Parameter Estimation}
+
 Our next goal is to estimate the unknown parameters of this model, given the available fluorescence time series data, by fitting predicted time series from the model to the data.
@@
 
+% NEW SUBSECTION
+
+\subsection{Initial Parameter Estimates}
+\label{Initial Parameter Estimates}
+
  We begin by fitting two of the datasets, \textit{13\_9} and \textit{14\_7}, by choosing 200 points, uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
@@
 \ref{InitialResults} demonstrates that the model is capable of quantitatively capturing the data - however, the fitting also suggests that some parameters are not tightly constrained, taking values right across the initial bounding range specified.
-Further, the correlation matrices in Fig.
-\ref{InitialResults} indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
+The correlation matrices in Fig.
+\ref{InitialResults} also indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
 Referring to eq.
@@
 
- There are two main reasons why a parameter may not be estimable  \cite{Mclean2012,Yao2003, Beck}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others - changes in one parameter may be balanced by adjustments to the values of several others.
+ We can test the predictive ability of our model by cross validating - taking the parameter values found in the fitting of one dataset, and using them to give model predictions for another.
+We thus take the parameter values giving the best fit for the  \textit{13\_9} dataset and use them to predict the \textit{14\_7} dataset, and vice versa.
+Results are shown in Fig.\ref{CrossValidation}.Though the predictions are reasonable, they are substantially worse than the predictions for the data they were trained on.
+The first reason for this may be experiment to experiment variability of parameter values \cite{Hu2015}.
+The second may be that the model is overfitting.
+
+ Taken together, these results suggest that some parameters may be inestimable, and that the data may be better described by a simplified model.
+
+ \subsection{Parameter Estimability}
+\label{Parameter Estimability}
+
+ There are two main reasons why a parameter may not be estimable  \cite{Mclean2012,Yao2003, Beck}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others.
 
@@
 
+It can be shown that if sensitivity co-efficients are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
+Related to this fact, a number of measures have been proposed to asses parameter estimability from the sensitivity matrix \cite{Mclean2012}
+
 Fig.
@@
 We see that many of the parameters give sensitivity curves of similar shapes - the effects of perturbing any one of these parameters all look similar in terms of model output.
-It is unsurprising that some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
+
+ It is unsurprising then that some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved, and as such that we should view estimates of these parameters with extreme caution.
 Note that the sensitivity curve that looks least similar to the others in Fig.
@@
 \ref{InitialResults}.
+ \todo[inline]{make more quantitivative with colinaerity idnex}
+
+Our inability to estimate some model parameters may stem from structural problems with the model (often termed \textit{structural identifiability}), in which two different parameter sets can give identical model predictions (eq.
+\ref{nonidentifiable}) .
+General methods exist to analyse the structural identifiability of an ODE model \cite{Jimenez-Hornero2008,Grewal1976 } \footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product can ever be estimated, a problem no amount of data can fix.}.
+
+ \begin{align}
+z(\boldsymbol{\theta}_{1},t) = z(\boldsymbol{\theta}_{2},t) \centernot\implies \theta_{1} = \theta_{2}
+\label{nonidentifiable}
+\end{align}
+
+Problems may also come from limitations in available data - differing experimental conditions may lift the strong correlations between parameter effects on model output (this is often termed \textit{practical identifiability} \cite{Mclean2012}).
+
+In our case, while the model may still turn out to have general structural problems, we suggest that part of the issue is the nature of the forcing used, and the timescales on which the system responds to it.
+We hypothesize that the system may have two timescales in it - a fast timescale in which eqs.
+\ref{eq:s} - \ref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in eqs.
+\ref{eq:p}, \ref{eq:g}, in which measured fluorescence changes in response to the forcing.
+If this is the case, then it may be that our experimental data can only probe the system via the fixed point of the models core, and that the parameters contained within those equations can only work to set the particular fixed point values of $p$ and $g$ that the system tends toward.
+If this were the case, the particular shape of the models response curve would be primarily determined by the parameters appearing in eqs.
+\ref{eq:p}, \ref{eq:g}, \ref{eq:z} - namely $\mu$, $\beta$, $f_{s}$ and $\Theta$.
+If all the information contained in the parameters in eqs.\ref{eq:s} - \ref{eq:c} can only be expressed in the form of the models fixed point, it would not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
+
+Fig.
+\ref{model_output} shows model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
+It demonstrates that, at least in some of the parameter sets initially estimated, this difference in time scales exists.
+The fixed points of eqs.
+\ref{eq:s} - \ref{eq:c} affects eqs.
+\ref{eq:p}, \ref{eq:g} through the $\beta m +f_{s}\beta c$ term - we would expect that, if the fixed point value of this term is what is relevant in determining model response, there would be some consistency in its value across the parameter sets found.
+Fig.
+\ref{fixedpoint_f} shows a scatterplot of model fixed point values against error function value, and demonstrates that, though individual parameter values can be spread across very large ranges, they are correlated in such a way as to give similar model fixed points.
+In Fig.
+\ref{fixedpoint_mu}, we also see a strong positive correlation between the fixed point values and the values of $\mu$ ($R = 0.9854$ for the tightly clustered data).
+Taken together with the relatively tightly constrained values of $\mu$ found in Fig.
+\ref{InitialResults}, these results suggest that in order to minimize error, the model is effectively varying
+
+\begin{figure}[h]	
+	\begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.26]{14_7_pred_13_9_data}
+        \caption{\textit{14\_7} prediction \textit{13\_9} data}
+        \label{SensitivityMatrix_unscaled} 
+    \end{subfigure}
+    \begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.26]{13_9_pred_14_7_data}
+        \caption{\textit{13\_9} prediction \textit{14\_7} data}
+                \label{SensitivityMatrix_scaled} 
+    \end{subfigure}
+    \caption{Cross validating data by taking parameter estimates from one dataset, and using them to predict another.
+Here we take the best parameter set fitted on the \textit{13\_9}, and fit it against the \textit{14\_7} dataset, and vice versa }
+\label{CrossValidation}   
+\end{figure}
 
@@
 
- \begin{itemize}
- \item discuss estimatbility criteria- uniqeuess and sensitivity
- \item put in some sensitivity analysis results
- \item discuss model fixed points
- \item discuss systematic paramater shift.
- \item discuss cross - validation and predicitive ability of parameter sets found.
- \item suggest two times, work out fixed point for all the solutions found, look at structure of soln,
- \item re run from starting pararms.
- \end{itemize}
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.3]{model_output}
+        \caption{}
+        \label{model_output} 
+\end{figure}
+
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.25]{fixedpoint_f}
+        \caption{}
+        \label{fixedpoint_f} 
+\end{figure}
+
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.25]{fixedpoint_mu}
+        \caption{}
+        \label{fixedpoint_mu} 
+\end{figure}
 
@@
 \clearpage
+
+ \begin{itemize}
+ \item discuss estimatbility criteria- uniqeuess and sensitivity
+ \item put in some sensitivity analysis results
+ \item discuss model fixed points
+ \item discuss systematic paramater shift.
+ \item discuss cross - validation and predicitive ability of parameter sets found.
+ \item suggest two times, work out fixed point for all the solutions found, look at structure of soln,
+ \item re run from starting pararms.
+ \end{itemize}
 

### 2015-09-03 MScWriteup.tex
more writing

@@
 
-% NEW SUBSECTION
-
-\subsection{Initial Parameter Estimates}
-\label{Initial Parameter Estimates}
-
- We begin by fitting two of the datasets, \textit{13\_9} and \textit{14\_7}, by choosing 200 points, uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
-Results are shown in Figs.
-\ref{InitialResults}, \ref{hist_f}.
-Fig.
-\ref{InitialResults} demonstrates that the model is capable of quantitatively capturing the data - however, the fitting also suggests that some parameters are not tightly constrained, taking values right across the initial bounding range specified.
-The correlation matrices in Fig.
-\ref{InitialResults} also indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
-Referring to eq.
-\ref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
-The results suggest a fitness landscape relatively flat to perturbations in certain combinations of parameters, and indicate we may have difficulty obtaining unique estimates of the model parameters.
-
- We can test the predictive ability of our model by cross validating - taking the parameter values found in the fitting of one dataset, and using them to give model predictions for another.
-We thus take the parameter values giving the best fit for the  \textit{13\_9} dataset and use them to predict the \textit{14\_7} dataset, and vice versa.
-Results are shown in Fig.\ref{CrossValidation}.Though the predictions are reasonable, they are substantially worse than the predictions for the data they were trained on.
-The first reason for this may be experiment to experiment variability of parameter values \cite{Hu2015}.
-The second may be that the model is overfitting.
-
- Taken together, these results suggest that some parameters may be inestimable, and that the data may be better described by a simplified model.
-
- \subsection{Parameter Estimability}
-\label{Parameter Estimability}
-
- There are two main reasons why a parameter may not be estimable  \cite{Mclean2012,Yao2003, Beck}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others.
-
- We may investigate whether our model suffers from such problems by performing a local sensitivity analysis about one of the solutions found in our initial parameter estimation.
-We numerically estimate the sensitivity matrix, $S$:
-
- \begin{align}
-S_{ij} = \hat{\theta}_{j} \frac{\partial z}{ \partial \theta_{j}}\Bigr|_{t_{i}}
-\end{align}
-
-where $S_{ij}$ is the derivative of the observed fluorescence, $z$, with parameter $\theta_{j}$, evaluated at timepoint $t_{i}$.
-This matrix contains information about how sensitive $z$ is to perturbations in parameter values, and at what times it is most sensitive.
-$\hat{\theta}_{j}$ is the value of the parameter that the derivative is being evaluated at.
-It is included to set the scale that parameters may vary at, to ensure that apparently small sensitivity values do not result from a poor choice of units.
-
-It can be shown that if sensitivity co-efficients are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
-Related to this fact, a number of measures have been proposed to asses parameter estimability from the sensitivity matrix \cite{Mclean2012}
-
-Fig.
-\ref{SensitivityMatrix} shows plots of each column of $S$, evaluated at the parameter set giving the lowest error in the \textit{13\_9} dataset - these show the evolution of the sensitivity matrix over time.
-Fig.
-\ref{SensitivityMatrix_unscaled} shows them unscaled, Fig.
-\ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$, so that their shapes may be more easily compared.
-We see that many of the parameters give sensitivity curves of similar shapes - the effects of perturbing any one of these parameters all look similar in terms of model output.
-
- It is unsurprising then that some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved, and as such that we should view estimates of these parameters with extreme caution.
-Note that the sensitivity curve that looks least similar to the others in Fig.
-\ref{SensitivityMatrix_scaled} - $\mu$ - corresponds to a relatively tightly estimated value in Fig.
-\ref{InitialResults}.
- \todo[inline]{make more quantitivative with colinaerity idnex}
-
-Our inability to estimate some model parameters may stem from structural problems with the model (often termed \textit{structural identifiability}), in which two different parameter sets can give identical model predictions (eq.
-\ref{nonidentifiable}) .
-General methods exist to analyse the structural identifiability of an ODE model \cite{Jimenez-Hornero2008,Grewal1976 } \footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product can ever be estimated, a problem no amount of data can fix.}.
-
- \begin{align}
-z(\boldsymbol{\theta}_{1},t) = z(\boldsymbol{\theta}_{2},t) \centernot\implies \theta_{1} = \theta_{2}
-\label{nonidentifiable}
-\end{align}
-
-Problems may also come from limitations in available data - differing experimental conditions may lift the strong correlations between parameter effects on model output (this is often termed \textit{practical identifiability} \cite{Mclean2012}).
-
-In our case, while the model may still turn out to have general structural problems, we suggest that part of the issue is the nature of the forcing used, and the timescales on which the system responds to it.
-We hypothesize that the system may have two timescales in it - a fast timescale in which eqs.
-\ref{eq:s} - \ref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in eqs.
-\ref{eq:p}, \ref{eq:g}, in which measured fluorescence changes in response to the forcing.
-If this is the case, then it may be that our experimental data can only probe the system via the fixed point of the models core, and that the parameters contained within those equations can only work to set the particular fixed point values of $p$ and $g$ that the system tends toward.
-If this were the case, the particular shape of the models response curve would be primarily determined by the parameters appearing in eqs.
-\ref{eq:p}, \ref{eq:g}, \ref{eq:z} - namely $\mu$, $\beta$, $f_{s}$ and $\Theta$.
-If all the information contained in the parameters in eqs.\ref{eq:s} - \ref{eq:c} can only be expressed in the form of the models fixed point, it would not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
-
-Fig.
-\ref{model_output} shows model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
-It demonstrates that, at least in some of the parameter sets initially estimated, this difference in time scales exists.
-The fixed points of eqs.
-\ref{eq:s} - \ref{eq:c} affects eqs.
-\ref{eq:p}, \ref{eq:g} through the $\beta m +f_{s}\beta c$ term - we would expect that, if the fixed point value of this term is what is relevant in determining model response, there would be some consistency in its value across the parameter sets found.
-Fig.
-\ref{fixedpoint_f} shows a scatterplot of model fixed point values against error function value, and demonstrates that, though individual parameter values can be spread across very large ranges, they are correlated in such a way as to give similar model fixed points.
-In Fig.
-\ref{fixedpoint_mu}, we also see a strong positive correlation between the fixed point values and the values of $\mu$ ($R = 0.9854$ for the tightly clustered data).
-Taken together with the relatively tightly constrained values of $\mu$ found in Fig.
-\ref{InitialResults}, these results suggest that in order to minimize error, the model is effectively varying
-
-\begin{figure}[h]	
-	\begin{subfigure}[h]{0.49\textwidth}
-    \centering
-        \includegraphics[scale = 0.26]{14_7_pred_13_9_data}
-        \caption{\textit{14\_7} prediction \textit{13\_9} data}
-        \label{SensitivityMatrix_unscaled} 
-    \end{subfigure}
-    \begin{subfigure}[h]{0.49\textwidth}
-    \centering
-        \includegraphics[scale = 0.26]{13_9_pred_14_7_data}
-        \caption{\textit{13\_9} prediction \textit{14\_7} data}
-                \label{SensitivityMatrix_scaled} 
-    \end{subfigure}
-    \caption{Cross validating data by taking parameter estimates from one dataset, and using them to predict another.
-Here we take the best parameter set fitted on the \textit{13\_9}, and fit it against the \textit{14\_7} dataset, and vice versa }
-\label{CrossValidation}   
-\end{figure}
-
-\begin{figure}[h]	
-	\begin{subfigure}[h]{0.49\textwidth}
-    \centering
-        \includegraphics[scale = 0.3]{Sensitivity_unscaled}
-        \caption{Unscaled}
-        \label{SensitivityMatrix_unscaled} 
-    \end{subfigure}
-    \begin{subfigure}[h]{0.49\textwidth}
-    \centering
-        \includegraphics[scale = 0.3]{Sensitivity_scaled}
-        \caption{Scaled}
-                \label{SensitivityMatrix_scaled} 
-    \end{subfigure}
-    \caption{Sensitivity co-efficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset.
-\ref{SensitivityMatrix_unscaled} shows them unscaled, \ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$.}
-\label{SensitivityMatrix}   
-\end{figure}
-
-\begin{figure}[h]	
-    \centering
-        \includegraphics[scale = 0.3]{model_output}
-        \caption{}
-        \label{model_output} 
-\end{figure}
-
-\begin{figure}[h]	
-    \centering
-        \includegraphics[scale = 0.25]{fixedpoint_f}
-        \caption{}
-        \label{fixedpoint_f} 
-\end{figure}
-
-\begin{figure}[h]	
-    \centering
-        \includegraphics[scale = 0.25]{fixedpoint_mu}
-        \caption{}
-        \label{fixedpoint_mu} 
-\end{figure}
-
 \begin{figure*}
@@
 
-\clearpage
+% NEW SUBSECTION
+
+\subsection{Initial Parameter Estimates}
+\label{Initial Parameter Estimates}
+
+ We begin by fitting two of the datasets, \textit{13\_9} and \textit{14\_7}, by choosing 200 points, uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
+Results are shown in Figs.
+\ref{InitialResults}, \ref{hist_f}.
+Fig.
+\ref{InitialResults} demonstrates that the model is capable of quantitatively capturing the data - however, the fitting also suggests that some parameters are not tightly constrained, taking values right across the initial bounding range specified.
+The correlation matrices in Fig.
+\ref{InitialResults} also indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
+Referring to eq.
+\ref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
+The results suggest a fitness landscape relatively flat to perturbations in certain combinations of parameters, and indicate we may have difficulty obtaining unique estimates of the model parameters.
+
+ We can test the predictive ability of our model by cross validating - taking the parameter values found in the fitting of one dataset, and using them to give model predictions for another.
+We thus take the parameter values giving the best fit for the  \textit{13\_9} dataset and use them to predict the \textit{14\_7} dataset, and vice versa.
+Results are shown in Fig.\ref{CrossValidation}.
+
+ \begin{figure}[h!]	
+	\begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.26]{14_7_pred_13_9_data}
+        \caption{\textit{14\_7} prediction \textit{13\_9} data}
+        \label{SensitivityMatrix_unscaled} 
+    \end{subfigure}
+    \begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.26]{13_9_pred_14_7_data}
+        \caption{\textit{13\_9} prediction \textit{14\_7} data}
+                \label{SensitivityMatrix_scaled} 
+    \end{subfigure}
+    \caption{Cross validating data by taking parameter estimates from one dataset, and using them to predict another.
+Here we take the best parameter set fitted on the \textit{13\_9}, and fit it against the \textit{14\_7} dataset, and vice versa }
+\label{CrossValidation}   
+\end{figure}
+
+Though the predictions are reasonable, they are substantially worse than the predictions for the data they were trained on.
+The first reason for this may be experiment to experiment variability of parameter values \cite{Hu2015}.
+The second may be that the model is overfitting.
+
+ Taken together, these results suggest that some parameters may be inestimable, and that the data may be better described by a simplified model.
+
+ \subsection{Parameter Estimability}
+\label{Parameter Estimability}
+
+ There are two main reasons why a parameter may not be estimable \cite{Mclean2012,Yao2003, Beck}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others.
+
+These problems may stem from structural inadequacies in the model (often termed \textit{structural identifiability}), in which two different parameter sets can give identical model predictions \cite{Jimenez-Hornero2008,Grewal1976 }.
+If this is the case, no amount of experimental data will allow us to estimate parameters, and we must consider reformulating the model.
+\footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product can ever be estimated, a problem no amount of data can fix.}
+
+Problems may also arise for more practical reasons (\textit{practical identifiability} \cite{Mclean2012}).
+For example, it is possible that in the experimental regime we operate in, parameter effects may be weak, or highly correlated, but in general this is not true.
+\footnote{\cite{Beck} gives an example in which a parameter only effects model predictions after several hours, though others will affect it at all timescales - in this example, if we only took data for a few minutes, the parameter would be inestimable, but in principle it is not.} In this case, parameter estimates be may improved by taking data in more varied experimental conditions, and attempting to observe as many components of the model output as possible.
+
+ We may begin to investigate these issues in our model by performing a local sensitivity analysis about one of the solutions found in our initial parameter estimation.
+We numerically estimate the sensitivity matrix, $S$:
+
+ \begin{align}
+S_{ij} = \hat{\theta}_{j} \frac{\partial z}{ \partial \theta_{j}}\Bigr|_{t_{i}}
+\end{align}
+
+where $S_{ij}$ is the derivative of the observed fluorescence, $z$, with parameter $\theta_{j}$, evaluated at timepoint $t_{i}$.
+This matrix contains information about how sensitive $z$ is to perturbations in parameter values, and at what times it is most sensitive.
+$\hat{\theta}_{j}$ is the value of the parameter that the derivative is being evaluated at.
+It is included to set the scale that parameters may vary at, to ensure that apparently small sensitivity values do not result from a poor choice of units.
+
+It can be shown that if sensitivity co-efficients are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
+Related to this fact, a number of measures have been proposed to asses parameter estimability from the sensitivity matrix \cite{Mclean2012}, the simplest of which is to simply plot the sensitivity co-effecients as a function of time, and visually check for obvious linear relations between the curves.
+
+Fig.
+\ref{SensitivityMatrix} shows plots of each column of $S$, evaluated at the parameter set giving the lowest error in the \textit{13\_9} dataset - these show the evolution of the sensitivity matrix over time.
+
+\begin{figure}[h]	
+	\begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.3]{Sensitivity_unscaled}
+        \caption{Unscaled}
+        \label{SensitivityMatrix_unscaled} 
+    \end{subfigure}
+    \begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.3]{Sensitivity_scaled}
+        \caption{Scaled}
+                \label{SensitivityMatrix_scaled} 
+    \end{subfigure}
+    \caption{Sensitivity co-efficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset.
+\ref{SensitivityMatrix_unscaled} shows them unscaled, \ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$.}
+\label{SensitivityMatrix}   
+\end{figure}
+
+ Fig.
+\ref{SensitivityMatrix_unscaled} shows them unscaled, Fig.
+\ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$, so that their shapes may be more easily compared.
+We see that many of the parameters give sensitivity curves of similar shapes - the effects of perturbing any one of these parameters all look similar in terms of model output.
+
+This may help to explain why some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
+As such, we should view estimates of these parameters with extreme caution.
+Note that the sensitivity curve that looks least similar to the others in Fig.
+\ref{SensitivityMatrix_scaled} - $\mu$ - corresponds to a relatively tightly estimated value in Fig.
+\ref{InitialResults}.
+
+We suggest that part of the issue is the step function forcing used in model predictions, and the timescales on which the system responds to it.
+We hypothesize that the system may have two timescales in it - a fast timescale in which eqs.
+\ref{eq:s} - \ref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in eqs.
+\ref{eq:p}, \ref{eq:g}, in which measured fluorescence changes in response to the forcing.
+If this is the case, then it may be that our experimental data can only probe the system via the fixed point of eqs.
+\ref{eq:s} - \ref{eq:c} , and that the parameters contained within those equations can only work to set the particular fixed point values of $p$ and $g$ that the system tends toward.
+If this were the case, the particular shape of the models response curve would be primarily determined by the parameters appearing in eqs.
+\ref{eq:p}, \ref{eq:g}, \ref{eq:z} - namely $\mu$, $\beta$, $f_{s}$ and $\Theta$.
+If all the information contained in the parameters in eqs.\ref{eq:s} - \ref{eq:c} can only be expressed in the form of the models fixed point, it would not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
+
+Fig. \ref{model_output} shows model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
+
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.3]{model_output}
+        \caption{}
+        \label{model_output} 
+\end{figure}
+
+It demonstrates that, at least in some of the parameter sets initially estimated, this difference in time scales exists.
+The fixed points of eqs.
+\ref{eq:s} - \ref{eq:c} affects eqs.
+\ref{eq:p}, \ref{eq:g} through the $\beta m +f_{s}\beta c$ term - we would expect that, if the fixed point value of this term is what is relevant in determining model response, there would be some consistency in its value across the parameter sets found.
+Fig.
+\ref{fixedpoint_f} shows a scatterplot of model fixed point values against error function value, and demonstrates that, though individual parameter values can be spread across very large ranges, they are correlated in such a way as to give similar model fixed points.
+
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.25]{fixedpoint_f}
+        \caption{}
+        \label{fixedpoint_f} 
+\end{figure}
+
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.25]{fixedpoint_mu}
+        \caption{}
+        \label{fixedpoint_mu} 
+\end{figure}
+
+In Fig.
+\ref{fixedpoint_mu}, we also see a strong positive correlation between the fixed point values and the values of $\mu$ ($R = 0.9854$ for the visually tightly clustered data).
+Taken together with the relatively tightly constrained values of $\mu$ found in Fig.
+\ref{InitialResults}, these results suggest that in order to minimize error, the algorithm is effectively trying to vary $\beta m +f_{s}\beta c$ and $\mu$, and finding a `trench' of highly correlated values, which is shown in Fig.
+\ref{fixedpoint_mu}.
+
+These results further suggest that, while we are trying to minimise model error over our initial high dimensional space, we are effectively working in a lower dimensional space, in which one dimension is the value of the models $\beta m +f_{s}\beta c$ .
 
  \begin{itemize}
+ \item explore degeneracies in fixed point equations
  \item discuss estimatbility criteria- uniqeuess and sensitivity
@@
 \label{Conclusions and Further work}
+
+ \begin{itemize}
+ \item discuss evidence for model being very over parametrized, and suggest a simpler model may work equally well - suggest biologically interesting parameters may be unobservable.
+ \item discuss general methodological flaws - throwing marbles in the air, local minima, is MLE even any good?
+say data shows even the EA doesn't avoid local minima.
+The space is so unconstrained.
+the algorithm can't be expected to just fix all these problems.
+ \item suggest additional forcing time data may not help due to two timing nature, 
+ \item suggest there MAY be other ways to fit the data, but the wide range of parameter bounds mean we just don't know - the landscape is complicated and huge, and there may be minima widely separated in the parameter space.
+whose to say which is right?
+ \item Further work - on experimnetal side, propoer lit review to rightly constrain some values, fix more.
+More state variables!
+should easily be able to fix mu, theta, maybe others too?
+other forcing functions?
+ \item a proper bloody methodology - gaussian processes.
+ \item mention this is how sys bio papers do it, and we were given some earlier fitting work which we think is now crap.
+ \item suggest a general structural analysis be carried out.
+
+ \end{itemize}
 

### 2015-09-04 MScWriteup.tex
more conclusion writing

@@
 
-\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e, epstopdf, caption, subcaption,centernot }
+\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e, epstopdf, caption, subcaption,centernot}
 
@@
 This regulation allows the cell to control protein expression, and so cell behaviour, in response to various environmental cues.
-The natural cell machinery which performs it takes the form a genetic circuits - networks of interacting gene expression regulators.
+The natural cell machinery which performs it takes the form of genetic circuits - networks of interacting gene expression regulators.
 This genetic circuitry offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.
@@
 This report will focus on one such sRNA system, introduced in \cite{Rodrigo2012}.
-It will extend existing understanding of it beyond the qualitative by proposing a quantitative model of gene expression, in the form of a set of ODE's, and fitting this model to available time series data to estimate its unknown parameters.
+It will extend existing understanding of it beyond the qualitative by first proposing a quantitative model of its behaviour in the form of a set of ODE's, and then fitting this ODE model to available time series data to estimate its unknown parameters.
 
 The report is structured as follows.
-In the remainder of this section we review the sRNA regulatory system we will consider, and discuss recent single cell fluorescence experiments performed on this system.
+In the remainder of this section we review the regulatory system we will consider, and discuss recent single cell fluorescence experiments performed on it.
 In section \ref{Model Derivation} we propose a set of ODE's to model the system.
-In sect \ref{Parameter Estimation} and estimate its unknown parameters by fitting to time series data.
+In section \ref{Parameter Estimation} we attempt to estimate its unknown parameters by fitting the model to time series data.
 Finally, in section \ref{Conclusions and Further work}, we conclude, and suggest directions for further work.
@@
 \label{The sRNA regulatory system}
-In bacteria, one mechanism by which gene expression is regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA (Fig.
+In bacteria, one mechanism by which gene expression is naturally regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA (Fig.
 \ref{RBS}).
 This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA.
-In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the UTR), upstream of the RBS.
-Translation may be self repressed by this `tail' of the mRNA folding over and binding across the RBS, forming a stem loop in the mRNA and preventing the Ribosome from binding (Fig.
+
+In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the UTR), upstream of the RBS \footnote{Both DNA and RNA are directional - the backbone of the molecule is not symmetric.
+This gives the molecule two distinct ends, denoted 5' and 3'.
+Translation can only occur in the 5' to 3' direction, hence the meaning of `upstream'}.
+Translation may be self repressed by this `tail' folding over and binding across the RBS, forming a  loop in the mRNA and preventing the Ribosome from binding (Fig.
 \ref{RBS}).
-This self repression may be released with an sRNA which binds to the same region on the mRNA - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind.
+This self repression may be released with an sRNA which binds to this looped tail - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind.
 In summary, the presence of the sRNA positively regulates gene expresssion.
@@
 \includegraphics[trim = 100 170 100 400,page=10,clip = true,scale = 0.6]{Appendix.pdf}
-\caption{A mechanism by which sRNA's can regulate gene expression.
+\caption{A mechanism by which sRNA can regulate gene expression.
 Initially, the 5' UTR of the mRNA is folded over the RBS, forming a loop and blocking Ribosome binding.
-The sRNA binds to this mRNA, causing a conformational change which uncovers the RBS, and allows translation to occur.Image reproduced from \cite{Rodrigo2012} }
+The sRNA binds to this loop in the mRNA, causing a conformational change which uncovers the RBS, and allows translation to occur.
+Image reproduced from \cite{Rodrigo2012}.}
 \label{RBS}
@@
 \ref{reactionscheme}.
-The two RNA's, originally in their own individually folded states, would initially interact via a small 'toehold' sequence of unpaired nucleotides to form an unstable transition state.
+The sRNA and mRNA, originally in their own individually folded states, would initially interact via a small
+`toehold' sequence of unpaired nucleotides to form an unstable transition state.
 This intermediate complex would then rapidly form a final, stable complex with the desired conformation.
-By suggesting sRNA and mRNA sequences which optimised this energy landscape, \cite{Rodrigo2012} suggested several devices which would form a stable hybrid with the RBS free, and experimentally validated their function in \textit{E.
-coli}, using an mRNA which codes for GFP for experimental ease (note the algorithm only optimises the 5' UTR of the mRNA, so the actual protein being coded for is unimportant).
+By finding sRNA and mRNA sequences which optimised the energy landscape shown in Fig.
+\ref{reactionscheme}, \cite{Rodrigo2012} suggested several sRNA-mRNA pairs which would work in tandem to form a stable hybrid with the RBS free.
 
-\begin{figure}[H]
+\begin{figure}[h]
 \centering
 \includegraphics[trim = 60 630 300 30,page=2,clip = true]{pnas1203831109.pdf}
-\caption{A network with an intuitively clear community structure, which is captured by the partition chosen, shown in gray. Image reproduced from \cite{Rodrigo2012}}
+\caption{A proposed reaction scheme between the synthetic sRNA-mRNA pairs proposed in \cite{Rodrigo2012}.
+The reaction co-ordinate is defined as the number of paired nucleotides, and the vertical axis denotes free energy.
+The RNA's initially interact via a small
+`toehold' sequence, forming an unstable transition state, which then stabilises to give the final compound. Image reproduced from \cite{Rodrigo2012}}
 \label{reactionscheme}
@@
 
-Further, by placing the concentrations of the sRNA and mRNA under the control of tuneable promoters, \cite{Rodrigo2012} constructs a logical AND gate from one of the proposed devices (RAJ11) \textit{in vivo} (Fig.
+The authors then experimentally validated their methodology by testing the suggested sRNA-mRNA pairs in \textit{E.
+coli}.
+Further, by placing the \textit{in vivo}  concentrations of the sRNA and mRNA under the control of tuneable promoters, they constructed a logical AND gate from one of the proposed pairs (Fig.
 \ref{ANDGate}).
+
+\todo[inline]{promoter? ).}
+
 In this system, transcription of the designed sRNA and mRNA are placed under the control of promoter regions, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}.
@@
 coli} considered.
-These repressors disable the promoter regions, and so by default transcription of the RNA's is turned off, and no protein is produced.
-These repressors can themselves be disabled by the presence of two chemicals, aTC and IPTG, which can be introduced externally into the cell (Fig.
+These repressors disable the promoter regions, and so by default transcription of both RNA's is turned off, and no protein is produced.
+These repressors can themselves be disabled by the presence of two chemicals, aT and IPTG, which can be introduced externally into the cell (Fig.
 \ref{ANDGate}).
@@
 But when both are present, the conformational change discussed above occurs, and protein is produced.
-\todo[inline]{promoter?}
 
@@
 Such an understanding would allow, for example, tailoring of the system in reponse to design requirements, by altering the values of the important parameters of the model.
-By changing which sRNA-mRNA device is used in the system, it would also allow exploration of the relationship between the thermodynamic properties of each device, and the model's rate constants.
+By changing which sRNA-mRNA pair is used in the system, it would also allow exploration of the relationship between the thermodynamic properties of each device, and the model's rate constants.
 
@@
 
-Recent experiments have used timelapse microscopy to observe the fluoresence of bacteria which contain the above device, as they are periodically forced with a varying aTc or IPTG concentration \cite{Jaramillo}.
+mRNA concentrations in this system shown in Fig.
+\ref{ANDGate} can be indirectly observed by designing the mRNA to code for GFP \footnote{ The methodology of \cite{Rodrigo2012} only optimises the 5' UTR of the mRNA, so the actual protein being coded for is unimportant}.
+Recent experiments have used timelapse microscopy to observe the fluoresence of bacteria which contain the above sRNA-mRNA pair, as they are periodically forced with a varying aTc or IPTG concentration \cite{Jaramillo}.
 
@@
 Image reproduced from \cite{Jaramillo}.}
-\label{reactionscheme}
+\label{}
 \end{figure}
 
-The data we will consider consists of three sets of individual cell timeseries, labelled \textit{13\_9}, \textit{14\_7} and \textit{14\_9}.
-They correspond to three different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response, and aTc concentrations were varied periodically.
+The data we will consider consists of two sets of individual cell timeseries, labelled \textit{13\_9}, \textit{14\_7}.
+They correspond to different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response, and aTc concentrations were varied periodically.
 Section \ref{Initial  Experimental Data} shows the full datasets, with their forcing functions.
@@
   Based on the reaction mechanism in Fig.
-\ref{reactionscheme}, the hybridization of the sRNA and mRNA first into an unstable complex, then a stable one, is modelled in eqs.\ref{eq:s} - \ref{eq:c}.
+\ref{reactionscheme}, the hybridization of the sRNA and mRNA first into an unstable complex, then a stable one, is modelled in eqs.
+\ref{eq:s}-\ref{eq:c}.
 The initial binding is modelled as a reversible reaction with forward and backward rates $k_{on}$ and $k_{off}$.
@@
 
- Control of the system by aTc and IPTG is modelled by $y(t)$ and $x(t)$ in eqs.\ref{eq:s}-\ref{eq:m}.
+ Control of the system by aTc and IPTG is modelled by $y(t)$ and $x(t)$ in eqs.
+\ref{eq:s}-\ref{eq:m}.
 $y(t)$ models the response of the sRNA transcription rate to a time varying aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}.
@@
  $N\frac{1}{f_{T}}$ and $N\frac{\alpha_{T}}{f}$, where $\alpha_{T}$ is the maximal transcription rate of the $P_{\mathrm{LtetO-1}}$ promoter.
-When engineering the system, many copies of the $P_{\mathrm{LtetO-1}}$ promoter and section of DNA coding for sRNA may be placed in the plasmid DNA - this is modelled by the copy number, $N$.
+When engineering the system, many copies of the $P_{\mathrm{LtetO-1}}$ promoter, and section of DNA coding for sRNA, may be placed in the bacterial DNA - this is modelled by the copy number, $N$.
 Identical considerations hold for $x(t)$ and IPTG concentrations.
@@
 \todo[inline]{para about clpX}
-Finally, eq. \ref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence.
+Finally, eq. \ref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence, assuming a linear response.
 
@@
 \end{align}
-where the experimental timeseries, $\mathbf{y}_{\mathrm{exp}}(t_{i})$ is evaluated at timepoints $t_{i}$, $i = 1 \hdots N$.
+where the experimental timeseries, $\mathbf{y}_{\mathrm{exp}}(t_{i})$ is recorded at timepoints $t_{i}$, $i = 1 \hdots N$.
 This error function defines a landscape in $\boldsymbol{\theta}$ space, and we seek to minimise it by varying $\boldsymbol{\theta}$.
@@
 To further reduce the search space, we assume that $\delta_m$, $\delta_{sm}$ and $\delta_c$ all take similar values, and set them equal.
-After this is done, we are left with a 9 dimensional search space, bounded by a hypercube (parameters to be estimated are shown in bold in \ref{ModelParameters}).
+After this is done, we are left with a 9 dimensional search space, bounded by a hypercube (parameters to be estimated are shown in bold in table \ref{ModelParameters}).
 
@@
 \ref{SensitivityMatrix} shows plots of each column of $S$, evaluated at the parameter set giving the lowest error in the \textit{13\_9} dataset - these show the evolution of the sensitivity matrix over time.
+Fig.
+\ref{SensitivityMatrix_unscaled} shows them unscaled, Fig.
+\ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$, so that their shapes may be more easily compared.
 
@@
 
- Fig.
-\ref{SensitivityMatrix_unscaled} shows them unscaled, Fig.
-\ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$, so that their shapes may be more easily compared.
 We see that many of the parameters give sensitivity curves of similar shapes - the effects of perturbing any one of these parameters all look similar in terms of model output.
-
 This may help to explain why some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
@@
     \centering
-        \includegraphics[scale = 0.3]{model_output}
+        \includegraphics[scale = 0.25]{model_output}
         \caption{}
@@
 
- \begin{itemize}
- \item explore degeneracies in fixed point equations
- \item discuss estimatbility criteria- uniqeuess and sensitivity
- \item put in some sensitivity analysis results
- \item discuss model fixed points
- \item discuss systematic paramater shift.
- \item discuss cross - validation and predicitive ability of parameter sets found.
- \item suggest two times, work out fixed point for all the solutions found, look at structure of soln,
- \item re run from starting pararms.
- \end{itemize}
-
 \section{Conclusions and Further work}
@@
 
- \begin{itemize}
- \item discuss evidence for model being very over parametrized, and suggest a simpler model may work equally well - suggest biologically interesting parameters may be unobservable.
- \item discuss general methodological flaws - throwing marbles in the air, local minima, is MLE even any good?
-say data shows even the EA doesn't avoid local minima.
-The space is so unconstrained.
-the algorithm can't be expected to just fix all these problems.
- \item suggest additional forcing time data may not help due to two timing nature, 
- \item suggest there MAY be other ways to fit the data, but the wide range of parameter bounds mean we just don't know - the landscape is complicated and huge, and there may be minima widely separated in the parameter space.
-whose to say which is right?
- \item Further work - on experimnetal side, propoer lit review to rightly constrain some values, fix more.
-More state variables!
-should easily be able to fix mu, theta, maybe others too?
-other forcing functions?
- \item a proper bloody methodology - gaussian processes.
- \item mention this is how sys bio papers do it, and we were given some earlier fitting work which we think is now crap.
- \item suggest a general structural analysis be carried out.
+In this report, we have proposed a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}.
+We have attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach, similar to that found in recent system biology literature \cite{Hu2015}.
+Where this approach has failed to give estimates, we have discussed in detail how the method has broken down.
 
- \end{itemize}
+The results in this paper suggest that, even though the model has a solid biological rationale, some parameters included in it may not be estimable.
+We have suggested practical reasons for this, namely that the step function nature of the forcing is giving highly correlated parameter effects, causing the minimisation algorithm to effectively work in a much lower dimensional space than our full parameter space.
+This result suggests that using the additional step function data  available may not help in determining the parameters - all of these step functions are applied at timescales over which the `fast' part of the model responds instantly, and so essentially tell us the same thing about the parameters contained in those equations.
+The different sets may give more information about $\mu$, which appears to work on a timescale comparable to the forcing.
 
-\todo[inline]{some stuff}
+We note that it is possible that this separation of timescales is an artefact of the fitting, and not necessarily true of the system.
+We can only evaluate model predictions on the parameter sets the algorithm returns us, and it seems the case that the parameter sets found by the algorithm have this feature.
+Hypothetically, there may exist parameter sets where all state variables vary on a similar timescale - all we can say is that, in the search region specified, the algorithm has not found any which give a low error.
+
+This point raises another issue - the arbitrary nature of the initial bounding box.
+Without good biological constraints on most parameter values, they have been bounded loosely and not necessarily sensibly.
+Rate constants can vary by many orders of magnitude, and it may be possible that the true optimum lies entirely outside the bounding box specified.
+Equally,  when our fitness function has turned out to be very flat in some directions, or multimodal, good biologically motivated parameter bounds may be needed to ensure our minimisation algorithm does not wander outside realistic parameter space - the algorithm may present several equally good regions of parameter space, and improved bounds will help us choose between them.
+We also note that while some parameters may be hard to get any realistic bounds on, values for others - $\mu, \Theta$ for example - may already be available.
+
+ The model may also have general structural issues.
+\cite{Mclean2012, Grewal1976} give general tests which can be applied to a system of ODE's to determine if all parameters are, in principle, estimable.
+We note that our system has few enough parameters to make these tests feasible.
+
+ Thus, further work may consist of tightening the bounds on parameter values, carrying out a structural analysis, and possibly simplifying the model.
+A much simpler model, consisting of  just eqs.
+\ref{eq:p} \ref{eq:g}, may explain existing data perfectly well without the need for inestimable parameters.
+Experimentally, one option  may be to use different forcing functions.
+The effect of this on parameter estimability might be anticipated by running a sensitivity analysis using this new forcing, to see if the form of Fig.
+\ref{SensitivityMatrix} changes.
+Another option, though perhaps experimentally infeasible, would be to directly observe other components of the state vector - $s$, $m$, $s:m$ and $c$.
+This would improve parameter estimates by giving direct observations of the effects of parameters contained in eqs.
+\ref{eq:s} - \ref{eq:c}.
+
+Some earlier fitting work has been done on this dataset via a similar method to the one used above.
+Our results suggest that those estimates should be viewed with caution - the estimates will not be unique, and the values of some parameters (those in eqs.
+\ref{eq:s} - \ref{eq:c}) may be very far off any values which are eventually found.
+
+These results also suggest general problems with our methodology - using a least squares minimisation approach.
+The first problem is local minima.
+Using the CMA-ES will give better results that a local minimisation algorithm, but you can never be sure you have found a global optimum - for example, \cite{Algorithms2003} compares several algorithms and finds, in the problem set, none locate the true global optimum.
+A related problem is that of working with single points rather than distributions on parameter space.
+The CMA-ES will only ever find a single optimum, but in the case where the error landscape is very flat -  for example a flat basin pocketed by many local minima - the exact minimum it finds paints a misleading picture, because it gives no idea of the uncertainty in the answer given.
+To surmount this problem, we may rerun the algorithm many times, from many starting locations, and interpret the resulting spread of parameter values found as an uncertainty \cite{Hu2015}, as we have done in this report.
+However this methodology is rather ad hoc, and a more systematic approach which gives a probability distribution on parameter space would be desirable.
+
+\todo[inline]{write section on GP's}.
 
@@
 Note the forcing curve's height is schematic - aTc concentation is switched between off, and a level which saturates the cell's response}
-\label{reactionscheme}
+\label{}
 \end{figure}
@@
 Note the forcing curve's height is schematic - aTc concentation is switched between off, and a level which saturates the cell's response}
-\label{reactionscheme}
+\label{}
 \end{figure}
 
-\begin{figure}[h]
-\centering
-\includegraphics[trim = 0 330 0 270 , scale = 0.5, clip = true]{14_9}
-\caption{The \textit{14\_9} dataset, with aTc forcing shown.
+%\begin{figure}[h]
+%\centering
+%\includegraphics[trim = 0 330 0 270 , scale = 0.5, clip = true]{14_9}
+%\caption{The \textit{14\_9} dataset, with aTc forcing shown.
 IPTG concentration is constant at a level which saturates the cell's response.
-Note the forcing curve's height is schematic - aTc concentation is switched between off, and a level which saturates the cell's response}
-\label{reactionscheme}
-\end{figure}
+Note the forcing %curve's height is schematic - aTc concentation is switched between off, and a level which saturates the cell's response}
+%\label{}
+%\end{figure}
 
@@
 
+ \begin{itemize}
+ \item look at degeneracies in the fixed points.
+ \item discuss evidence for model being very over parametrized, and suggest a simpler model may work equally well - suggest biologically interesting parameters may be unobservable.
+ \item discuss general methodological flaws - throwing marbles in the air, local minima, is MLE even any good?
+say data shows even the EA doesn't avoid local minima.
+The space is so unconstrained.
+the algorithm can't be expected to just fix all these problems.
+ \item suggest additional forcing time data may not help due to two timing nature, 
+ \item suggest there MAY be other ways to fit the data, but the wide range of parameter bounds mean we just don't know - the landscape is complicated and huge, and there may be minima widely separated in the parameter space.
+whose to say which is right?
+ \item Further work - on experimnetal side, propoer lit review to rightly constrain some values, fix more.
+More state variables!
+should easily be able to fix mu, theta, maybe others too?
+other forcing functions?
+ \item a proper bloody methodology - gaussian processes.
+ \item mention this is how sys bio papers do it, and we were given some earlier fitting work which we think is now crap.
+ \item suggest a general structural analysis be carried out.
+
+ \end{itemize}
+
+  \begin{itemize}
+ \item explore degeneracies in fixed point equations
+ \item discuss estimatbility criteria- uniqeuess and sensitivity
+ \item put in some sensitivity analysis results
+ \item discuss model fixed points
+ \item discuss systematic paramater shift.
+ \item discuss cross - validation and predicitive ability of parameter sets found.
+ \item suggest two times, work out fixed point for all the solutions found, look at structure of soln,
+ \item re run from starting pararms.
+ \end{itemize}
+
 \end{document}

### 2015-09-05 MScWriteup.tex
getting closer to a first draft

@@
 \markboth{hi}%
-{Computational design of RNA-based oscillatory circuits}
+{Parameter estimation of RNA-based oscillatory circuits}
 
@@
 \begin{abstract}
-genetic circuitry, RNA's offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription.
-In comparison to proteins, it is relatively straightfor
+For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA is an attractive tool, due to the ease of predicting its behaviour from physiochemical models.
+In this report, we propose a system of ODE's to model a recently designed synthetic regulatory circuit, before attempting to estimate its unknown parameters using recent time series data.
+We provide estimates for a subset of parameters, but find many of them to be inestimable.
+We discuss why this is this case for our system, and suggest how this problem might be resolved.
 \end{abstract}
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
 \centering
@@
 \includegraphics[trim = 60 630 300 30,page=2,clip = true]{pnas1203831109.pdf}
-\caption{A proposed reaction scheme between the synthetic sRNA-mRNA pairs proposed in \cite{Rodrigo2012}.
+\caption{A proposed reaction scheme between the synthetic sRNA-mRNA pairs designed in \cite{Rodrigo2012}.
 The reaction co-ordinate is defined as the number of paired nucleotides, and the vertical axis denotes free energy.
@@
 coli}.
-Further, by placing the \textit{in vivo}  concentrations of the sRNA and mRNA under the control of tuneable promoters, they constructed a logical AND gate from one of the proposed pairs (Fig.
+Further, by placing the \textit{in vivo}  concentrations of the sRNA and mRNA under the control of tuneable promoter regions \footnote{A promoter region in a DNA sequence is a sequence of nucleotides, found upstream of where transcription of a gene begins, which can influence the transcription rate of the gene.}, they constructed a logical AND gate from one of the proposed pairs (Fig.
 \ref{ANDGate}).
-
-\todo[inline]{promoter? ).}
 
@@
 These repressors disable the promoter regions, and so by default transcription of both RNA's is turned off, and no protein is produced.
-These repressors can themselves be disabled by the presence of two chemicals, aT and IPTG, which can be introduced externally into the cell (Fig.
+These repressors can themselves be disabled by the presence of two chemicals, aTc and IPTG, which can be introduced externally into the cell (Fig.
 \ref{ANDGate}).
@@
 \caption{A schematic of an experimental setup which allows single cell fluorescences to be recorded over time.
-Bacteria are grown in chambers, with a constant supply of medium flowing through them from the media inlet.
-This inlet also allows the introduction of aTc IPTG.
+Bacteria are grown in chambers, with a constant supply of medium flowing through them, which allows the introduction of aTc and IPTG.
 Software then traces these cells, allowing fluorescence timeseries to be built.
@@
 Firstly a generic degradation rate $\delta_{g}$, assumed identical for the mature and immature species, with the dilution rate $\mu$ shared by all species.
-\todo[inline]{para about clpX}
+\todo[inline]{need a description about the terms representing clpX degradation.}
 Finally, eq. \ref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence, assuming a linear response.
@@
     \centering
-        \includegraphics[scale = 0.25]{model_output}
-        \caption{}
+        \includegraphics[scale = 0.22]{model_output}
+        \caption{model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
+Note $g$ is not explicitly shown, but is simply a rescaling of $z$ and as such would lie over it.
+}
         \label{model_output} 
@@
         \includegraphics[scale = 0.25]{fixedpoint_f}
-        \caption{}
+        \caption{A scatterplot of the fixed point value of $\beta m +f_{s}\beta c$, for a given parameter set, against the error value of that set. }
         \label{fixedpoint_f} 
@@
         \includegraphics[scale = 0.25]{fixedpoint_mu}
-        \caption{}
+        \caption{A scatterplot of the fixed point value of $\beta m +f_{s}\beta c$, for a given parameter set, against the value of $\mu$ in that set.}
         \label{fixedpoint_mu} 
@@
 other forcing functions?
- \item a proper bloody methodology - gaussian processes.
- \item mention this is how sys bio papers do it, and we were given some earlier fitting work which we think is now crap.
+ \item a proper methodology - gaussian processes.
+ \item mention this is how sys bio papers do it, and we were given some earlier fitting work which we think is now wrong.
  \item suggest a general structural analysis be carried out.

### 2015-09-07 MScWriteup.tex
making illiana's changes, beginning to make manish's

@@
 In the notation of the diagram, a barred line indicates repression, and an arrowed line indicates production.
-We see a 'double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA).
-Thus presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.
+We see a `double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA).
+Thus the presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.
 Image reproduced from \cite{Rodrigo2012}.}
@@
 In this system, transcription of the designed sRNA and mRNA are placed under the control of promoter regions, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}.
-These are in turn controlled by two transcriptional repressors, TetR and LacI, which are naturally present in the strain of \text{E.
+These are in turn controlled by two transcriptional repressors, TetR and LacI, which are naturally present in the strain of \textit{E.
 coli} considered.
@@
 mRNA concentrations in this system shown in Fig.
-\ref{ANDGate} can be indirectly observed by designing the mRNA to code for GFP \footnote{ The methodology of \cite{Rodrigo2012} only optimises the 5' UTR of the mRNA, so the actual protein being coded for is unimportant}.
+\ref{ANDGate} can be indirectly observed by designing the mRNA to code for GFP \footnote{ The methodology of \cite{Rodrigo2012} only optimises the 5' UTR of the mRNA, so the actual protein being coded for is unimportant.}.
 Recent experiments have used timelapse microscopy to observe the fluoresence of bacteria which contain the above sRNA-mRNA pair, as they are periodically forced with a varying aTc or IPTG concentration \cite{Jaramillo}.
 
-The bacteria are grown a single layer thick in rows of chambers.
+A single layer of the bacteria are grown in rows of chambers.
 A medium constantly flows through these chambers, allowing normal feeding of the bacteria, and the introduction of aTc or IPTG.
-The chambers are monitored with software which traces the position of each cell over time, allowing timeseries of individual cell fluorescences to be recorded.
+The chambers are monitored with software which traces the position of each cell over time, allowing time series of individual cell fluorescences to be recorded.
 
@@
 \centering
-\includegraphics[trim = 0 0 0  85 ,page=17 ,clip = true, scale=0.3]{WCPM_Warwick-talk45-220115.pdf}
+\includegraphics[trim = 280 0 0  85 ,page=17 ,clip = true, scale=0.4]{WCPM_Warwick-talk45-220115.pdf}
 \centering
-\includegraphics[trim = 0 0 0  90 ,page=18 ,clip = true, scale=0.3]{WCPM_Warwick-talk45-220115.pdf}
+\includegraphics[trim = 0 0 0  90 ,page=18 ,clip = true, scale=0.35]{WCPM_Warwick-talk45-220115.pdf}
 \caption{A schematic of an experimental setup which allows single cell fluorescences to be recorded over time.
 Bacteria are grown in chambers, with a constant supply of medium flowing through them, which allows the introduction of aTc and IPTG.
-Software then traces these cells, allowing fluorescence timeseries to be built.
+Software then traces these cells, allowing fluorescence timesseries to be built.
 Image reproduced from \cite{Jaramillo}.}
@@
 
-The data we will consider consists of two sets of individual cell timeseries, labelled \textit{13\_9}, \textit{14\_7}.
+The data we will consider consists of two sets of individual cell time series, labelled \textit{13\_9} and \textit{14\_7}.
 They correspond to different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response, and aTc concentrations were varied periodically.
-Section \ref{Initial  Experimental Data} shows the full datasets, with their forcing functions.
+Appendix \ref{Initial  Experimental Data} shows the full datasets, with their forcing functions.
 
@@
 
-We propose a simple model, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}, to describe the system.
-Tables \ref{ModelParameters}, \ref{StateVariables} give complete descriptions of the parameters and state variables.
+In eqs.
+\ref{eq:s} - \ref{eq:z}, we propose a simple model consisting of a set of ODE's with mass action kinetics to describe the system \cite{UriAlon}.
+Its state is given by the vector $(s,m,s:m,c,p,g,z)$, with all other variables representing model parameters.
+Tables \ref{ModelParameters} and \ref{StateVariables} give complete descriptions of the parameters and state variables.
 
@@
 \ref{eq:s}-\ref{eq:c}.
-The initial binding is modelled as a reversible reaction with forward and backward rates $k_{on}$ and $k_{off}$.
+The initial binding is modelled as a reversible reaction with forward and backward rates $k_\mathrm{on}$ and $k_\mathrm{off}$.
 
@@
 \mathrm{
-sRNA \chemsign+ mRNA\chemrel[$k_{on}$][$k_{off}$]{<>} sRNA:mRNA\textsubscript{unstable}
+sRNA \chemsign+ mRNA\chemrel[$k_\mathrm{on}$][$k_\mathrm{off}$]{<>} sRNA:mRNA\textsubscript{unstable}
 }
 \end{align*}
-after which the stabilization is modelled as an irreversible reaction with rate $k_{hyb}$.
+after which the stabilization is modelled as an irreversible reaction with rate $k_\mathrm{hyb}$:
 
@@
 \mathrm{
-sRNA:mRNA\textsubscript{unstable} \chemrel[$k_{hyb}$]{->} sRNA:mRNA\textsubscript{stable} 
+sRNA:mRNA\textsubscript{unstable} \chemrel[$k_\mathrm{hyb}$]{->} sRNA:mRNA\textsubscript{stable} 
 }
@@
 
- In addition, these complexes are given degradation rates, $\delta_{s}$, $\delta_{m}$, $\delta_{sm}$, $\delta_{c}$, and dilution of chemical concentrations due to cell growth are modelled with a dilution rate $\mu$.
+ In addition, these complexes are given degradation rates, $\delta_{s}$, $\delta_{m}$, $\delta_{sm}$, $\delta_{c}$, and dilutions of chemical concentrations due to cell growth are modelled with a dilution rate $\mu$.
 
@@
 \hline \hline N & & Number of copies of promoter existing on plasmid DNA  \\
-\hline $z_{0}$ &  AFU & Baseline experimental fluorescence  \\
+\hline $z_{0}$ &  Arbitrary (AU) & Baseline experimental fluorescence  \\
 \hline $\alpha_{L}$ & nM/min & Maximal transcription rate of $P_\mathrm{LlacO-1}$ promoter\\
@@
 \hline $\gamma$ &  /min & GFP maturation rate  \\
-\hline $v_{z}$ & nM/min & degradation constant of clpx  \\
+\hline $v_{z}$ & nM/min & Degradation constant of clpx  \\
 \hline $K_{z}$   &   nM/min & Dissociation constant of clpx  \\
@@
 
-Our next goal is to estimate the unknown parameters of this model, given the available fluorescence time series data, by fitting predicted time series from the model to the data.
+Our next goal is to estimate the unknown parameters of this model, given the available fluorescence time series data, by fitting the predicted time series from the model to the data.
 Typically, this is done by minimising the least squares error between model prediction and the experimental data \cite{Brewer2008,Algorithms2003, Hu2015}.
@@
 \end{align}
-where the experimental timeseries, $\mathbf{y}_{\mathrm{exp}}(t_{i})$ is recorded at timepoints $t_{i}$, $i = 1 \hdots N$.
+where the experimental time series, $\mathbf{y}_{\mathrm{exp}}(t_{i})$ is recorded at timepoints $t_{i}$, $i = 1 \hdots N$.
 This error function defines a landscape in $\boldsymbol{\theta}$ space, and we seek to minimise it by varying $\boldsymbol{\theta}$.
 In our case, we do not have experimental data on the full state vector, but only one component of it - the observed fluorescence, $z(t)$.
-In addition, rather than a single experimental run, we have many, corresponding to a timeseries from each cell.
+In addition, rather than a single experimental run, we have many, corresponding to a time series from each cell.
 We incorporate this by fitting to the experimental mean of the data, and only minimising over the observed component.
@@
 To try and surmount this problem, \cite{Algorithms2003} suggests the use of a global optimisation algorithm, and in particular recommends several Evolutionary Algorithms, of which we choose one, the CMA-ES \cite{Hansen2006,Hansen2011}.
+\footnote{CMA-ES stands for covariance matrix evolutionary strategy.
+This algorithm generates a population of test points according to a multivariate gaussian distribution.
+It then ranks the points based on their fitness function scores, and  considers some number of the highest ranked points as a sample from a new multivariate gaussian.
+Using this sample, it then estimates a new mean and covariance matrix for this gaussian and the process is iterated.}
 
 In order to reduce the dimensionality of our search space, we can perform a literature search for existing values of some of our parameters, simplify our model to remove others, and place bounds on those that remain.
-Section \ref{Parameter literature review} contains a list of parameter values found in the literature, where available, and their reference, as well as initial bounds placed on parameters not found in the literature.
+Appendix \ref{Parameter literature review} contains a list of parameter values found in the literature, where available, and their reference, as well as initial bounds placed on parameters not found in the literature.
 To further reduce the search space, we assume that $\delta_m$, $\delta_{sm}$ and $\delta_c$ all take similar values, and set them equal.
@@
     \centering
-        \includegraphics[scale = 0.35, clip = true, trim = 80 0 80 0]{13_9_hist.eps}
-        \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset. }
+        \includegraphics[scale = 0.3, clip = true, trim = 80 0 74 0]{13_9_hist.eps}
+        \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset.}
     \end{subfigure}
@@
     \centering
-        \includegraphics[scale = 0.38,clip = true, trim = 60 12 80 0]{14_7_hist.eps}
+        \includegraphics[scale = 0.38,clip = true, trim = 60 12 74 0]{14_7_hist.eps}
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{14\_7} dataset. }
@@
         \centering
-    \includegraphics[scale = 0.25, clip = true, trim = 100 0 100 0]{13_9_heatmap.eps}
+    \includegraphics[scale = 0.25, clip = true, trim = 100 0 90 0]{13_9_heatmap.eps}
         \caption{Correlation matrix of estimated parameter sets, from the  \textit{13\_9} dataset. }
@@
     \centering
-        \includegraphics[scale = 0.25, clip = true, trim = 80 0 80 0]{14_7_heatmap.eps}
+        \includegraphics[scale = 0.25, clip = true, trim = 80 0 70 0]{14_7_heatmap.eps}
         \caption{Correlation matrix of estimated parameter sets, from the  \textit{14\_7} dataset.}
@@
     \includegraphics[scale = 0.37]{13_9_bestPlot}
-        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset. A}
+        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset.}
     \end{subfigure}
@@
 Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response.
-The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response}
+The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response \todo[inline]{colour forcings differntly, the 14 7 has the wrong forcing plotted, put a colorbar on the maps.}}
 \label{InitialResults}
@@
 Though the predictions are reasonable, they are substantially worse than the predictions for the data they were trained on.
-The first reason for this may be experiment to experiment variability of parameter values \cite{Hu2015}.
+The first reason for this may be that parameter values will vary between experimental runs \cite{Hu2015}.
 The second may be that the model is overfitting.
+\todo[inline]{follow this up properly.}
 
@@
 If this is the case, no amount of experimental data will allow us to estimate parameters, and we must consider reformulating the model.
-\footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product can ever be estimated, a problem no amount of data can fix.}
+\footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product $\beta_{1}\beta_{2}$ can ever be estimated, a problem no amount of data can fix.}
 
@@
 For example, it is possible that in the experimental regime we operate in, parameter effects may be weak, or highly correlated, but in general this is not true.
-\footnote{\cite{Beck} gives an example in which a parameter only effects model predictions after several hours, though others will affect it at all timescales - in this example, if we only took data for a few minutes, the parameter would be inestimable, but in principle it is not.} In this case, parameter estimates be may improved by taking data in more varied experimental conditions, and attempting to observe as many components of the model output as possible.
+\footnote{\cite{Beck} gives an example in which a parameter only affects model predictions after several hours, though others will affect it at all timescales - in this example, if we only took data for a few minutes, the parameter would be inestimable, but in principle it is not.} In this case, parameter estimates may be improved by taking data in more varied experimental conditions, and attempting to observe as many components of the model output as possible.
 
@@
 
-It can be shown that if sensitivity co-efficients are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
-Related to this fact, a number of measures have been proposed to asses parameter estimability from the sensitivity matrix \cite{Mclean2012}, the simplest of which is to simply plot the sensitivity co-effecients as a function of time, and visually check for obvious linear relations between the curves.
+It can be shown that if sensitivity coefficients are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
+Related to this fact, a number of measures have been proposed to assess parameter estimability from the sensitivity matrix \cite{Mclean2012}, the simplest of which is to plot the sensitivity coefficients as a function of time, and visually check for obvious linear relations between the curves.
 
@@
     \end{subfigure}
-    \caption{Sensitivity co-efficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset.
+    \caption{Sensitivity coefficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset.
 \ref{SensitivityMatrix_unscaled} shows them unscaled, \ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$.}
@@
 If this were the case, the particular shape of the models response curve would be primarily determined by the parameters appearing in eqs.
-\ref{eq:p}, \ref{eq:g}, \ref{eq:z} - namely $\mu$, $\beta$, $f_{s}$ and $\Theta$.
-If all the information contained in the parameters in eqs.\ref{eq:s} - \ref{eq:c} can only be expressed in the form of the models fixed point, it would not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
+\ref{eq:p} - \ref{eq:z} - namely $\mu$, $\beta$, $f_{s}$ and $\Theta$.
+If all the information contained in the parameters in eqs.\ref{eq:s} - \ref{eq:c} can only be expressed in the form of the model's fixed point, it would not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
 
@@
 It demonstrates that, at least in some of the parameter sets initially estimated, this difference in time scales exists.
-The fixed points of eqs.
+The fixed point of eqs.
 \ref{eq:s} - \ref{eq:c} affects eqs.
@@
 The first problem is local minima.
-Using the CMA-ES will give better results that a local minimisation algorithm, but you can never be sure you have found a global optimum - for example, \cite{Algorithms2003} compares several algorithms and finds, in the problem set, none locate the true global optimum.
+Using the CMA-ES will give better results than a local minimisation algorithm, but you can never be sure you have found a global optimum - for example, \cite{Algorithms2003} compares several algorithms and finds, in the problem set, none locate the true global optimum.
 A related problem is that of working with single points rather than distributions on parameter space.

### 2015-09-07 MScWriteup.tex
all but the last of manishe's changes made
beginning on annabelle's changes

@@
 
-\section{Model Derivation}
-\label{Model Derivation}
+\section{ODE Model}
+\label{ODE Model}
 
 In eqs.
-\ref{eq:s} - \ref{eq:z}, we propose a simple model consisting of a set of ODE's with mass action kinetics to describe the system \cite{UriAlon}.
+\ref{eq:s} - \ref{eq:z}, we present a modified version an existing model which describes the system, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}.
 Its state is given by the vector $(s,m,s:m,c,p,g,z)$, with all other variables representing model parameters.
 Tables \ref{ModelParameters} and \ref{StateVariables} give complete descriptions of the parameters and state variables.
+\todo[inline]{give the original model and discuss modifications in appendix.}
 
@@
 
- Control of the system by aTc and IPTG is modelled by $y(t)$ and $x(t)$ in eqs.
-\ref{eq:s}-\ref{eq:m}.
-$y(t)$ models the response of the sRNA transcription rate to a time varying aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}.
-Thus the forcing may vary between
- $N\frac{1}{f_{T}}$ and $N\frac{\alpha_{T}}{f}$, where $\alpha_{T}$ is the maximal transcription rate of the $P_{\mathrm{LtetO-1}}$ promoter.
-When engineering the system, many copies of the $P_{\mathrm{LtetO-1}}$ promoter, and section of DNA coding for sRNA, may be placed in the bacterial DNA - this is modelled by the copy number, $N$.
-Identical considerations hold for $x(t)$ and IPTG concentrations.
+ Control of the system by aTc is modelled by $y(t)$ in eq.
+\ref{eq:s}.
+This function models the response of the sRNA transcription rate to a time varying aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}.
+$\alpha_{T}$ is the maximal transcription rate of the $P_{\mathrm{LtetO-1}}$ promoter, and $N$ is the copy number, which models the fact that when engineering the system, many copies of the $P_{\mathrm{LtetO-1}}$ promoter may be placed in the bacterial DNA.
+Thus the transcription rate varies as a sigmoid bounded by $N\frac{1}{f_{T}}$ and $N\frac{\alpha_{T}}{f}$.
+Identical considerations hold for $x(t)$ in eq.
+\ref{eq:m}, and IPTG concentration.
 
@@
 \renewcommand{\arraystretch}{1.3}
+\caption{State Variables}
+\label{StateVariables}
+\centering
+\begin{tabular}{| l | l | l|}
+\hline \textbf{State variable} & Units &  \textbf{Definition}  \\
+\hline\hline $s$  & nM & sRNA concentration \\
+\hline $m$ & nM & mRNA concentration  \\
+\hline $s:m$ &  nM & Unstable sRNA:mRNA complex concentration  \\
+\hline $c$ &  nM & Stable sRNA:mRNA complex concentration  \\
+\hline $p$ & nM & Immature GFP concentration  \\
+\hline $g$ &  nm & Mature GFP concentration  \\
+\hline $z$ & AFU & Observed fluorescence  \\
+\hline $y(t)$ & & Unitless aTc forcing function  \\
+\hline $x(t)$ &  & Unitless IPTG forcing function  \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[h]
+\renewcommand{\arraystretch}{1.3}
 \caption{Model Parameters (those to be estimated shown in bold)}
@@
 
-\begin{table}[h]
-\renewcommand{\arraystretch}{1.3}
-\caption{State Variables}
-\label{StateVariables}
-\centering
-\begin{tabular}{| l | l | l|}
-\hline \textbf{State variable} & Units &  \textbf{Definition}  \\
-\hline\hline $s$  & nM & sRNA concentration \\
-\hline $m$ & nM & mRNA concentration  \\
-\hline $s:m$ &  nM & Unstable sRNA:mRNA complex concentration  \\
-\hline $c$ &  nM & Stable sRNA:mRNA complex concentration  \\
-\hline $p$ & nM & Immature GFP concentration  \\
-\hline $g$ &  nm & Mature GFP concentration  \\
-\hline $z$ & AFU & Observed fluorescence  \\
-\hline $y(t)$ & & Unitless aTc forcing function  \\
-\hline $x(t)$ &  & Unitless IPTG forcing function  \\
-\hline
-\end{tabular}
-\end{table}
-
 %%%
@@
 The model may be integrated numerically, giving a prediction $\mathbf{y}(t,\boldsymbol{\theta})$.
-An error between the model prediction and an experimental time series is defined as
+The least squares error between the model prediction and an experimental time series is defined as
 \begin{align}
-J(\boldsymbol{\theta}) = \sum_{i =1}^{N} (\mathbf{y}_{\mathrm{exp}}(t_{i}) - \mathbf{y}(t_{i},\boldsymbol{\theta}))^2
+J(\boldsymbol{\theta}) = \sum_{i =1}^{n} (\mathbf{y}_{\mathrm{exp}}(t_{i}) - \mathbf{y}(t_{i},\boldsymbol{\theta}))^2
 \end{align}
-where the experimental time series, $\mathbf{y}_{\mathrm{exp}}(t_{i})$ is recorded at timepoints $t_{i}$, $i = 1 \hdots N$.
+where the experimental time series, $\mathbf{y}_{\mathrm{exp}}(t_{i})$ is recorded at timepoints $t_{i}$, $i = 1 \hdots n$.
 This error function defines a landscape in $\boldsymbol{\theta}$ space, and we seek to minimise it by varying $\boldsymbol{\theta}$.
@@
 Our minimisation problem is thus
+\todo[inline]{explain why not fit each curve individually.}
 \begin{align}
-\min_{\boldsymbol{\theta}} \sum_{i =1}^{N} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2
+\min_{\boldsymbol{\theta}} \sum_{i =1}^{n} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2
 \end{align}
 The next step is performing the minimisation.
-In general, the landscape defined by the error function is multimodal, and may be very rugged.
+In general, the landscape defined by the error function may be very rugged, and contain many local minima.
 \footnote{this is true even if ODE model is linear in its parameters, as is almost the case for us - though the ODE model is linear, the resulting solutions are in general not.
@@
 Using this sample, it then estimates a new mean and covariance matrix for this gaussian and the process is iterated.}
+\todo[inline]{why use CMA-ES? Some refs}
 
@@
 Appendix \ref{Parameter literature review} contains a list of parameter values found in the literature, where available, and their reference, as well as initial bounds placed on parameters not found in the literature.
-To further reduce the search space, we assume that $\delta_m$, $\delta_{sm}$ and $\delta_c$ all take similar values, and set them equal.
+To further reduce the search space, we simplify the model by assuming that $\delta_m$, $\delta_{sm}$ and $\delta_c$ all take similar values, and set them equal.
 After this is done, we are left with a 9 dimensional search space, bounded by a hypercube (parameters to be estimated are shown in bold in table \ref{ModelParameters}).
@@
     \centering
-        \includegraphics[scale = 0.3, clip = true, trim = 80 0 74 0]{13_9_hist.eps}
+        \includegraphics[scale = 0.38, clip = true, trim = 80 0 74 0]{13_9_hist.eps}
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset.}
+        \label{InitialResults_13_9_hist}
     \end{subfigure}
@@
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{14\_7} dataset. }
+        \label{InitialResults_14_7_hist}
     \end{subfigure}
@@
         \caption{Correlation matrix of estimated parameter sets, from the  \textit{13\_9} dataset. }
+        \label{InitialResults_13_9_correlation}
     \end{subfigure}
@@
         \caption{Correlation matrix of estimated parameter sets, from the  \textit{14\_7} dataset.}
+        \label{InitialResults_14_7_correlation}
     \end{subfigure}
@@
         \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset.}
+                \label{InitialResults_13_9_best}
     \end{subfigure}
@@
         \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{14\_7} dataset.}
+        \label{InitialResults_14_7_best}
     \end{subfigure}
@@
 Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response.
-The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response \todo[inline]{colour forcings differntly, the 14 7 has the wrong forcing plotted, put a colorbar on the maps.}}
+The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response.}
 \label{InitialResults}
@@
 
- We begin by fitting two of the datasets, \textit{13\_9} and \textit{14\_7}, by choosing 200 points, uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
+ We begin by fitting each dataset individually, by choosing 200 parameter sets uniformly distributed over our initial parameter bounds and running the CMA-ES starting from them.
 Results are shown in Figs.
 \ref{InitialResults}, \ref{hist_f}.
-Fig.
-\ref{InitialResults} demonstrates that the model is capable of quantitatively capturing the data - however, the fitting also suggests that some parameters are not tightly constrained, taking values right across the initial bounding range specified.
-The correlation matrices in Fig.
-\ref{InitialResults} also indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
+Figs.
+\ref{InitialResults_13_9_best}, \ref{InitialResults_14_7_best}  demonstrate that the model is capable of quantitatively capturing the data.
+However, histograms of estimated parameter values found from fitting the above 200 sets indicate many parameter estimates are not tightly constrained, taking values right across the initial bounding range specified - in particular, $k_\mathrm{off}$, $\delta_{m}$ and $\delta_{s}$ are very weakly constrained, with substantial numbers of runs constrained only by the initial bounding box.
+By contrast, some parameter estimates ($\mu$ and $\Theta$ in particular) are much tighter.
+
+Figs.
+\ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices, computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
+An element of the matrix $R_{ij}$ is then the correlation co-efficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters.
+These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
 Referring to eq.
 \ref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
+Biologically, an increased production rate of GFP being balance by an increased dilution rate.
 The results suggest a fitness landscape relatively flat to perturbations in certain combinations of parameters, and indicate we may have difficulty obtaining unique estimates of the model parameters.
@@
 
- Taken together, these results suggest that some parameters may be inestimable, and that the data may be better described by a simplified model.
+ Taken together, these results suggest that some parameters may be inestimable.  
+ \todo[inline]{move this down into the discussion, and that the data may be better described by a simplified model}
 
@@
 Problems may also arise for more practical reasons (\textit{practical identifiability} \cite{Mclean2012}).
-For example, it is possible that in the experimental regime we operate in, parameter effects may be weak, or highly correlated, but in general this is not true.
+For example, it is possible that in the experimental regime we operate in, parameter effects may be weak, or highly correlated, but in other regimes this is not true.
 \footnote{\cite{Beck} gives an example in which a parameter only affects model predictions after several hours, though others will affect it at all timescales - in this example, if we only took data for a few minutes, the parameter would be inestimable, but in principle it is not.} In this case, parameter estimates may be improved by taking data in more varied experimental conditions, and attempting to observe as many components of the model output as possible.
@@
 where $S_{ij}$ is the derivative of the observed fluorescence, $z$, with parameter $\theta_{j}$, evaluated at timepoint $t_{i}$.
-This matrix contains information about how sensitive $z$ is to perturbations in parameter values, and at what times it is most sensitive.
+Each column of $S$ is a time series of sensitivity co-efficients, which describe how sensitive $z$ is to perturbations in the parameter associated with that column, and at what times it is most sensitive.
 $\hat{\theta}_{j}$ is the value of the parameter that the derivative is being evaluated at.
@@
 
-It can be shown that if sensitivity coefficients are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
+It can be shown that if the sensitivity coefficients (columns of $S$) are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
 Related to this fact, a number of measures have been proposed to assess parameter estimability from the sensitivity matrix \cite{Mclean2012}, the simplest of which is to plot the sensitivity coefficients as a function of time, and visually check for obvious linear relations between the curves.
@@
 Fig.
-\ref{SensitivityMatrix} shows plots of each column of $S$, evaluated at the parameter set giving the lowest error in the \textit{13\_9} dataset - these show the evolution of the sensitivity matrix over time.
+\ref{SensitivityMatrix} shows plots of each column of $S$, evaluated at the parameter set giving the lowest error in the \textit{13\_9} dataset.
 Fig.
@@
 
-We see that many of the parameters give sensitivity curves of similar shapes - the effects of perturbing any one of these parameters all look similar in terms of model output.
-This may help to explain why some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
+We see that many of the parameters give sensitivity curves of similar shapes, and have near linear dependence - this implies that the effects of perturbing any one of these parameters all look similar in terms of model output, and are hard to distinguish between.
+This may help to explain why some of our initial parameter estimates are very loose, and why there are correlations between the estimated parameter sets - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
 As such, we should view estimates of these parameters with extreme caution.
@@
 \ref{InitialResults}.
+
+\subsection{Differing timescales within the system}
 

### 2015-09-08 MScWriteup.tex
more of annabelle's changes

@@
 
-\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e, epstopdf, caption, subcaption,centernot}
+\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e, epstopdf, subcaption,centernot}
+\usepackage[font=small,labelfont = sf, textfont=sf,labelfont=bf]{caption}
+
+\newcounter{MYtempeqncnt}
 
@@
 Its state is given by the vector $(s,m,s:m,c,p,g,z)$, with all other variables representing model parameters.
-Tables \ref{ModelParameters} and \ref{StateVariables} give complete descriptions of the parameters and state variables.
+Tables \ref{StateVariables} and \ref{ModelParameters} give complete descriptions of the parameters and state variables.
 \todo[inline]{give the original model and discuss modifications in appendix.}
@@
     \centering
-        \includegraphics[scale = 0.38, clip = true, trim = 80 0 74 0]{13_9_hist.eps}
+        \includegraphics[scale = 0.27, clip = true, trim = 120 0 125 0]{13_9_hist}
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset.}
@@
     \centering
-        \includegraphics[scale = 0.38,clip = true, trim = 60 12 74 0]{14_7_hist.eps}
+        \includegraphics[scale = 0.3,clip = true, trim = 120 0 125 0]{14_7_hist}
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{14\_7} dataset. }
@@
 
-We suggest that part of the issue is the step function forcing used in model predictions, and the timescales on which the system responds to it.
-We hypothesize that the system may have two timescales in it - a fast timescale in which eqs.
+Fig.
+\ref{model_output} shows model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
+We see that $s,m,s:m$ and $c$ respond rapidly to the forcing function, flipping between the two fixed points defined by the step function forcing almost instantly.
+By contrast, there is delay in the response of $p$ and $g$, on a timescale comparable to the period of the forcing.
+This result suggests that, at least in some of the parameter sets initially estimated, the system may have two timescales in it - a fast timescale in which eqs.
 \ref{eq:s} - \ref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in eqs.
 \ref{eq:p}, \ref{eq:g}, in which measured fluorescence changes in response to the forcing.
-If this is the case, then it may be that our experimental data can only probe the system via the fixed point of eqs.
-\ref{eq:s} - \ref{eq:c} , and that the parameters contained within those equations can only work to set the particular fixed point values of $p$ and $g$ that the system tends toward.
-If this were the case, the particular shape of the models response curve would be primarily determined by the parameters appearing in eqs.
-\ref{eq:p} - \ref{eq:z} - namely $\mu$, $\beta$, $f_{s}$ and $\Theta$.
-If all the information contained in the parameters in eqs.\ref{eq:s} - \ref{eq:c} can only be expressed in the form of the model's fixed point, it would not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
-
-Fig. \ref{model_output} shows model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
 
@@
 
-It demonstrates that, at least in some of the parameter sets initially estimated, this difference in time scales exists.
-The fixed point of eqs.
-\ref{eq:s} - \ref{eq:c} affects eqs.
-\ref{eq:p}, \ref{eq:g} through the $\beta m +f_{s}\beta c$ term - we would expect that, if the fixed point value of this term is what is relevant in determining model response, there would be some consistency in its value across the parameter sets found.
-Fig.
+If this is the case, then it may be that our experimental data can only probe the system via the fixed point of eqs.
+\ref{eq:s} - \ref{eq:c} , and that the parameters contained within those equations can only work to set the particular fixed point values of $p$ and $g$ that the system tends toward.
+This would explain the similarity between the sensitivity curves of many parameters - if the parameters in eqs.\ref{eq:s} - \ref{eq:c} can only act to alter the model's fixed point, they would all alter the sensitivity curve in exactly the same way, since a small perturbation in one of these parameters would only act to perturb the location of the models fixed points.
+It would then not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
+
+The fixed point values of $s,m, s:m$ and $c$ affect eqs.
+\ref{eq:p}, \ref{eq:g} through the $\beta m +f_{s}\beta c$ term in \ref{eq:p}.
+Explicit forms for $m$ and $c$ are given in eqs.
+\ref{eq:m_fixed}, \ref{eq:c_fixed}.
+The fact that this term is  the only one in eqs.\ref{eq:p}, \ref{eq:g} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in eqs.
+\ref{eq:s} - \ref{eq:c} - all that $p$ and $g$ see is this $\beta m +f_{s}\beta c$ flipping between its two fixed point values, so any parameters that only enter the system via this term will have indistinguishable roles.
+
+Although we have seen that parameter sets with similar least squares error values can contain very different parameter values, this explanation implies that we would expect similar values of $\beta m +f_{s}\beta c$ across all the sets.Fig.
 \ref{fixedpoint_f} shows a scatterplot of model fixed point values against error function value, and demonstrates that, though individual parameter values can be spread across very large ranges, they are correlated in such a way as to give similar model fixed points.
+
+In Fig.
+\ref{fixedpoint_mu}, we also see a strong positive correlation between the fixed point values and the values of $\mu$ ($R = 0.9854$ for the visually tightly clustered data).
+Taken together with the relatively tightly constrained values of $\mu$ found in Fig.
+\ref{InitialResults}, these results suggest that in order to minimize error, the algorithm is effectively trying to vary $\beta m +f_{s}\beta c$ and $\mu$, and finding a `trench' of highly correlated values, which is shown in Fig.
+\ref{fixedpoint_mu}.
+
+These results further suggest that, while we are trying to minimise model error over our initial high dimensional space, we are effectively working in a lower dimensional space, in which one dimension is the value of the models $\beta m +f_{s}\beta c$ .
+
+\begin{figure*}[!t]
+% ensure that we have normalsize text
+\tiny
+% Store the current equation number.
+\setcounter{MYtempeqncnt}{\value{equation}}
+% Set the equation number to one less than the one
+% desired for the first equation here.
+% The value here will have to changed if equations
+% are added or removed prior to the place these
+% equations are referenced in the main text.
+\setcounter{equation}{5}
+\begin{equation}
+\label{eq:m_fixed}
+m\to -\frac{\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}-(\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)}{2 k_{\text{on}} \left(\mu +\delta _m\right)
+   \left(k_{\text{hyb}}+\delta _m\right)}
+\end{equation}
+\begin{equation}
+\label{eq:c_fixed}
+\frac{k_{\text{hyb}} \left(\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)\right)}{2 k_{\text{on}} \left(\mu +\delta
+   _m\right) \left(k_{\text{hyb}}+\delta _m\right){}^2}
+\end{equation}
+% Restore the current equation number.
+\setcounter{equation}{\value{MYtempeqncnt}}
+% IEEE uses as a separator
+\hrulefill
+% The spacer can be tweaked to stop underfull vboxes.
+\vspace*{4pt}
+\end{figure*}
 
@@
 \end{figure}
-
-In Fig.
-\ref{fixedpoint_mu}, we also see a strong positive correlation between the fixed point values and the values of $\mu$ ($R = 0.9854$ for the visually tightly clustered data).
-Taken together with the relatively tightly constrained values of $\mu$ found in Fig.
-\ref{InitialResults}, these results suggest that in order to minimize error, the algorithm is effectively trying to vary $\beta m +f_{s}\beta c$ and $\mu$, and finding a `trench' of highly correlated values, which is shown in Fig.
-\ref{fixedpoint_mu}.
-
-These results further suggest that, while we are trying to minimise model error over our initial high dimensional space, we are effectively working in a lower dimensional space, in which one dimension is the value of the models $\beta m +f_{s}\beta c$ .
 
@@
 
+\section{Model Fixed Point}
+\label{ModelFixedPoint}
+\setcounter{figure}{0}
+
+\begin{figure*}[!t]
+% ensure that we have normalsize text
+\tiny
+% Store the current equation number.
+\setcounter{MYtempeqncnt}{\value{equation}}
+% Set the equation number to one less than the one
+% desired for the first equation here.
+% The value here will have to changed if equations
+% are added or removed prior to the place these
+% equations are referenced in the main text.
+\setcounter{equation}{5}
+\begin{equation}
+\label{eqn_dbl_x}
+y\to \frac{\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)}{2 k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right){}^2}
+\end{equation}
+\begin{equation}
+\label{eqn_dbl_y}
+m\to -\frac{\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}-(\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)}{2 k_{\text{on}} \left(\mu +\delta _m\right)
+   \left(k_{\text{hyb}}+\delta _m\right)}
+\end{equation}
+\begin{equation}
+\label{eqn_dbl_y}
+s\to -\frac{\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}+(\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)}{2 k_{\text{on}} \left(\mu +\delta _s\right)
+   \left(k_{\text{hyb}}+\delta _m\right)}
+\end{equation}
+\begin{equation}
+\frac{k_{\text{hyb}} \left(\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)\right)}{2 k_{\text{on}} \left(\mu +\delta
+   _m\right) \left(k_{\text{hyb}}+\delta _m\right){}^2}
+\end{equation}
+% Restore the current equation number.
+\setcounter{equation}{\value{MYtempeqncnt}}
+% IEEE uses as a separator
+\hrulefill
+% The spacer can be tweaked to stop underfull vboxes.
+\vspace*{4pt}
+\end{figure*}
+
 \end{document}

### 2015-09-08 MScWriteup.tex
tweaking sizes and labels

@@
 \usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e, epstopdf, subcaption,centernot}
-\usepackage[font=small,labelfont = sf, textfont=sf,labelfont=bf]{caption}
+\usepackage[font=footnotesize,labelfont = sf, textfont=sf,labelfont=bf]{caption,subcaption}
 
@@
 In the remainder of this section we review the regulatory system we will consider, and discuss recent single cell fluorescence experiments performed on it.
-In section \ref{Model Derivation} we propose a set of ODE's to model the system.
+In section \ref{ODEModel} we introduce a set of ODE's to model the system.
 In section \ref{Parameter Estimation} we attempt to estimate its unknown parameters by fitting the model to time series data.
@@
 
-mRNA concentrations in this system shown in Fig.
+mRNA concentrations in the system shown in Fig.
 \ref{ANDGate} can be indirectly observed by designing the mRNA to code for GFP \footnote{ The methodology of \cite{Rodrigo2012} only optimises the 5' UTR of the mRNA, so the actual protein being coded for is unimportant.}.
@@
 
-A single layer of the bacteria are grown in rows of chambers.
+The experimental setup is as follows (Fig.
+\ref{ExperimentalSetup}): A single layer of the bacteria are grown in rows of chambers.
 A medium constantly flows through these chambers, allowing normal feeding of the bacteria, and the introduction of aTc or IPTG.
@@
 \begin{figure}[h]
+\begin{subfigure}[h]{0.49\textwidth}
     \centering
 		\includegraphics[trim = 280 0 0  85 ,page=17 ,clip = true, scale=0.4]{WCPM_Warwick-talk45-220115.pdf}
+        \caption{An overview of the entire setup, including bacterial growth chambers (labelled `Microfabricated Microchemstat device'), the microscope imaging them, and the software constructing fluorescence time series.
+}
+        \label{InitialResults_13_9_hist}
+    \end{subfigure}
+    \begin{subfigure}[c]{0.49\textwidth}
     \centering
 	\includegraphics[trim = 0 0 0  90 ,page=18 ,clip = true, scale=0.35]{WCPM_Warwick-talk45-220115.pdf}
+        \caption{A diagram of the bacterial growth chambers.
+The schematic shows the chambers themselves, and the medium inlet where aTc and IPTG are let in.
+Shown also are photographs of a row of chambers, and an individual chamber with bacteria growing in it.}
+        \label{InitialResults_14_7_hist}
+    \end{subfigure}
     \caption{A schematic of an experimental setup which allows single cell fluorescences to be recorded over time.
@@
 Image reproduced from \cite{Jaramillo}.}
-\label{}
+    \label{ExperimentalSetup}
 \end{figure}
@@
 \section{ODE Model}
-\label{ODE Model}
+\label{ODEModel}
 
@@
 Degradation of the immature and mature GFP is modelled in two ways.
-Firstly a generic degradation rate $\delta_{g}$, assumed identical for the mature and immature species, with the dilution rate $\mu$ shared by all species.
-\todo[inline]{need a description about the terms representing clpX degradation.}
-Finally, eq. \ref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence, assuming a linear response.
+Firstly a generic degradation rate $\delta_{g}$ is included, assumed identical for the mature and immature species, along with the dilution rate $\mu$ shared by all species.
+Secondly, in the experimental setup we can arrange for GFP molecules to be produced with a \textit{degradation tag} attached to them \cite{Hersch2004}.
+This tag is sought out by a protein, ClpX, which will then degrade the molecule the tag is  attached to.
+This degradation process is modelled by the final terms in eqs.
+\ref{eq:p} and \ref{eq:g}.
+Finally, eq.
+\ref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence, assuming a linear response.
 
@@
     \centering
-        \includegraphics[scale = 0.27, clip = true, trim = 120 0 125 0]{13_9_hist}
+        \includegraphics[scale = 0.24, clip = true, trim = 140 0 125 0]{13_9_hist}
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset.}
@@
     \centering
-        \includegraphics[scale = 0.3,clip = true, trim = 120 0 125 0]{14_7_hist}
+        \includegraphics[scale = 0.24,clip = true, trim = 140 0 125 0]{14_7_hist}
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{14\_7} dataset. }
@@
     \centering
-    \includegraphics[scale = 0.37]{13_9_bestPlot}
+    \includegraphics[scale = 0.26,clip = true, trim = 90 0 70 0]{13_9_bestPlot}
         \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset.}
@@
     \centering
-        \includegraphics[scale = 0.37]{14_7_bestPlot}
+        \includegraphics[scale = 0.26,clip = true, trim = 90 0 70 0]{14_7_bestPlot}
         \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{14\_7} dataset.}
@@
     \centering
-        \includegraphics[scale = 0.26]{14_7_pred_13_9_data}
+        \includegraphics[scale = 0.3,clip = true, trim = 87 0 70 0]{14_7_pred_13_9_data}
         \caption{\textit{14\_7} prediction \textit{13\_9} data}
@@
     \centering
-        \includegraphics[scale = 0.26]{13_9_pred_14_7_data}
+        \includegraphics[scale = 0.3,clip = true, trim = 87 0 70 0]{13_9_pred_14_7_data}
         \caption{\textit{13\_9} prediction \textit{14\_7} data}
@@
  \todo[inline]{move this down into the discussion, and that the data may be better described by a simplified model}
+
+ \begin{figure*}[!t]
+% ensure that we have normalsize text
+\tiny
+% Store the current equation number.
+\setcounter{MYtempeqncnt}{\value{equation}}
+% Set the equation number to one less than the one
+% desired for the first equation here.
+% The value here will have to changed if equations
+% are added or removed prior to the place these
+% equations are referenced in the main text.
+\setcounter{equation}{11}
+\begin{align}
+\label{eq:m_fixed}  
+m &= \frac{1}{{2 k_{\text{on}} \left(\mu +\delta _m\right)
+   \left(k_{\text{hyb}}+\delta _m\right)}}
+   \Bigg(\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2} \\ \nonumber  
+ &- (\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)\Bigg)
+\end{align}
+\begin{align}
+\label{eq:c_fixed}
+c &= \frac{k_{\text{hyb}}}{2 k_{\text{on}} \left(\mu +\delta
+   _m\right) \left(k_{\text{hyb}}+\delta _m\right){}^2}
+   \Bigg( \sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2} \\ \nonumber
+   &+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right) \Bigg)
+\end{align}
+% Restore the current equation number.
+\setcounter{equation}{\value{MYtempeqncnt}}
+% IEEE uses as a separator
+\hrulefill
+% The spacer can be tweaked to stop underfull vboxes.
+\vspace*{4pt}
+\end{figure*}
 
@@
     \centering
-        \includegraphics[scale = 0.3]{Sensitivity_unscaled}
+        \includegraphics[scale = 0.35,clip = true, trim = 30 0 60 0]{Sensitivity_unscaled}
         \caption{Unscaled}
@@
     \centering
-        \includegraphics[scale = 0.3]{Sensitivity_scaled}
+        \includegraphics[scale = 0.35,clip = true, trim = 15 0 60 0]{Sensitivity_scaled}
         \caption{Scaled}
@@
 These results further suggest that, while we are trying to minimise model error over our initial high dimensional space, we are effectively working in a lower dimensional space, in which one dimension is the value of the models $\beta m +f_{s}\beta c$ .
-
-\begin{figure*}[!t]
-% ensure that we have normalsize text
-\tiny
-% Store the current equation number.
-\setcounter{MYtempeqncnt}{\value{equation}}
-% Set the equation number to one less than the one
-% desired for the first equation here.
-% The value here will have to changed if equations
-% are added or removed prior to the place these
-% equations are referenced in the main text.
-\setcounter{equation}{5}
-\begin{equation}
-\label{eq:m_fixed}
-m\to -\frac{\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
-   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
-   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}-(\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)}{2 k_{\text{on}} \left(\mu +\delta _m\right)
-   \left(k_{\text{hyb}}+\delta _m\right)}
-\end{equation}
-\begin{equation}
-\label{eq:c_fixed}
-\frac{k_{\text{hyb}} \left(\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
-   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
-   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)\right)}{2 k_{\text{on}} \left(\mu +\delta
-   _m\right) \left(k_{\text{hyb}}+\delta _m\right){}^2}
-\end{equation}
-% Restore the current equation number.
-\setcounter{equation}{\value{MYtempeqncnt}}
-% IEEE uses as a separator
-\hrulefill
-% The spacer can be tweaked to stop underfull vboxes.
-\vspace*{4pt}
-\end{figure*}
 

### 2015-09-08 MScWriteup.tex
moving figures around

@@
 For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA is an attractive tool, due to the ease of predicting its behaviour from physiochemical models.
-In this report, we propose a system of ODE's to model a recently designed synthetic regulatory circuit, before attempting to estimate its unknown parameters using recent time series data.
+In this report, we introduce a system of ODE's to model a recently designed synthetic regulatory circuit, before attempting to estimate its unknown parameters using recent time series data.
 We provide estimates for a subset of parameters, but find many of them to be inestimable.
@@
 
-In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the UTR), upstream of the RBS \footnote{Both DNA and RNA are directional - the backbone of the molecule is not symmetric.
-This gives the molecule two distinct ends, denoted 5' and 3'.
-Translation can only occur in the 5' to 3' direction, hence the meaning of `upstream'}.
-Translation may be self repressed by this `tail' folding over and binding across the RBS, forming a  loop in the mRNA and preventing the Ribosome from binding (Fig.
-\ref{RBS}).
-This self repression may be released with an sRNA which binds to this looped tail - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind.
-In summary, the presence of the sRNA positively regulates gene expresssion.
-
-\begin{figure*}[t]
+\begin{figure*}[t!]
 \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centering
@@
 \end{figure}
+
+In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the UTR), upstream of the RBS \footnote{Both DNA and RNA are directional - the backbone of the molecule is not symmetric.
+This gives the molecule two distinct ends, denoted 5' and 3'.
+Translation can only occur in the 5' to 3' direction, hence the meaning of `upstream'}.
+Translation may be self repressed by this `tail' folding over and binding across the RBS, forming a  loop in the mRNA and preventing the Ribosome from binding (Fig.
+\ref{RBS}).
+This self repression may be released with an sRNA which binds to this looped tail - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind.
+In summary, the presence of the sRNA positively regulates gene expresssion.
 
@@
 
-The experimental setup is as follows (Fig.
-\ref{ExperimentalSetup}): A single layer of the bacteria are grown in rows of chambers.
+The experimental setup is as follows (Figs.
+\ref{ExperimentalSetup_overview},  \ref{ExperimentalSetup_chambers}): A single layer of the bacteria are grown in rows of chambers.
 A medium constantly flows through these chambers, allowing normal feeding of the bacteria, and the introduction of aTc or IPTG.
@@
 
-\begin{figure}[h]
-\begin{subfigure}[h]{0.49\textwidth}
+\begin{figure}
     \centering
 		\includegraphics[trim = 280 0 0  85 ,page=17 ,clip = true, scale=0.4]{WCPM_Warwick-talk45-220115.pdf}
-        \caption{An overview of the entire setup, including bacterial growth chambers (labelled `Microfabricated Microchemstat device'), the microscope imaging them, and the software constructing fluorescence time series.
+        \caption{An overview of the experimental setup which allows single cell fluorescences to be recorded over time.
+Shown are the bacterial growth chambers (labelled `Microfabricated Microchemstat device'), the microscope imaging them, and the software constructing fluorescence time series.
+Image reproduced from \cite{Jaramillo}.
 }
-        \label{InitialResults_13_9_hist}
-    \end{subfigure}
-    \begin{subfigure}[c]{0.49\textwidth}
+            \label{ExperimentalSetup_overview}
+\end{figure}
+\begin{figure}
     \centering
 	\includegraphics[trim = 0 0 0  90 ,page=18 ,clip = true, scale=0.35]{WCPM_Warwick-talk45-220115.pdf}
-        \caption{A diagram of the bacterial growth chambers.
+        \caption{A diagram of the bacterial growth chambers shown in Fig.
+\ref{ExperimentalSetup_overview}.
 The schematic shows the chambers themselves, and the medium inlet where aTc and IPTG are let in.
-Shown also are photographs of a row of chambers, and an individual chamber with bacteria growing in it.}
-        \label{InitialResults_14_7_hist}
-    \end{subfigure}
-    \caption{A schematic of an experimental setup which allows single cell fluorescences to be recorded over time.
-Bacteria are grown in chambers, with a constant supply of medium flowing through them, which allows the introduction of aTc and IPTG.
-Software then traces these cells, allowing fluorescence timesseries to be built.
+Shown also are photographs of a row of chambers, and an individual chamber with bacteria growing in it.
 Image reproduced from \cite{Jaramillo}.}
-    \label{ExperimentalSetup}
+    \label{ExperimentalSetup_chambers}
 \end{figure}
@@
 
-In eqs.
-\ref{eq:s} - \ref{eq:z}, we present a modified version an existing model which describes the system, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}.
+In \eqref{eq:s} - \eqref{eq:z}, we present a modified version an existing model which describes the system, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}.
 Its state is given by the vector $(s,m,s:m,c,p,g,z)$, with all other variables representing model parameters.
@@
 \end{align}
+
 Based on the reaction mechanism in Fig.
-\ref{reactionscheme}, the hybridization of the sRNA and mRNA first into an unstable complex, then a stable one, is modelled in eqs.
-\ref{eq:s}-\ref{eq:c}.
-The initial binding is modelled as a reversible reaction with forward and backward rates $k_\mathrm{on}$ and $k_\mathrm{off}$.
+\ref{reactionscheme}, the hybridization of the sRNA and mRNA first into an unstable complex, then a stable one, is modelled in \eqref{eq:s} - \eqref{eq:c}.
+The initial binding is modelled as a reversible reaction with forward and backward rates $k_\mathrm{on}$ and $k_\mathrm{off}$:
 
@@
 sRNA \chemsign+ mRNA\chemrel[$k_\mathrm{on}$][$k_\mathrm{off}$]{<>} sRNA:mRNA\textsubscript{unstable}
-}
+},
 \end{align*}
@@
 sRNA:mRNA\textsubscript{unstable} \chemrel[$k_\mathrm{hyb}$]{->} sRNA:mRNA\textsubscript{stable} 
-}
+}.
 \end{align*}
@@
 
- Control of the system by aTc is modelled by $y(t)$ in eq.
-\ref{eq:s}.
+ Control of the system by aTc is modelled by $y(t)$ in \eqref{eq:s}.
 This function models the response of the sRNA transcription rate to a time varying aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}.
@@
 Thus the transcription rate varies as a sigmoid bounded by $N\frac{1}{f_{T}}$ and $N\frac{\alpha_{T}}{f}$.
-Identical considerations hold for $x(t)$ in eq.
-\ref{eq:m}, and IPTG concentration.
+Identical considerations hold for $x(t)$ in \eqref{eq:m}, and IPTG concentration.
 
-We explicitly model translation as a simple one step process in eqs.
-\ref{eq:p}-\ref{eq:z}.
+We explicitly model translation as a simple one step process in \eqref{eq:p} - \eqref{eq:z}.
 There is a small rate of translation of the self repressed mRNA \cite{Rodrigo2012}, which is modelled at rate $\beta$, and a larger one for translation of the stable complex, $\beta f_s$.
-Here $f_s$ represents the fractional change in translation rate between the repressed mRNA and the unrepressed complex.
+Here $f_s$ represents the fractional change in translation rate between the repressed mRNA and the unrepressed complex:
 
@@
 sRNA:mRNA\textsubscript{Stable} \chemrel[$\beta f_{s}$]{->} GFP\textsubscript{Immature}
-}
+},
 \end{align*}
-
 \begin{align*}
@@
 mRNA \chemrel[$\beta$ ]{->} GFP\textsubscript{Immature}
-}
+}.
 \end{align*}
 
-Initially, the translated GFP is in an immature state, and will not fluoresce. To account for this, we include a maturation rate, $\gamma$
+Initially, the translated GFP is in an immature state, and will not fluoresce. To account for this, we include a maturation rate, $\gamma$:
 
@@
 GFP\textsubscript{Immature}  \chemrel[$\gamma$]{->} GFP\textsubscript{Mature}
-}
+}.
 \end{align*}
@@
 This tag is sought out by a protein, ClpX, which will then degrade the molecule the tag is  attached to.
-This degradation process is modelled by the final terms in eqs.
-\ref{eq:p} and \ref{eq:g}.
-Finally, eq.
-\ref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence, assuming a linear response.
+This degradation process is modelled by the final terms in \eqref{eq:p} and \eqref{eq:g}.
+Finally, \eqref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence, assuming a linear response.
 
@@
 \begin{align}
-\frac{\mathrm{d}\mathbf{y}}{\mathrm{d}t} = \mathbf{f}(\mathbf{y},\boldsymbol{\theta},t)
+\frac{\mathrm{d}\mathbf{y}}{\mathrm{d}t} = \mathbf{f}(\mathbf{y},\boldsymbol{\theta},t),
 \end{align}
@@
 \begin{align}
-J(\boldsymbol{\theta}) = \sum_{i =1}^{n} (\mathbf{y}_{\mathrm{exp}}(t_{i}) - \mathbf{y}(t_{i},\boldsymbol{\theta}))^2
+J(\boldsymbol{\theta}) = \sum_{i =1}^{n} (\mathbf{y}_{\mathrm{exp}}(t_{i}) - \mathbf{y}(t_{i},\boldsymbol{\theta}))^2,
 \end{align}
@@
 \begin{align}
-\min_{\boldsymbol{\theta}} \sum_{i =1}^{n} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2
+\min_{\boldsymbol{\theta}} \sum_{i =1}^{n} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2.
 \end{align}
@@
 These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
-Referring to eq.
-\ref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
+Referring to \eqref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
 Biologically, an increased production rate of GFP being balance by an increased dilution rate.
@@
  \begin{align}
-S_{ij} = \hat{\theta}_{j} \frac{\partial z}{ \partial \theta_{j}}\Bigr|_{t_{i}}
+S_{ij} = \hat{\theta}_{j} \frac{\partial z}{ \partial \theta_{j}}\Bigr|_{t_{i}},
 \end{align}
-
 where $S_{ij}$ is the derivative of the observed fluorescence, $z$, with parameter $\theta_{j}$, evaluated at timepoint $t_{i}$.
@@
 By contrast, there is delay in the response of $p$ and $g$, on a timescale comparable to the period of the forcing.
-This result suggests that, at least in some of the parameter sets initially estimated, the system may have two timescales in it - a fast timescale in which eqs.
-\ref{eq:s} - \ref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in eqs.
-\ref{eq:p}, \ref{eq:g}, in which measured fluorescence changes in response to the forcing.
+This result suggests that, at least in some of the parameter sets initially estimated, the system may have two timescales in it - a fast timescale in which \eqref{eq:s} - \eqref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in \eqref{eq:p}, \eqref{eq:g}, in which measured fluorescence changes in response to the forcing.
 
@@
     \centering
-        \includegraphics[scale = 0.22]{model_output}
+        \includegraphics[scale = 0.26,clip = true, trim = 90 0 60 0]{model_output}
         \caption{model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
@@
 
-If this is the case, then it may be that our experimental data can only probe the system via the fixed point of eqs.
-\ref{eq:s} - \ref{eq:c} , and that the parameters contained within those equations can only work to set the particular fixed point values of $p$ and $g$ that the system tends toward.
-This would explain the similarity between the sensitivity curves of many parameters - if the parameters in eqs.\ref{eq:s} - \ref{eq:c} can only act to alter the model's fixed point, they would all alter the sensitivity curve in exactly the same way, since a small perturbation in one of these parameters would only act to perturb the location of the models fixed points.
+If this is the case, then it may be that our experimental data can only probe the system via the fixed point of \eqref{eq:s} - \eqref{eq:c} , and that the parameters contained within those equations can only work to set the particular fixed point values of $p$ and $g$ that the system tends toward.
+This would explain the similarity between the sensitivity curves of many parameters - if the parameters in \eqref{eq:s} - \eqref{eq:c} can only act to alter the model's fixed point, they would all alter the sensitivity curve in exactly the same way, since a small perturbation in one of these parameters would only act to perturb the location of the models fixed points.
 It would then not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
 
-The fixed point values of $s,m, s:m$ and $c$ affect eqs.
-\ref{eq:p}, \ref{eq:g} through the $\beta m +f_{s}\beta c$ term in \ref{eq:p}.
-Explicit forms for $m$ and $c$ are given in eqs.
-\ref{eq:m_fixed}, \ref{eq:c_fixed}.
-The fact that this term is  the only one in eqs.\ref{eq:p}, \ref{eq:g} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in eqs.
-\ref{eq:s} - \ref{eq:c} - all that $p$ and $g$ see is this $\beta m +f_{s}\beta c$ flipping between its two fixed point values, so any parameters that only enter the system via this term will have indistinguishable roles.
+The fixed point values of $s,m, s:m$ and $c$ affect \eqref{eq:p}, \eqref{eq:g} through the $\beta m +f_{s}\beta c$ term in \eqref{eq:p}.
+Explicit forms for $m$ and $c$ are given in \eqref{eq:m_fixed}, \eqref{eq:c_fixed}.
+The fact that this term is  the only one in \eqref{eq:p}, \eqref{eq:g} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in \eqref{eq:s} - \eqref{eq:c} - all that $p$ and $g$ see is this $\beta m +f_{s}\beta c$ flipping between its two fixed point values, so any parameters that only enter the system via this term will have indistinguishable roles.
 
@@
  Thus, further work may consist of tightening the bounds on parameter values, carrying out a structural analysis, and possibly simplifying the model.
-A much simpler model, consisting of  just eqs.
-\ref{eq:p} \ref{eq:g}, may explain existing data perfectly well without the need for inestimable parameters.
+A much simpler model, consisting of  just \eqref{eq:p} and \eqref{eq:g}, may explain existing data perfectly well without the need for inestimable parameters.
 Experimentally, one option  may be to use different forcing functions.
@@
 Some earlier fitting work has been done on this dataset via a similar method to the one used above.
-Our results suggest that those estimates should be viewed with caution - the estimates will not be unique, and the values of some parameters (those in eqs.
-\ref{eq:s} - \ref{eq:c}) may be very far off any values which are eventually found.
+Our results suggest that those estimates should be viewed with caution - the estimates will not be unique, and the values of some parameters (those in \eqref{eq:s} - \eqref{eq:c}) may be very far off any values which are eventually found.
 

### 2015-09-09 MScWriteup.tex
more changes

@@
 The next step is performing the minimisation.
-In general, the landscape defined by the error function may be very rugged, and contain many local minima.
-\footnote{this is true even if ODE model is linear in its parameters, as is almost the case for us - though the ODE model is linear, the resulting solutions are in general not.
-A counterexample is the harmonic oscillator.} A local optimisation algorithm will often get stuck in local minima.
+In general, the landscape defined by the error function may be rugged and contain many local minima, which a local optimisation algorithm will may get stuck in \footnote{this is true even if ODE model is linear in its parameters, as is almost the case for us - though the ODE model is linear, the resulting solutions are in general not.
+A counterexample is the harmonic oscillator.}.
 To try and surmount this problem, \cite{Algorithms2003} suggests the use of a global optimisation algorithm, and in particular recommends several Evolutionary Algorithms, of which we choose one, the CMA-ES \cite{Hansen2006,Hansen2011}.
@@
 Figs.
-\ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices, computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
-An element of the matrix $R_{ij}$ is then the correlation co-efficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters.
+\ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices between parameter sets, computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
+An element of the correlation matrix $R_{ij}$ is then the correlation co-efficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters.
 These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
 Referring to \eqref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
-Biologically, an increased production rate of GFP being balance by an increased dilution rate.
+Biologically, an increased production rate of GFP is being balance by an increased dilution rate.
 The results suggest a fitness landscape relatively flat to perturbations in certain combinations of parameters, and indicate we may have difficulty obtaining unique estimates of the model parameters.
@@
 Though the predictions are reasonable, they are substantially worse than the predictions for the data they were trained on.
-The first reason for this may be that parameter values will vary between experimental runs \cite{Hu2015}.
-The second may be that the model is overfitting.
-\todo[inline]{follow this up properly.}
+One reason for this may be that parameter values will vary between experimental runs \cite{Hu2015}.
+Another may be that the model is inadequate.
 
- Taken together, these results suggest that some parameters may be inestimable.  
- \todo[inline]{move this down into the discussion, and that the data may be better described by a simplified model}
+ Taken together, these results suggest that simply performing a least squares fit on the available data will not give unique parameter estimates.
+In the following sections we discuss why this might be.
 
@@
 
-It can be shown that if the sensitivity coefficients (columns of $S$) are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
-Related to this fact, a number of measures have been proposed to assess parameter estimability from the sensitivity matrix \cite{Mclean2012}, the simplest of which is to plot the sensitivity coefficients as a function of time, and visually check for obvious linear relations between the curves.
+It can be shown that if the sensitivity curves (columns of $S$) are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
+Related to this fact, a number of measures have been proposed to assess parameter estimability from the sensitivity matrix \cite{Mclean2012}, the simplest of which is to plot the sensitivity curves as a function of time, and visually check for obvious linear relations between them.
 
@@
     \end{subfigure}
-    \caption{Sensitivity coefficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset.
+    \caption{Sensitivity coefficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset, for a single oscillation.
 \ref{SensitivityMatrix_unscaled} shows them unscaled, \ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$.}
@@
         \includegraphics[scale = 0.26,clip = true, trim = 90 0 60 0]{model_output}
-        \caption{model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
+        \caption{Model output for all state variables over two oscillations, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
 Note $g$ is not explicitly shown, but is simply a rescaling of $z$ and as such would lie over it.
@@
 This would explain the similarity between the sensitivity curves of many parameters - if the parameters in \eqref{eq:s} - \eqref{eq:c} can only act to alter the model's fixed point, they would all alter the sensitivity curve in exactly the same way, since a small perturbation in one of these parameters would only act to perturb the location of the models fixed points.
-It would then not be surprising if there were large spaces of parameters, all giving the same model fixed point, which were indistinguishable.
+It would also suggest that all parameter sets giving the same value of the models fixed point will give similar model predictions - and there may be many, very different, parameter sets which all give the same fixed point.
 
@@
 Explicit forms for $m$ and $c$ are given in \eqref{eq:m_fixed}, \eqref{eq:c_fixed}.
-The fact that this term is  the only one in \eqref{eq:p}, \eqref{eq:g} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in \eqref{eq:s} - \eqref{eq:c} - all that $p$ and $g$ see is this $\beta m +f_{s}\beta c$ flipping between its two fixed point values, so any parameters that only enter the system via this term will have indistinguishable roles.
+The fact that this term is  the only one in \eqref{eq:p}, \eqref{eq:g} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in \eqref{eq:s} - \eqref{eq:c} (Fig.
+\ref{SensitivityMatrix_scaled}) - all that $p$ and $g$ see is this $\beta m +f_{s}\beta c$ flipping between its two fixed point values, so any parameters that only enter the system via this term will have indistinguishable roles.
 
-Although we have seen that parameter sets with similar least squares error values can contain very different parameter values, this explanation implies that we would expect similar values of $\beta m +f_{s}\beta c$ across all the sets.Fig.
-\ref{fixedpoint_f} shows a scatterplot of model fixed point values against error function value, and demonstrates that, though individual parameter values can be spread across very large ranges, they are correlated in such a way as to give similar model fixed points.
+Although we have seen that parameter sets with similar least squares error values can contain very different parameter values, this explanation implies that we would expect similar values of $\beta m +f_{s}\beta c$ across all the sets.
+Fig.
+\ref{fixedpoint_f} shows a scatterplot of $\beta m +f_{s}\beta c$  values against error function value, and demonstrates that, though individual parameter values can be spread across very large ranges, they are correlated in such a way as to give similar model fixed points.
 
@@
 
-In this report, we have proposed a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}.
+In this report, we have presented a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}.
 We have attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach, similar to that found in recent system biology literature \cite{Hu2015}.
@@
 The results in this paper suggest that, even though the model has a solid biological rationale, some parameters included in it may not be estimable.
-We have suggested practical reasons for this, namely that the step function nature of the forcing is giving highly correlated parameter effects, causing the minimisation algorithm to effectively work in a much lower dimensional space than our full parameter space.
-This result suggests that using the additional step function data  available may not help in determining the parameters - all of these step functions are applied at timescales over which the `fast' part of the model responds instantly, and so essentially tell us the same thing about the parameters contained in those equations.
+We have suggested practical reasons for this, namely that the effects of the parameters are highly correlated with one another, because the relevant quantity for determining a model prediction is not the full parameter set, but only the fixed point values of $\beta m +f_{s}\beta c$, and $\mu$.
+Thus the parameters in \eqref{eq:s} - \eqref{eq:c} only work to alter the fixed point value of $\beta m +f_{s}\beta c$, and so all have a qualitatively identical effects on model predictions.
+Effectively our algorithm is only working in a 2 dimensional space, varying $\beta m +f_{s}\beta c$ and $\mu$ to minimize error.
+This result suggests that using any additional step function data available may not help in determining the parameters - all of these step functions are applied at timescales over which the `fast' part of the model responds instantly, and so essentially tell us the same thing about the parameters contained in those equations.
 The different sets may give more information about $\mu$, which appears to work on a timescale comparable to the forcing.
@@
  Thus, further work may consist of tightening the bounds on parameter values, carrying out a structural analysis, and possibly simplifying the model.
-A much simpler model, consisting of  just \eqref{eq:p} and \eqref{eq:g}, may explain existing data perfectly well without the need for inestimable parameters.
+A much simpler phenomenological model, consisting of  just \eqref{eq:p} and \eqref{eq:g}, may explain existing data perfectly well without the need for inestimable parameters.
 Experimentally, one option  may be to use different forcing functions.
 The effect of this on parameter estimability might be anticipated by running a sensitivity analysis using this new forcing, to see if the form of Fig.
-\ref{SensitivityMatrix} changes.
+\ref{SensitivityMatrix} (which shows model predictions for all states) changes.
 Another option, though perhaps experimentally infeasible, would be to directly observe other components of the state vector - $s$, $m$, $s:m$ and $c$.
-This would improve parameter estimates by giving direct observations of the effects of parameters contained in eqs.
-\ref{eq:s} - \ref{eq:c}.
+This would improve parameter estimates by giving direct observations of the effects of parameters contained in \eqref{eq:s} - \eqref{eq:c}.
 
@@
 To surmount this problem, we may rerun the algorithm many times, from many starting locations, and interpret the resulting spread of parameter values found as an uncertainty \cite{Hu2015}, as we have done in this report.
-However this methodology is rather ad hoc, and a more systematic approach which gives a probability distribution on parameter space would be desirable.
+However this methodology is rather ad hoc.
+
+A more systematic approach might be to use Markov Chain Monte Carlo (MCMC) for the parameter estimation \cite{Jitjareonchai2006,Andrieu2003}.
+This method would explicitly provide us with marginal distributions of parameters, giving a more complete and systematic picture of the best parameter sets and their uncertainties than we currently have - we would still be able to pick out a Maximum Likelihood point estimator, as we do now, but rather than getting an ad hoc picture of the uncertainty in the parameters by repeatedly running the minimisation algorithm, the marginals would give us this information directly.
 
@@
 
- \begin{itemize}
- \item look at degeneracies in the fixed points.
- \item discuss evidence for model being very over parametrized, and suggest a simpler model may work equally well - suggest biologically interesting parameters may be unobservable.
- \item discuss general methodological flaws - throwing marbles in the air, local minima, is MLE even any good?
-say data shows even the EA doesn't avoid local minima.
-The space is so unconstrained.
-the algorithm can't be expected to just fix all these problems.
- \item suggest additional forcing time data may not help due to two timing nature, 
- \item suggest there MAY be other ways to fit the data, but the wide range of parameter bounds mean we just don't know - the landscape is complicated and huge, and there may be minima widely separated in the parameter space.
-whose to say which is right?
- \item Further work - on experimnetal side, propoer lit review to rightly constrain some values, fix more.
-More state variables!
-should easily be able to fix mu, theta, maybe others too?
-other forcing functions?
- \item a proper methodology - gaussian processes.
- \item mention this is how sys bio papers do it, and we were given some earlier fitting work which we think is now wrong.
- \item suggest a general structural analysis be carried out.
-
- \end{itemize}
-
-  \begin{itemize}
- \item explore degeneracies in fixed point equations
- \item discuss estimatbility criteria- uniqeuess and sensitivity
- \item put in some sensitivity analysis results
- \item discuss model fixed points
- \item discuss systematic paramater shift.
- \item discuss cross - validation and predicitive ability of parameter sets found.
- \item suggest two times, work out fixed point for all the solutions found, look at structure of soln,
- \item re run from starting pararms.
- \end{itemize}
-
 \section{Model Fixed Point}

### 2015-09-10 MScWriteup.tex
adding cross validation stuff

@@
 Appendix \ref{Initial  Experimental Data} shows the full datasets, with their forcing functions.
+Note that \textit{13\_9} contains two different forcing periods.
 
@@
 \min_{\boldsymbol{\theta}} \sum_{i =1}^{n} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2.
+\label{eq:min}
 \end{align}
@@
 However, histograms of estimated parameter values found from fitting the above 200 sets indicate many parameter estimates are not tightly constrained, taking values right across the initial bounding range specified - in particular, $k_\mathrm{off}$, $\delta_{m}$ and $\delta_{s}$ are very weakly constrained, with substantial numbers of runs constrained only by the initial bounding box.
-By contrast, some parameter estimates ($\mu$ and $\Theta$ in particular) are much tighter.
+By contrast, some parameter estimates ($\mu$ in particular) are much tighter.
+$\Theta$ appears tightly constrained, but many of the estimated values are close to the upper bound set, and must be treated with caution.
 
@@
 
- We can test the predictive ability of our model by cross validating - taking the parameter values found in the fitting of one dataset, and using them to give model predictions for another.
-We thus take the parameter values giving the best fit for the  \textit{13\_9} dataset and use them to predict the \textit{14\_7} dataset, and vice versa.
-Results are shown in Fig.\ref{CrossValidation}.
+ We can test the predictive ability of our model by cross validating, either by taking the parameter values found in the fitting of one dataset and using them to give model predictions for another, or only fitting part of a single dataset and predicting the rest.
+We begin by fitting to only the data for the first forcing period in \textit{13\_9}, and then predicting the full time series.
+Results are shown in Fig.
+\ref{CrossValidation_1}, for the parameter set giving the lowest error on the training data.
+We see the prediction is very similar to that obtained by fitting to the full dataset (Fig.
+\ref{InitialResults_13_9_best}).
+Next, we take the parameter values giving the best fit for the  \textit{13\_9} dataset and use them to predict the \textit{14\_7} dataset, and vice versa.
+Results are shown in Figs.
+\ref{CrossValidation_2}, \ref{CrossValidation_3}.
+Though the predictions are reasonable, they are substantially worse than the predictions for the data they were trained on.
+One reason for this may be that parameter values will vary between experimental runs \cite{Hu2015}.
+We can also fit to both datasets combined, by extending \eqref{eq:min} to a sum over multiple datasets.
+Results are shown in Appendix \ref{Bothdatasetfitting}.
+We are unable to achieve error values as low as those found when fitting individual datasets.
+In addition, we do not see a tightening of the estimates on our parameters, as we might expect from including the additional data.
+Instead, many parameter estimates remain spread across the initial bounding region.
+We also note that all three of the cases considered (\textit{13\_9}, \textit{14\_7}, both) see a shift in the estimated value of $\mu$.
 
- \begin{figure}[h!]	
+ \begin{figure}[h]	
+ 	\begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.3,clip = true, trim = 87 0 70 0]{13_9_partial_data_13_9_pred}
+        \caption{Model trained on \textit{13\_9}  60/60 data only, full prediction}
+        \label{CrossValidation_1} 
+    \end{subfigure}
 	\begin{subfigure}[h]{0.49\textwidth}
@@
         \caption{\textit{14\_7} prediction \textit{13\_9} data}
-        \label{SensitivityMatrix_unscaled} 
+        \label{CrossValidation_2} 
     \end{subfigure}
@@
         \caption{\textit{13\_9} prediction \textit{14\_7} data}
-                \label{SensitivityMatrix_scaled} 
+                \label{CrossValidation_3} 
     \end{subfigure}
     \caption{Cross validating data by taking parameter estimates from one dataset, and using them to predict another.
-Here we take the best parameter set fitted on the \textit{13\_9}, and fit it against the \textit{14\_7} dataset, and vice versa }
+\ref{CrossValidation_1} shows model predictions for the full \textit{13\_9} dataset, when only trained on the first forcing period data.
+\ref{CrossValidation_2}, \ref{CrossValidation_3} show model predictions for fitting to one dataset, and predicting the other.}
 \label{CrossValidation}   
@@
 
-Though the predictions are reasonable, they are substantially worse than the predictions for the data they were trained on.
-One reason for this may be that parameter values will vary between experimental runs \cite{Hu2015}.
-Another may be that the model is inadequate.
-
  Taken together, these results suggest that simply performing a least squares fit on the available data will not give unique parameter estimates.
-In the following sections we discuss why this might be.
+In the following sections we investigate why this might be.
 
@@
 % references section
+\begin{footnotesize}
 \bibliographystyle{./IEEE_bibtex_class/IEEEtran}
 \bibliography{./IEEE_bibtex_class/IEEEabrv,library}
-
+\end{footnotesize}
 \clearpage
@@
 \centering
-\includegraphics[trim = 0 310 0 300 , scale = 0.5, clip = true]{13_9}
+\includegraphics[trim = 0 310 0 300 , scale = 0.7, clip = true]{13_9}
 \caption{The \textit{13\_9} dataset, with aTc forcing shown.
@@
 \centering
-\includegraphics[trim = 0 270 0 360 , scale = 0.5, clip = true]{14_7}
+\includegraphics[trim = 0 270 0 250 , scale = 0.7, clip = true]{14_7}
 \caption{The \textit{14\_7} dataset, with aTc forcing shown.
@@
 %\end{figure}
-
-\section{Histogram of Initial Fit Error values}
-\label{Initial fit Error}
-\setcounter{figure}{0}
-
-\begin{figure}[H]
-    \begin{subfigure}[c]{\textwidth}
-    \centering
-        \includegraphics[scale = 0.5]{13_9_f_hist}
-        \caption{Error values found from 200 initial parameter estimates, \textit{13\_9 dataset}.}
-    \end{subfigure}
-
-    \begin{subfigure}[c]{\textwidth}
-    \centering
-        \includegraphics[scale = 0.5]{14_7_f_hist}
-        \caption{Error values found from 200 initial parameter estimates, \textit{14\_7 dataset}}
-    \end{subfigure}
-    \caption{}
-\label{hist_f}
-\end{figure}
 
@@
 \end{table}
+
+\section{Histogram of Initial Fit Error values}
+\label{Initial fit Error}
+\setcounter{figure}{0}
+
+\begin{figure}[H]
+    \begin{subfigure}[c]{\textwidth}
+    \centering
+        \includegraphics[scale = 0.5]{13_9_f_hist}
+        \caption{Error values found from 200 initial parameter estimates, \textit{13\_9 dataset}.}
+    \end{subfigure}
+
+    \begin{subfigure}[c]{\textwidth}
+    \centering
+        \includegraphics[scale = 0.5]{14_7_f_hist}
+        \caption{Error values found from 200 initial parameter estimates, \textit{14\_7 dataset}}
+    \end{subfigure}
+    \caption{}
+\label{hist_f}
+\end{figure}
+
+\section{Initial parameter Estimates, both datasets}
+\label{Bothdatasetfitting}
+\setcounter{figure}{0}
+
+\begin{figure}[H]
+    \begin{subfigure}[c]{\textwidth}
+    \centering
+        \includegraphics[scale = 0.3]{13_9_14_7_with_13_9_bestplot}
+        \caption{Model prediction from parameter set which gives lowest error when trained on the combination of both datasets, plotted against the  \textit{13\_9} dataset. }
+    \end{subfigure}
+
+    \begin{subfigure}[c]{\textwidth}
+    \centering
+        \includegraphics[scale = 0.3]{13_9_14_7_with_14_7_bestplot}
+        \caption{Model prediction from parameter set which gives lowest error when trained on the combination of both datasets, plotted against the  \textit{14\_7} dataset.}
+    \end{subfigure}
+    \caption{}
+\label{hist_f}
+\end{figure}
+
+\begin{figure}[H]
+   \centering
+        \includegraphics[scale = 0.4]{13_9_14_7_hist}
+        \caption{Histogram of estimated parameter values, found from 100 runs of the CMA-ES algorithm. Fitted to both the  \textit{13\_9} and \textit{14\_7} datasets.}
+        \label{}
+\end{figure}
+
+\begin{figure}[H]
+   \centering
+        \includegraphics[scale = 0.3]{13_9_14_7_f_hist}
+        \caption{Error values found from 100 initial parameter estimates. Fitted to both the  \textit{13\_9} and \textit{14\_7} datasets.}
+        \label{}
+\end{figure}
+
+\end{document}
 
@@
 \end{figure*}
-
-\end{document}

### 2015-09-10 MScWriteup.tex
adding appendix

@@
         \centering
-    \includegraphics[scale = 0.25, clip = true, trim = 100 0 90 0]{13_9_heatmap.eps}
+    \includegraphics[scale = 0.24, clip = true, trim = 120 0 70 0]{13_9_heatmap.eps}
         \caption{Correlation matrix of estimated parameter sets, from the  \textit{13\_9} dataset. }
@@
     \centering
-        \includegraphics[scale = 0.25, clip = true, trim = 80 0 70 0]{14_7_heatmap.eps}
+        \includegraphics[scale = 0.24, clip = true, trim = 80 0 70 0]{14_7_heatmap.eps}
         \caption{Correlation matrix of estimated parameter sets, from the  \textit{14\_7} dataset.}
@@
     \centering
-        \includegraphics[scale = 0.3,clip = true, trim = 87 0 70 0]{13_9_partial_data_13_9_pred}
+        \includegraphics[scale = 0.29,clip = true, trim = 82 0 70 0]{13_9_partial_data_13_9_pred}
         \caption{Model trained on \textit{13\_9}  60/60 data only, full prediction}
@@
     \centering
-        \includegraphics[scale = 0.35,clip = true, trim = 30 0 60 0]{Sensitivity_unscaled}
+        \includegraphics[scale = 0.23,clip = true, trim = 110 0 60 0]{Sensitivity_unscaled}
         \caption{Unscaled}
@@
     \centering
-        \includegraphics[scale = 0.35,clip = true, trim = 15 0 60 0]{Sensitivity_scaled}
+        \includegraphics[scale = 0.23,clip = true, trim = 70 0 60 0]{Sensitivity_scaled}
         \caption{Scaled}
@@
 
+\clearpage
+
 \section{Parameter literature review}
@@
 
-\begin{table}[H]
+\begin{table}[h!]
 \renewcommand{\arraystretch}{1.3}
@@
 \end{table}
+
+\clearpage
 

### 2015-09-12 MScWriteup.tex
rewriting final section

@@
 \newcounter{MYtempeqncnt}
+\renewcommand{\thefootnote}{\roman{footnote}}
 
@@
 \begin{abstract}
-For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA is an attractive tool, due to the ease of predicting its behaviour from physiochemical models.
-In this report, we introduce a system of ODE's to model a recently designed synthetic regulatory circuit, before attempting to estimate its unknown parameters using recent time series data.
+For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA is an attractive tool due to the ease of predicting its behaviour from physiochemical models.
+In this report, we introduce a system of ODEs to model a recently designed synthetic regulatory circuit, before attempting to estimate its unknown parameters using recent single cell time series data.
 We provide estimates for a subset of parameters, but find many of them to be inestimable.
-We discuss why this is this case for our system, and suggest how this problem might be resolved.
+We discuss why this is the case for our system, and suggest how this problem might be resolved.
 \end{abstract}
@@
 For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription.
-In comparison to proteins, it is relatively straightforward to predict the structure and function of an RNA from its sequence using physiochemical models.
-Recently, this has been exploited to computationally design synthetic sRNA's - small RNA's which do not code for a protein, but rather have some direct regulatory function -  with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.
+In comparison to proteins, it is relatively straightforward to predict the intra and intermolecular interactions of an RNA from its sequence using physiochemical models.
+Recently, this has been exploited to computationally design synthetic sRNAs - small RNAs which do not code for a protein, but rather have some direct regulatory function -  with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.
 
 This report will focus on one such sRNA system, introduced in \cite{Rodrigo2012}.
-It will extend existing understanding of it beyond the qualitative by first proposing a quantitative model of its behaviour in the form of a set of ODE's, and then fitting this ODE model to available time series data to estimate its unknown parameters.
+It will extend existing understanding of it beyond the qualitative by first proposing a quantitative model of its behaviour in the form of a set of ODEs, and then fitting this ODE model to available time series data to estimate its unknown parameters.
 
@@
 In the remainder of this section we review the regulatory system we will consider, and discuss recent single cell fluorescence experiments performed on it.
-In section \ref{ODEModel} we introduce a set of ODE's to model the system.
+In section \ref{ODEModel} we introduce a set of ODEs to model the system.
 In section \ref{Parameter Estimation} we attempt to estimate its unknown parameters by fitting the model to time series data.
@@
 \includegraphics[trim = 60 630 300 30,page=2,clip = true]{pnas1203831109.pdf}
-\caption{A proposed reaction scheme between the synthetic sRNA-mRNA pairs designed in \cite{Rodrigo2012}.
+\caption{The reaction scheme between the synthetic sRNA-mRNA pairs designed in \cite{Rodrigo2012}.
 The reaction co-ordinate is defined as the number of paired nucleotides, and the vertical axis denotes free energy.
@@
 coli}.
-Further, by placing the \textit{in vivo}  concentrations of the sRNA and mRNA under the control of tuneable promoter regions \footnote{A promoter region in a DNA sequence is a sequence of nucleotides, found upstream of where transcription of a gene begins, which can influence the transcription rate of the gene.}, they constructed a logical AND gate from one of the proposed pairs (Fig.
+Further, by placing the \textit{in vivo}  concentrations of the sRNA and mRNA under the control of tuneable promoters \footnote{A promoter in a DNA sequence is a sequence of nucleotides, found upstream of where transcription of a gene begins, which can influence the transcription rate of the gene.}, they constructed a logical AND gate from one of the proposed pairs (Fig.
 \ref{ANDGate}).
 
-In this system, transcription of the designed sRNA and mRNA are placed under the control of promoter regions, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}.
-These are in turn controlled by two transcriptional repressors, TetR and LacI, which are naturally present in the strain of \textit{E.
-coli} considered.
+In this system, transcription of the designed sRNA and mRNA are placed under the control of promoters, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}.
+These are in turn controlled by two transcriptional repressors, TetR and LacI respectively, which are naturally present in the strain of \textit{E.
+coli} used.
 These repressors disable the promoter regions, and so by default transcription of both RNA's is turned off, and no protein is produced.
@@
 \ref{ANDGate}).
-So transcription of the two RNA's is indirectly controlled by the presence of two chemicals - if neither is present, sRNA and mRNA transcription is repressed, and no protein is produced.
+So transcription of the two RNAs is indirectly controlled by the presence of two chemicals - if neither is present, sRNA and mRNA transcription is repressed, and no protein is produced.
 If only one is present, the AND gate remains off, either because there is no mRNA to be translated into protein, or because the mRNA is self repressed.
@@
 
-The data we will consider consists of two sets of individual cell time series, labelled \textit{13\_9} and \textit{14\_7}.
+The data we will consider consists of two sets of individual cell time series, labelled \textit{13\_9} and \textit{14\_7} \cite{Jaramilloa}.
 They correspond to different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response, and aTc concentrations were varied periodically.
@@
 
-In \eqref{eq:s} - \eqref{eq:z}, we present a modified version an existing model which describes the system, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}.
+In \eqref{eq:s} - \eqref{eq:z}, we present a modified version an existing model which describes the system \cite{Jaramilloa}, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}.
 Its state is given by the vector $(s,m,s:m,c,p,g,z)$, with all other variables representing model parameters.
 Tables \ref{StateVariables} and \ref{ModelParameters} give complete descriptions of the parameters and state variables.
-\todo[inline]{give the original model and discuss modifications in appendix.}
 
@@
 Secondly, in the experimental setup we can arrange for GFP molecules to be produced with a \textit{degradation tag} attached to them \cite{Hersch2004}.
-This tag is sought out by a protein, ClpX, which will then degrade the molecule the tag is  attached to.
+This tag is identified by an enzyme, ClpX, which will then degrade the molecule the tag is  attached to.
 This degradation process is modelled by the final terms in \eqref{eq:p} and \eqref{eq:g}.
@@
 Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response.
-The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response.}
+The forcing curve's height is schematic - aTc concentration is switched between off, and on a level which saturates the cell's response.}
 \label{InitialResults}
@@
 
- We begin by fitting each dataset individually, by choosing 200 parameter sets uniformly distributed over our initial parameter bounds and running the CMA-ES starting from them.
+ We begin by fitting each dataset individually, by choosing 200 sets of the parameters, with each parameter choice uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
 Results are shown in Figs.
@@
 \ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices between parameter sets, computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
-An element of the correlation matrix $R_{ij}$ is then the correlation co-efficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters.
+An element of the correlation matrix $R_{ij}$ is then the Pearson correlation coefficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters.
 These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
 Referring to \eqref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
-Biologically, an increased production rate of GFP is being balance by an increased dilution rate.
 The results suggest a fitness landscape relatively flat to perturbations in certain combinations of parameters, and indicate we may have difficulty obtaining unique estimates of the model parameters.
@@
 
- \begin{figure*}[!t]
-% ensure that we have normalsize text
-\tiny
-% Store the current equation number.
-\setcounter{MYtempeqncnt}{\value{equation}}
-% Set the equation number to one less than the one
-% desired for the first equation here.
-% The value here will have to changed if equations
-% are added or removed prior to the place these
-% equations are referenced in the main text.
-\setcounter{equation}{11}
-\begin{align}
-\label{eq:m_fixed}  
-m &= \frac{1}{{2 k_{\text{on}} \left(\mu +\delta _m\right)
-   \left(k_{\text{hyb}}+\delta _m\right)}}
-   \Bigg(\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
-   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
-   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2} \\ \nonumber  
- &- (\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)\Bigg)
-\end{align}
-\begin{align}
-\label{eq:c_fixed}
-c &= \frac{k_{\text{hyb}}}{2 k_{\text{on}} \left(\mu +\delta
-   _m\right) \left(k_{\text{hyb}}+\delta _m\right){}^2}
-   \Bigg( \sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
-   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
-   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2} \\ \nonumber
-   &+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right) \Bigg)
-\end{align}
-% Restore the current equation number.
-\setcounter{equation}{\value{MYtempeqncnt}}
-% IEEE uses as a separator
-\hrulefill
-% The spacer can be tweaked to stop underfull vboxes.
-\vspace*{4pt}
-\end{figure*}
-
  \subsection{Parameter Estimability}
@@
 Fig.
-\ref{SensitivityMatrix} shows plots of each column of $S$, evaluated at the parameter set giving the lowest error in the \textit{13\_9} dataset.
+\ref{SensitivityMatrix} shows plots of each column of $S$, evaluated at the parameter set giving the lowest error in the \textit{13\_9} dataset (Fig \ref{InitialResults_13_9_best}).
 Fig.
@@
 This result suggests that, at least in some of the parameter sets initially estimated, the system may have two timescales in it - a fast timescale in which \eqref{eq:s} - \eqref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in \eqref{eq:p}, \eqref{eq:g}, in which measured fluorescence changes in response to the forcing.
+Biologically, this would correspond to \eqref{eq:p}, representing translation, being a rate limiting step.
 
@@
     \centering
-        \includegraphics[scale = 0.26,clip = true, trim = 90 0 60 0]{model_output}
+        \includegraphics[scale = 0.25,clip = true, trim = 95 0 60 0]{model_output}
         \caption{Model output for all state variables over two oscillations, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
@@
 
-If this is the case, then it may be that our experimental data can only probe the system via the fixed point of \eqref{eq:s} - \eqref{eq:c} , and that the parameters contained within those equations can only work to set the particular fixed point values of $p$ and $g$ that the system tends toward.
-This would explain the similarity between the sensitivity curves of many parameters - if the parameters in \eqref{eq:s} - \eqref{eq:c} can only act to alter the model's fixed point, they would all alter the sensitivity curve in exactly the same way, since a small perturbation in one of these parameters would only act to perturb the location of the models fixed points.
-It would also suggest that all parameter sets giving the same value of the models fixed point will give similar model predictions - and there may be many, very different, parameter sets which all give the same fixed point.
+If this is the case, then our experimental data can only probe the system via the fixed point of \eqref{eq:s} - \eqref{eq:c}, whose value is communicated to the rest of the model via the $\beta m +f_{s}\beta c$ term in \eqref{eq:p} \footnote{ Explicit forms for $m$ and $c$ are given in appendix \ref{ModelFixedPoint}}.
+$\beta m +f_{s}\beta c$ will flip between two values, which will cause the fixed point values of $p$ and $g$ to flip, with $p$ and $g$ tending towards them in response.
+This is what we observe in Fig \ref{model_output}.
 
-The fixed point values of $s,m, s:m$ and $c$ affect \eqref{eq:p}, \eqref{eq:g} through the $\beta m +f_{s}\beta c$ term in \eqref{eq:p}.
-Explicit forms for $m$ and $c$ are given in \eqref{eq:m_fixed}, \eqref{eq:c_fixed}.
-The fact that this term is  the only one in \eqref{eq:p}, \eqref{eq:g} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in \eqref{eq:s} - \eqref{eq:c} (Fig.
-\ref{SensitivityMatrix_scaled}) - all that $p$ and $g$ see is this $\beta m +f_{s}\beta c$ flipping between its two fixed point values, so any parameters that only enter the system via this term will have indistinguishable roles.
+This would explain the similarity between the sensitivity curves of many parameters - if the parameters in \eqref{eq:s} - \eqref{eq:c} can only act to alter the fixed point values of $p$ and $g$ which the system tends toward, they all play qualitatively identical roles on the time series of $z$.
+The fact that $\beta m +f_{s}\beta c$ is the only term in \eqref{eq:p}, \eqref{eq:g} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in \eqref{eq:s} - \eqref{eq:c} (Fig.
+\ref{SensitivityMatrix_scaled}), by the same reasoning.
+This explanation also suggests that all parameter sets giving the same value of the model's fixed point will give similar model predictions - and there may be many, very different, parameter sets which all give the same fixed point.
+Fig.
+\ref{fixedpoint_f} shows a scatterplot of error function value against $\beta m +f_{s}\beta c$ for the 200 estimated parameter sets for the \textit{13\_9} dataset, and demonstrates that, although individual parameter values can be spread across very large ranges, they are correlated in such a way as to give similar model fixed points.
 
-Although we have seen that parameter sets with similar least squares error values can contain very different parameter values, this explanation implies that we would expect similar values of $\beta m +f_{s}\beta c$ across all the sets.
-Fig.
-\ref{fixedpoint_f} shows a scatterplot of $\beta m +f_{s}\beta c$  values against error function value, and demonstrates that, though individual parameter values can be spread across very large ranges, they are correlated in such a way as to give similar model fixed points.
+These results suggest we may simplify our model, by replacing $\beta m +f_{s}\beta c$ in \eqref{eq:p} with a phenomenological forcing function, and removing \eqref{eq:s} - \eqref{eq:c} entirely.
+The simplified model is shown in \eqref{eq:p_simplified} - \eqref{eq:z_simplified}
 
-In Fig.
-\ref{fixedpoint_mu}, we also see a strong positive correlation between the fixed point values and the values of $\mu$ ($R = 0.9854$ for the visually tightly clustered data).
-Taken together with the relatively tightly constrained values of $\mu$ found in Fig.
-\ref{InitialResults}, these results suggest that in order to minimize error, the algorithm is effectively trying to vary $\beta m +f_{s}\beta c$ and $\mu$, and finding a `trench' of highly correlated values, which is shown in Fig.
-\ref{fixedpoint_mu}.
+\begin{align}
+\frac{dp}{dt} & = Fy(t)-(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}  \label{eq:p_simplified} \\
+\frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \label{eq:g_simplified} \\
+z &= z_{0} +\frac{g}{\Theta} \label{eq:z_simplified}
+\end{align}
 
-These results further suggest that, while we are trying to minimise model error over our initial high dimensional space, we are effectively working in a lower dimensional space, in which one dimension is the value of the models $\beta m +f_{s}\beta c$ .
-
-\begin{figure}[h]	
-    \centering
-        \includegraphics[scale = 0.25]{fixedpoint_f}
-        \caption{A scatterplot of the fixed point value of $\beta m +f_{s}\beta c$, for a given parameter set, against the error value of that set. }
-        \label{fixedpoint_f} 
-\end{figure}
-
-\begin{figure}[h]	
-    \centering
-        \includegraphics[scale = 0.25]{fixedpoint_mu}
-        \caption{A scatterplot of the fixed point value of $\beta m +f_{s}\beta c$, for a given parameter set, against the value of $\mu$ in that set.}
-        \label{fixedpoint_mu} 
-\end{figure}
+where $F$ is a phenomenological scaling factor.
+We now have three free parameters, $F,\mu,\Theta$, which we may estimate as before.
+Results are shown in Fig.\ref{SimplifiedModelResults}.
+In contrast to the results of Fig.\ref{InitialResults}, we see relatively tightly estimated parameters, with no estimated parameter values near the initial bounding box.
+Further, this simplified model achieves error values as low as the more complex model, and predictions from both are similar.
+In fact, the simplified algorithm is more consistently able to find the minimal function value identified in Fig.
+\ref{Initial fit Error}, possibly because the low dimension of the new model makes parameter space easier to search.
+We also note positive correlation between $F$ and $\Theta$, suggesting that both play similar roles in determining the fit.
+This may be explained as follows: neglecting the ClpX degradation term, \eqref{eq:p_simplified} is a simple exponential decay to a fixed value.
+Here, $F$ will determine the value the $p$ will tend to, and $\mu$ will determine its rate constant.
+$F$ is then a linear scaling factor, just as $\Theta$ is - both scale the value the exponential will tend to.
 
@@
 In this report, we have presented a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}.
-We have attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach, similar to that found in recent system biology literature \cite{Hu2015}.
-Where this approach has failed to give estimates, we have discussed in detail how the method has broken down.
+We have attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach, similar to that found in recent synthetic biology literature \cite{Hu2015}.
+Where this approach has failed to give estimates, we have discussed why, and suggested that a simpler model may provide an equally good description of the data with fewer parameters than the initial model.
 
-The results in this paper suggest that, even though the model has a solid biological rationale, some parameters included in it may not be estimable.
-We have suggested practical reasons for this, namely that the effects of the parameters are highly correlated with one another, because the relevant quantity for determining a model prediction is not the full parameter set, but only the fixed point values of $\beta m +f_{s}\beta c$, and $\mu$.
-Thus the parameters in \eqref{eq:s} - \eqref{eq:c} only work to alter the fixed point value of $\beta m +f_{s}\beta c$, and so all have a qualitatively identical effects on model predictions.
-Effectively our algorithm is only working in a 2 dimensional space, varying $\beta m +f_{s}\beta c$ and $\mu$ to minimize error.
-This result suggests that using any additional step function data available may not help in determining the parameters - all of these step functions are applied at timescales over which the `fast' part of the model responds instantly, and so essentially tell us the same thing about the parameters contained in those equations.
-The different sets may give more information about $\mu$, which appears to work on a timescale comparable to the forcing.
+The results in this paper demonstrate that, even though the original model has a solid biological rationale, some parameters included in it may not be estimable.
+The suggested reason for this is that \eqref{eq:p} represents a rate limiting step, and so the parameters in \eqref{eq:s} - \eqref{eq:c} only work to alter the fixed point value of $\beta m +f_{s}\beta c$ in \eqref{eq:p}, causing them to give qualitatively identical effects on model predictions.
+A simplified model, consisting of three free parameters, gives model predictions with errors as low as the full model, but contains parameters which can be estimated unambiguously.
 
-We note that it is possible that this separation of timescales is an artefact of the fitting, and not necessarily true of the system.
-We can only evaluate model predictions on the parameter sets the algorithm returns us, and it seems the case that the parameter sets found by the algorithm have this feature.
-Hypothetically, there may exist parameter sets where all state variables vary on a similar timescale - all we can say is that, in the search region specified, the algorithm has not found any which give a low error.
+These results suggest that, if estimates of the parameters contained in \eqref{eq:s} - \eqref{eq:c} are desired, fluorescence time series data alone will not be sufficient.
+One option, though perhaps experimentally infeasible, would be to directly observe other components of the state vector - $s$, $m$, $s:m$ and $c$.
+This would improve parameter estimates by giving direct observations of the `fast' timescale.
+On the other hand, if a good description of the fluorescence data is what is really of interest, further work may consist of using the simplified model, but improving the modelling of the translation step.
+\cite{Hu2015}, for example, uses a more detailed model than our single step description of translation.
+We also note that $\mu$ and $\Theta$ are both parameters which can be easily estimated independent of our datasets - it may be of interest to compare those values to the ones estimated here.
 
-This point raises another issue - the arbitrary nature of the initial bounding box.
-Without good biological constraints on most parameter values, they have been bounded loosely and not necessarily sensibly.
-Rate constants can vary by many orders of magnitude, and it may be possible that the true optimum lies entirely outside the bounding box specified.
-Equally,  when our fitness function has turned out to be very flat in some directions, or multimodal, good biologically motivated parameter bounds may be needed to ensure our minimisation algorithm does not wander outside realistic parameter space - the algorithm may present several equally good regions of parameter space, and improved bounds will help us choose between them.
-We also note that while some parameters may be hard to get any realistic bounds on, values for others - $\mu, \Theta$ for example - may already be available.
-
- The model may also have general structural issues.
-\cite{Mclean2012, Grewal1976} give general tests which can be applied to a system of ODE's to determine if all parameters are, in principle, estimable.
-We note that our system has few enough parameters to make these tests feasible.
-
- Thus, further work may consist of tightening the bounds on parameter values, carrying out a structural analysis, and possibly simplifying the model.
-A much simpler phenomenological model, consisting of  just \eqref{eq:p} and \eqref{eq:g}, may explain existing data perfectly well without the need for inestimable parameters.
-Experimentally, one option  may be to use different forcing functions.
-The effect of this on parameter estimability might be anticipated by running a sensitivity analysis using this new forcing, to see if the form of Fig.
-\ref{SensitivityMatrix} (which shows model predictions for all states) changes.
-Another option, though perhaps experimentally infeasible, would be to directly observe other components of the state vector - $s$, $m$, $s:m$ and $c$.
-This would improve parameter estimates by giving direct observations of the effects of parameters contained in \eqref{eq:s} - \eqref{eq:c}.
-
-Some earlier fitting work has been done on this dataset via a similar method to the one used above.
-Our results suggest that those estimates should be viewed with caution - the estimates will not be unique, and the values of some parameters (those in \eqref{eq:s} - \eqref{eq:c}) may be very far off any values which are eventually found.
-
-These results also suggest general problems with our methodology - using a least squares minimisation approach.
-The first problem is local minima.
-Using the CMA-ES will give better results than a local minimisation algorithm, but you can never be sure you have found a global optimum - for example, \cite{Algorithms2003} compares several algorithms and finds, in the problem set, none locate the true global optimum.
-A related problem is that of working with single points rather than distributions on parameter space.
+Further work may also consist of improving our methodology.
+Using a least squares minimisation approach has several problems, the first of which is local minima.
+Using the CMA-ES will give better results than a local minimisation algorithm, but one can never be sure that global optimum has been found - for example, \cite{Algorithms2003} compares several algorithms and finds, in the test case considered, none locate the true global optimum.
+A related problem is that of working with point estimates of parameters rather than distributions.
 The CMA-ES will only ever find a single optimum, but in the case where the error landscape is very flat -  for example a flat basin pocketed by many local minima - the exact minimum it finds paints a misleading picture, because it gives no idea of the uncertainty in the answer given.
@@
 
-A more systematic approach might be to use Markov Chain Monte Carlo (MCMC) for the parameter estimation \cite{Jitjareonchai2006,Andrieu2003}.
-This method would explicitly provide us with marginal distributions of parameters, giving a more complete and systematic picture of the best parameter sets and their uncertainties than we currently have - we would still be able to pick out a Maximum Likelihood point estimator, as we do now, but rather than getting an ad hoc picture of the uncertainty in the parameters by repeatedly running the minimisation algorithm, the marginals would give us this information directly.
-
-\todo[inline]{write section on GP's}.
+A more systematic approach might be to use Markov Chain Monte Carlo (MCMC) to perform a Bayesian parameter estimation \cite{Jitjareonchai2006,Andrieu2003}.
+This method would explicitly provide us with marginal distributions of parameters, giving a more complete and systematic picture of the best parameter sets and their uncertainties than we currently have - we would still be able to pick out a maximum likelihood point estimator, as we do now, but rather than getting an ad hoc picture of the uncertainty in the parameters by repeatedly running the minimisation algorithm, the marginals would give us this information directly.
 
@@
 
+\clearpage
+
 \begin{figure}[H]
@@
 \end{figure}
-
-\end{document}
 
@@
 % equations are referenced in the main text.
-\setcounter{equation}{5}
-\begin{equation}
-\label{eqn_dbl_x}
-y\to \frac{\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+\setcounter{equation}{11}
+\begin{align}
+\label{eq:m_fixed}  
+m &= \frac{1}{{2 k_{\text{on}} \left(\mu +\delta _m\right)
+   \left(k_{\text{hyb}}+\delta _m\right)}}
+   \Bigg(\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
    \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
-   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)}{2 k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right){}^2}
-\end{equation}
-\begin{equation}
-\label{eqn_dbl_y}
-m\to -\frac{\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2} \\ \nonumber  
+ &- (\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)\Bigg)
+\end{align}
+\begin{align}
+\label{eq:c_fixed}
+c &= \frac{k_{\text{hyb}}}{2 k_{\text{on}} \left(\mu +\delta
+   _m\right) \left(k_{\text{hyb}}+\delta _m\right){}^2}
+   \Bigg( \sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
    \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
-   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}-(\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)}{2 k_{\text{on}} \left(\mu +\delta _m\right)
-   \left(k_{\text{hyb}}+\delta _m\right)}
-\end{equation}
-\begin{equation}
-\label{eqn_dbl_y}
-s\to -\frac{\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
-   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
-   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}+(\text{am}-\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)}{2 k_{\text{on}} \left(\mu +\delta _s\right)
-   \left(k_{\text{hyb}}+\delta _m\right)}
-\end{equation}
-\begin{equation}
-\frac{k_{\text{hyb}} \left(\sqrt{2 (\text{am}+\text{as}) k_{\text{on}} \left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+\delta _m\right)
-   \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)+(\text{am}-\text{as})^2 k_{\text{on}}^2 \left(k_{\text{hyb}}+\delta _m\right){}^2+\left(\mu +\delta _m\right){}^2
-   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2}+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
-   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right)\right)}{2 k_{\text{on}} \left(\mu +\delta
-   _m\right) \left(k_{\text{hyb}}+\delta _m\right){}^2}
-\end{equation}
+   \left(\mu +\delta _s\right){}^2 \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right){}^2} \\ \nonumber
+   &+(\text{am}+\text{as}) k_{\text{on}} \left(k_{\text{hyb}}+\delta
+   _m\right)+\left(\mu +\delta _m\right) \left(\mu +\delta _s\right) \left(k_{\text{hyb}}+k_{\text{off}}+\delta _m\right) \Bigg)
+\end{align}
 % Restore the current equation number.
@@
 \end{figure*}
+
+\section{Fixed point scatterplot}
+\label{Fixedpointscatterplot}
+\setcounter{figure}{0}
+
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.25]{fixedpoint_f}
+        \caption{A scatterplot of the fixed point value of $\beta m +f_{s}\beta c$, for a given parameter set, against the error value of that set. }
+        \label{fixedpoint_f} 
+\end{figure}
+
+\end{document}
+
+We note that it is possible that this separation of timescales is an artefact of the fitting, and not necessarily true of the system.
+We can only evaluate model predictions on the parameter sets the algorithm returns us, and it seems the case that the parameter sets found by the algorithm have this feature.
+Hypothetically, there may exist parameter sets where all state variables vary on a similar timescale - all we can say is that, in the search region specified, the algorithm has not found any which give a low error.
+
+We also note that while some parameters may be hard to get any realistic bounds on, values for others - $\mu, \Theta$ for example - may already be available.
+
+This point raises another issue - the arbitrary nature of the initial bounding box.
+Without good biological constraints on most parameter values, they have been bounded loosely and not necessarily sensibly.
+Rate constants can vary by many orders of magnitude, and it may be possible that the true optimum lies entirely outside the bounding box specified.
+Equally,  when our fitness function has turned out to be very flat in some directions, or multimodal, good biologically motivated parameter bounds may be needed to ensure our minimisation algorithm does not wander outside realistic parameter space - the algorithm may present several equally good regions of parameter space, and improved bounds will help us choose between them.
+
+ The model may also have general structural issues.
+\cite{Mclean2012, Grewal1976} give general tests which can be applied to a system of ODEs to determine if all parameters are in principle estimable, and apply them successfully to models with comparable numbers of parameters to ours.
+
+ Thus, further work may consist of tightening the bounds on parameter values, carrying out a structural analysis, and possibly simplifying the model.
+A much simpler phenomenological model, consisting of  just \eqref{eq:p} and \eqref{eq:g}, may explain existing data perfectly well without the need for inestimable parameters.
+Experimentally, one option  may be to use different forcing functions.
+The effect of this on parameter estimability might be anticipated by running a sensitivity analysis using this new forcing, to see if the form of Fig.
+\ref{SensitivityMatrix} (which shows model predictions for all states) changes.
+Another option, though perhaps experimentally infeasible, would be to directly observe other components of the state vector - $s$, $m$, $s:m$ and $c$.
+This would improve parameter estimates by giving direct observations of the effects of parameters contained in \eqref{eq:s} - \eqref{eq:c}.

### 2015-09-13 MScWriteup.tex
still rewriting last section

@@
 The next step is performing the minimisation.
-In general, the landscape defined by the error function may be rugged and contain many local minima, which a local optimisation algorithm will may get stuck in \footnote{this is true even if ODE model is linear in its parameters, as is almost the case for us - though the ODE model is linear, the resulting solutions are in general not.
-A counterexample is the harmonic oscillator.}.
+In general, the landscape defined by the error function may be rugged and contain many local minima, which a local optimisation algorithm will may get stuck in.
 To try and surmount this problem, \cite{Algorithms2003} suggests the use of a global optimisation algorithm, and in particular recommends several Evolutionary Algorithms, of which we choose one, the CMA-ES \cite{Hansen2006,Hansen2011}.
-\footnote{CMA-ES stands for covariance matrix evolutionary strategy.
-This algorithm generates a population of test points according to a multivariate gaussian distribution.
-It then ranks the points based on their fitness function scores, and  considers some number of the highest ranked points as a sample from a new multivariate gaussian.
-Using this sample, it then estimates a new mean and covariance matrix for this gaussian and the process is iterated.}
 \todo[inline]{why use CMA-ES? Some refs}
@@
 Figs.
-\ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices between parameter sets, computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
-An element of the correlation matrix $R_{ij}$ is then the Pearson correlation coefficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters.
+\ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices between parameter sets\footnote{These are computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
+An element of the correlation matrix $R_{ij}$ is then the Pearson correlation coefficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters}.
 These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
@@
 
- There are two main reasons why a parameter may not be estimable \cite{Mclean2012,Yao2003, Beck}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others.
+ There are two main reasons why a parameter may not be estimable \cite{Mclean2012,Yao2003, Beck,Jimenez-Hornero2008,Grewal1976}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others.
 
@@
 For example, it is possible that in the experimental regime we operate in, parameter effects may be weak, or highly correlated, but in other regimes this is not true.
-\footnote{\cite{Beck} gives an example in which a parameter only affects model predictions after several hours, though others will affect it at all timescales - in this example, if we only took data for a few minutes, the parameter would be inestimable, but in principle it is not.} In this case, parameter estimates may be improved by taking data in more varied experimental conditions, and attempting to observe as many components of the model output as possible.
+In this case, parameter estimates may be improved by taking data in more varied experimental conditions, or attempting to increase the number of model outputs observed - in our case, observing more of the state vector than simply $z$.
 
@@
 
+\subsection{A simplified model}
+
 These results suggest we may simplify our model, by replacing $\beta m +f_{s}\beta c$ in \eqref{eq:p} with a phenomenological forcing function, and removing \eqref{eq:s} - \eqref{eq:c} entirely.
@@
 We now have three free parameters, $F,\mu,\Theta$, which we may estimate as before.
-Results are shown in Fig.\ref{SimplifiedModelResults}.
-In contrast to the results of Fig.\ref{InitialResults}, we see relatively tightly estimated parameters, with no estimated parameter values near the initial bounding box.
+Results are shown in Fig.\ref{SimplifiedModelResults}, for the \textit{13\_9} dataset only.
+In contrast to the results of Fig.
+\ref{InitialResults}, we see relatively tightly estimated parameters, with no estimated parameter values near the initial bounding box \footnote{Initial bounding box values given in appendix \ref{Simplified model initial bounding box}.}.
 Further, this simplified model achieves error values as low as the more complex model, and predictions from both are similar.
@@
 We also note positive correlation between $F$ and $\Theta$, suggesting that both play similar roles in determining the fit.
+Fig.
+\ref{Sensitivity_scaled_simplified} shows the normalised sensitivity curves using this simplified model for the \textit{13\_9} dataset, and also suggests $F$ and $\Theta$ play similar, though not identical, roles.
+
   This may be explained as follows: neglecting the ClpX degradation term, \eqref{eq:p_simplified} is a simple exponential decay to a fixed value.
@@
 $F$ is then a linear scaling factor, just as $\Theta$ is - both scale the value the exponential will tend to.
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale = 0.23,clip = true, trim = 90 0 60 0]{Sensitivity_scaled_simplified}
+\caption{Using the simplified model \eqref{eq:p_simplified} - \eqref{eq:z_simplified}, normalised sensitivity coefficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset, for a single oscillation.}
+\label{Sensitivity_scaled_simplified}
+\end{figure}
 
@@
 
+\clearpage
+
+\section{Simplified model initial bounding box}
+\label{Simplified model initial bounding box}
+\setcounter{figure}{0}
+
+\begin{table}[h!]
+\renewcommand{\arraystretch}{1.3}
+\caption{Initial bounds on parameters to be estimated in the simplified model, \eqref{eq:p_simplified} - \eqref{eq:z_simplified}.}
+\centering
+\begin{tabular}{| l | l | p{0.3\linewidth} | p{0.1\linewidth} | p{0.1\linewidth} |}
+\hline \textbf{Parameter} &  \textbf{Units} & \textbf{Definition}  & \textbf{Initial Bounds}  \\
+\hline ${\mu}$ &  /min & Dilution rate  & 0.001-0.1\\
+\hline ${F}$ &  nM /min & Phenomenological forcing  & 10-1500\\
+\hline ${\Theta}$  &   nM/AFU & Ratio between GFP concentration and observed fluoresence  & 100 - 10000  \\
+\hline
+\end{tabular}
+\end{table}
+
 \end{document}

### 2015-09-13 MScWriteup.tex
rewriting latter sections

@@
 \begin{figure*}
+    \begin{subfigure}[c]{0.49\textwidth}
+    \centering
+    \includegraphics[scale = 0.26,clip = true, trim = 90 0 70 0]{13_9_bestPlot}
+        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset.}
+                \label{InitialResults_13_9_best}
+    \end{subfigure}
+    \begin{subfigure}[c]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.26,clip = true, trim = 90 0 70 0]{14_7_bestPlot}
+        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{14\_7} dataset.}
+        \label{InitialResults_14_7_best}
+    \end{subfigure}
         \begin{subfigure}[h]{0.49\textwidth}
@@
     \end{subfigure}
-
-        \begin{subfigure}[c]{0.49\textwidth}
-        \centering
-    \includegraphics[scale = 0.24, clip = true, trim = 120 0 70 0]{13_9_heatmap.eps}
-        \caption{Correlation matrix of estimated parameter sets, from the  \textit{13\_9} dataset. }
-        \label{InitialResults_13_9_correlation}
-    \end{subfigure}
-    \begin{subfigure}[c]{0.49\textwidth}
-    \centering
-        \includegraphics[scale = 0.24, clip = true, trim = 80 0 70 0]{14_7_heatmap.eps}
-        \caption{Correlation matrix of estimated parameter sets, from the  \textit{14\_7} dataset.}
-        \label{InitialResults_14_7_correlation}
-    \end{subfigure}
-
-    \begin{subfigure}[c]{0.49\textwidth}
-    \centering
-    \includegraphics[scale = 0.26,clip = true, trim = 90 0 70 0]{13_9_bestPlot}
-        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset.}
-                \label{InitialResults_13_9_best}
-    \end{subfigure}
-    \begin{subfigure}[c]{0.49\textwidth}
-    \centering
-        \includegraphics[scale = 0.26,clip = true, trim = 90 0 70 0]{14_7_bestPlot}
-        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{14\_7} dataset.}
-        \label{InitialResults_14_7_best}
-    \end{subfigure}
-    \caption{Parameter estimates, inter-parameter correlation values, and model predictions for the \textit{13\_9} and \textit{14\_7} datasets.
+    \caption{Parameter estimates and model predictions for the \textit{13\_9} and \textit{14\_7} datasets.
 Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response.
@@
 
- We begin by fitting each dataset individually, by choosing 200 sets of the parameters, with each parameter choice uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
+ We begin by fitting each dataset individually, by choosing 200 sets of the parameters uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
 Results are shown in Figs.
-\ref{InitialResults}, \ref{hist_f}.
+\ref{InitialResults}, \ref{InitialResults_additional}.
 Figs.
 \ref{InitialResults_13_9_best}, \ref{InitialResults_14_7_best}  demonstrate that the model is capable of quantitatively capturing the data.
-However, histograms of estimated parameter values found from fitting the above 200 sets indicate many parameter estimates are not tightly constrained, taking values right across the initial bounding range specified - in particular, $k_\mathrm{off}$, $\delta_{m}$ and $\delta_{s}$ are very weakly constrained, with substantial numbers of runs constrained only by the initial bounding box.
+However, histograms of the parameter estimates (Figs.
+\ref{InitialResults_13_9_hist}, \ref{InitialResults_14_7_hist}) indicate that many are not tightly constrained, taking values right across the initial bounding range specified - in particular, $k_\mathrm{off}$, $\delta_{m}$ and $\delta_{s}$ are very weakly constrained, with substantial numbers of runs constrained only by the initial bounding box.
 By contrast, some parameter estimates ($\mu$ in particular) are much tighter.
@@
 These problems may stem from structural inadequacies in the model (often termed \textit{structural identifiability}), in which two different parameter sets can give identical model predictions \cite{Jimenez-Hornero2008,Grewal1976 }.
-If this is the case, no amount of experimental data will allow us to estimate parameters, and we must consider reformulating the model.
-\footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product $\beta_{1}\beta_{2}$ can ever be estimated, a problem no amount of data can fix.}
-
+If this is the case, no amount of experimental data will allow us to estimate parameters, and we must consider reformulating the model\footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product $\beta_{1}\beta_{2}$ can ever be estimated, a problem no amount of data can fix.}.
 Problems may also arise for more practical reasons (\textit{practical identifiability} \cite{Mclean2012}).
@@
 where $F$ is a phenomenological scaling factor.
-We now have three free parameters, $F,\mu,\Theta$, which we may estimate as before.
-Results are shown in Fig.\ref{SimplifiedModelResults}, for the \textit{13\_9} dataset only.
+We now have three free parameters, $F, \mu, \Theta$, which we may estimate as before.
+Based on the results of earlier fitting, we set a new bounding box for these parameters (shown in table \ref{Simplified_model_initial_bounding_box}) and run the CMA-ES 400 times.
+Results are shown in Fig.
+\ref{SimplifiedModelResults}, for the \textit{13\_9} dataset only.
+
+\begin{figure}[h]	
+	\begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.25 ,clip = true, trim = 100 0 60 400]{13_9_hist_simplified}
+        \caption{Histograms of estimated parameter values, using the simplified model \eqref{eq:p_simplified} - \eqref{eq:z_simplified}.}
+        \label{13_9_hist_simplified} 
+    \end{subfigure}
+    \begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale =  0.25 ,clip = true, trim = 100 0 60 400]{13_9_scatterplot_simplified}
+        \caption{Scatterplots of estimated parameters shown in \ref{13_9_hist_simplified}  against one another}
+                \label{13_9_scatterplot_simplified} 
+    \end{subfigure}
+    \caption{Histograms and scatterplots of parameter estimates, found using the simplified model \eqref{eq:p_simplified} - \eqref{eq:z_simplified}.} 
+\label{SimplifiedModelResults}   
+\end{figure}
+
  In contrast to the results of Fig.
-\ref{InitialResults}, we see relatively tightly estimated parameters, with no estimated parameter values near the initial bounding box \footnote{Initial bounding box values given in appendix \ref{Simplified model initial bounding box}.}.
-Further, this simplified model achieves error values as low as the more complex model, and predictions from both are similar.
-In fact, the simplified algorithm is more consistently able to find the minimal function value identified in Fig.
-\ref{Initial fit Error}, possibly because the low dimension of the new model makes parameter space easier to search.
-We also note positive correlation between $F$ and $\Theta$, suggesting that both play similar roles in determining the fit.
+\ref{InitialResults}, we see relatively tightly estimated parameters, with no estimated parameter values near the initial bounding box.
+Further, this simplified model achieves error values as low as the more complex model.
+In fact, Fig \ref{13_9_f_simplified} shows that the simplified algorithm is more consistently able to find the minimal function value identified in Fig.
+\ref{InitialResults_13_9_f}, possibly because the low dimension of the new model makes parameter space easier to search.
+
+ We also note that \ref{13_9_scatterplot_simplified} shows two distinct clusters of parameter sets, and a strong positive correlation between between $F$ and $\Theta$.
+The latter suggests that both play similar roles in determining the fit.
 Fig.
 \ref{Sensitivity_scaled_simplified} shows the normalised sensitivity curves using this simplified model for the \textit{13\_9} dataset, and also suggests $F$ and $\Theta$ play similar, though not identical, roles.
-
   This may be explained as follows: neglecting the ClpX degradation term, \eqref{eq:p_simplified} is a simple exponential decay to a fixed value.
@@
 $F$ is then a linear scaling factor, just as $\Theta$ is - both scale the value the exponential will tend to.
+
+  The clustering of values suggests that there may be several local minima, all with similar error values.
+We further investigate by fixing the value of $\Theta$, and plotting the error as a function of $\mu$ and $F$.
 
@@
 This method would explicitly provide us with marginal distributions of parameters, giving a more complete and systematic picture of the best parameter sets and their uncertainties than we currently have - we would still be able to pick out a maximum likelihood point estimator, as we do now, but rather than getting an ad hoc picture of the uncertainty in the parameters by repeatedly running the minimisation algorithm, the marginals would give us this information directly.
+
+\begin{itemize}
+\item systematic mode comparison ? or subset of paramter estimation if we really want to keep this model.
+\end{itemize}
 
@@
 
-\section{Parameter literature review}
+\section{Parameter literature review, and rough initial Parameter bounds}
 \label{Parameter literature review}
@@
 
-\section{Histogram of Initial Fit Error values}
-\label{Initial fit Error}
+\section{Additional Initial Results figures}
+\label{Additional Initial Results figures}
 \setcounter{figure}{0}
@@
 \begin{figure}[H]
-    \begin{subfigure}[c]{\textwidth}
+    \begin{subfigure}[c]{0.49\textwidth}
         \centering
-        \includegraphics[scale = 0.5]{13_9_f_hist}
-        \caption{Error values found from 200 initial parameter estimates, \textit{13\_9 dataset}.}
+    \includegraphics[scale = 0.24, clip = true, trim = 120 0 70 0]{13_9_heatmap.eps}
+        \caption{Correlation matrix of estimated parameter sets, from the  \textit{13\_9} dataset. }
+        \label{InitialResults_13_9_correlation}
+    \end{subfigure}
+    \begin{subfigure}[c]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.24, clip = true, trim = 80 0 70 0]{14_7_heatmap.eps}
+        \caption{Correlation matrix of estimated parameter sets, from the  \textit{14\_7} dataset.}
+        \label{InitialResults_14_7_correlation}
     \end{subfigure}
 
-    \begin{subfigure}[c]{\textwidth}
+       \begin{subfigure}[c]{0.49\textwidth}
         \centering
-        \includegraphics[scale = 0.5]{14_7_f_hist}
+    \includegraphics[scale = 0.35, clip = true, trim = 0 0 0 0]{13_9_f_hist}
+        \caption{Error values found from 200 initial parameter estimates, \textit{13\_9 dataset}. }
+        \label{InitialResults_13_9_f}
+    \end{subfigure}
+    \begin{subfigure}[c]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.35, clip = true, trim = 0 0 0 0]{14_7_f_hist}
         \caption{Error values found from 200 initial parameter estimates, \textit{14\_7 dataset}}
+        \label{InitialResults_14_7_f}
     \end{subfigure}
-    \caption{}
-\label{hist_f}
+    \caption{Inter-parameter correlation values, and error function values, for the \textit{13\_9} and \textit{14\_7} datasets.}     
+\label{InitialResults_additional}
 \end{figure}
+
+\clearpage
 
@@
 
+\clearpage
+
 \section{Model Fixed Point}
@@
 
- \begin{figure*}[!t]
+ \begin{figure}[h!]
 % ensure that we have normalsize text
@@
 \vspace*{4pt}
-\end{figure*}
+\end{figure}
 
@@
 
-\section{Simplified model initial bounding box}
-\label{Simplified model initial bounding box}
+\section{Simplified model additional results}
+\label{Simplified model additional results}
 \setcounter{figure}{0}
@@
 \begin{table}[h!]
+\caption{}
+\label{Simplified_model_initial_bounding_box}
 \renewcommand{\arraystretch}{1.3}
@@
 \end{table}
+
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.25]{13_9_f_simplified}
+        \caption{}
+        \label{13_9_f_simplified} 
+\end{figure}
 

### 2015-09-14 MScWriteup.tex
further rewrites

@@
 Further, this simplified model achieves error values as low as the more complex model.
-In fact, Fig \ref{13_9_f_simplified} shows that the simplified algorithm is more consistently able to find the minimal function value identified in Fig.
+Fig \ref{13_9_f_simplified} shows that the simplified algorithm is more consistently able to find the minimal function value identified in Fig.
 \ref{InitialResults_13_9_f}, possibly because the low dimension of the new model makes parameter space easier to search.
@@
 
-In this report, we have presented a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}.
-We have attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach, similar to that found in recent synthetic biology literature \cite{Hu2015}.
-Where this approach has failed to give estimates, we have discussed why, and suggested that a simpler model may provide an equally good description of the data with fewer parameters than the initial model.
+In this report, we have presented a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}, and attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach similar to that found in recent synthetic biology literature \cite{Hu2015}.
+We have found many of the parameters contained in the full model to be inestimable, and suggested reasons  for why this might be.
+We have further suggested a simplified model, which provides an equally good description of the data with fewer parameters than the initial model.
 
@@
 
-These results suggest that, if estimates of the parameters contained in \eqref{eq:s} - \eqref{eq:c} are desired, fluorescence time series data alone will not be sufficient.
+These results suggest that, if estimates of the parameters contained in \eqref{eq:s} - \eqref{eq:c} are desired, fluorescence time series data alone may not be sufficient.
 One option, though perhaps experimentally infeasible, would be to directly observe other components of the state vector - $s$, $m$, $s:m$ and $c$.
 This would improve parameter estimates by giving direct observations of the `fast' timescale.
-On the other hand, if a good description of the fluorescence data is what is really of interest, further work may consist of using the simplified model, but improving the modelling of the translation step.
+Another option would be further investigation of the methods presented in \cite{Mclean2012,Yao2003,Mclean2012-2}.
+Given a complex chemical model based on physical principles, and limited data, these methods rank parameters in order of estimability, and can determine if a subset of parameters might be estimable.
+
+On the other hand, if a good description of the fluorescence data is what is really of interest, further work may consist of reformulating the model, perhaps beginning with the simplified model in \eqref{eq:p_simplified} - \eqref{eq:z_simplified}, but improving the modelling of the translation step.
 \cite{Hu2015}, for example, uses a more detailed model than our single step description of translation.
+When constructing this revised model, a systematic approach to model comparison,
+\todo[inline]{talk about AIC, bayesian mode comparison here? https://sciencehouse.files.wordpress.com/2013/07/gordonconf13.pdf}
 We also note that $\mu$ and $\Theta$ are both parameters which can be easily estimated independent of our datasets - it may be of interest to compare those values to the ones estimated here.
 
-Further work may also consist of improving our methodology.
+Further work may also consist of improving our methodology for parameter estimation.
 Using a least squares minimisation approach has several problems, the first of which is local minima.
@@
 \item systematic mode comparison ? or subset of paramter estimation if we really want to keep this model.
+\item literature bounds?
 \end{itemize}
@@
     \end{subfigure}
-    \caption{Inter-parameter correlation values, and error function values, for the \textit{13\_9} and \textit{14\_7} datasets.}     
+    \caption{Inter-parameter correlation values and error function values, found from 200 initial parameter estimates, for the \textit{13\_9} and \textit{14\_7} datasets.}     
 \label{InitialResults_additional}
@@
 
-\section{Initial parameter Estimates, both datasets}
+\section{Initial parameter Estimates, fit using both datasets}
 \label{Bothdatasetfitting}
@@
         \includegraphics[scale = 0.3]{13_9_14_7_with_13_9_bestplot}
-        \caption{Model prediction from parameter set which gives lowest error when trained on the combination of both datasets, plotted against the  \textit{13\_9} dataset. }
+        \caption{Model prediction from the parameter set which gives lowest error when trained on the combination of both datasets, plotted against the  \textit{13\_9} dataset. }
     \end{subfigure}
@@
         \includegraphics[scale = 0.3]{13_9_14_7_with_14_7_bestplot}
-        \caption{Model prediction from parameter set which gives lowest error when trained on the combination of both datasets, plotted against the  \textit{14\_7} dataset.}
+        \caption{Model prediction from the parameter set which gives lowest error when trained on the combination of both datasets, plotted against the  \textit{14\_7} dataset.}
     \end{subfigure}
@@
         \includegraphics[scale = 0.4]{13_9_14_7_hist}
-        \caption{Histogram of estimated parameter values, found from 100 runs of the CMA-ES algorithm. Fitted to both the  \textit{13\_9} and \textit{14\_7} datasets.}
+        \caption{Histograms of estimated parameter values, found from 100 runs of the CMA-ES algorithm. Fitted to both the  \textit{13\_9} and \textit{14\_7} datasets.}
         \label{}
@@
 % equations are referenced in the main text.
-\setcounter{equation}{11}
+\setcounter{equation}{14}
 \begin{align}

### 2015-09-14 MScWriteup.tex
adding error profiles

@@
 In general, the landscape defined by the error function may be rugged and contain many local minima, which a local optimisation algorithm will may get stuck in.
-To try and surmount this problem, \cite{Algorithms2003} suggests the use of a global optimisation algorithm, and in particular recommends several Evolutionary Algorithms, of which we choose one, the CMA-ES \cite{Hansen2006,Hansen2011}.
+To try and surmount this problem, \cite{Algorithms2003} suggests the use of a global optimisation algorithm, and in particular demonstrates that several Evolutionary Algorithms perform well on a test problem involving a set of ODEs modelling a biological system.
+We choose one of those recommended, the CMA-ES \cite{Hansen2006,Hansen2011}, which in addition.
 \todo[inline]{why use CMA-ES? Some refs}
@@
 $\Theta$ appears tightly constrained, but many of the estimated values are close to the upper bound set, and must be treated with caution.
+We also note a shift in the estimated value of $\mu$ between the two datasets.
+Since the two sets of estimates for $\mu$ are relatively sharp, this variation may be a genuine feature of the data - one explanation for it is a variation in cell growth rate between the two experimental runs \cite{Hu2015}.
 
@@
 Though the predictions are reasonable, they are substantially worse than the predictions for the data they were trained on.
-One reason for this may be that parameter values will vary between experimental runs \cite{Hu2015}.
+One reason for this may be the variation in parameter values between experiments suggested above.
 We can also fit to both datasets combined, by extending \eqref{eq:min} to a sum over multiple datasets.
@@
 Instead, many parameter estimates remain spread across the initial bounding region.
-We also note that all three of the cases considered (\textit{13\_9}, \textit{14\_7}, both) see a shift in the estimated value of $\mu$.
 
@@
 
- Taken together, these results suggest that simply performing a least squares fit on the available data will not give unique parameter estimates.
-In the following sections we investigate why this might be.
+ Taken together, these results suggest that simply performing a least squares fit on the available data will not give unique estimates for many of the parameters.
+In the following sections we investigate this further.
 
@@
 
-It can be shown that if the sensitivity curves (columns of $S$) are linearly dependent over the range of observation values, the associated parameters cannot be simultaneously estimated \cite{Beck}.
+Suppose we take a set of parameters that give a local minima in the error function, and plot the sensitivity curves (columns of $S$)  over the range of observation times.
+It can be shown that, if two curves are linearly dependent, there exists a degenerate line of minima in parameter space, and the two associated parameters cannot be simultaneously estimated \cite{Beck}.
 Related to this fact, a number of measures have been proposed to assess parameter estimability from the sensitivity matrix \cite{Mclean2012}, the simplest of which is to plot the sensitivity curves as a function of time, and visually check for obvious linear relations between them.
@@
 \ref{InitialResults}.
+Note also that the curve for $\Theta$ is very similar to those for the remaining parameters.
+This suggests that our estimated value of $\Theta$ in Fig.
+\ref{InitialResults_13_9_hist} is misleading, and likely a consequence of the initial bounding box.
 
@@
 
-These results suggest we may simplify our model, by replacing $\beta m +f_{s}\beta c$ in \eqref{eq:p} with a phenomenological forcing function, and removing \eqref{eq:s} - \eqref{eq:c} entirely.
+These results suggest we may simplify our model, by replacing $\beta m +f_{s}\beta c$ in \eqref{eq:p} with the forcing function used in \eqref{eq:s}, and removing \eqref{eq:s} - \eqref{eq:c} entirely - in other words, moving the forcing term directly to the translation step.
 The simplified model is shown in \eqref{eq:p_simplified} - \eqref{eq:z_simplified}
@@
 
-where $F$ is a phenomenological scaling factor.
+where $F$ is a phenomenological scaling factor, and $y(t)$ is the same forcing function used in \eqref{eq:s}.
 We now have three free parameters, $F, \mu, \Theta$, which we may estimate as before.
@@
  In contrast to the results of Fig.
-\ref{InitialResults}, we see relatively tightly estimated parameters, with no estimated parameter values near the initial bounding box.
+\ref{InitialResults}, we see relatively tightly estimated parameters, with no estimated values near the initial bounding box.
 Further, this simplified model achieves error values as low as the more complex model.
@@
 
- We also note that \ref{13_9_scatterplot_simplified} shows two distinct clusters of parameter sets, and a strong positive correlation between between $F$ and $\Theta$.
-The latter suggests that both play similar roles in determining the fit.
-Fig.
-\ref{Sensitivity_scaled_simplified} shows the normalised sensitivity curves using this simplified model for the \textit{13\_9} dataset, and also suggests $F$ and $\Theta$ play similar, though not identical, roles.
+ We note that \ref{13_9_scatterplot_simplified} shows a strong positive correlation between the minima found for $F$ and $\Theta$.
+This is consistent with the  normalised sensitivity curves for the \textit{13\_9} dataset, shown in Fig.
+\ref{Sensitivity_scaled_simplified}.
+As before, the similarity of the $F$ and $\Theta$ curves suggests that any minima in $F - \Theta$ space will be locally non-unique \cite{Beck}.
 This may be explained as follows: neglecting the ClpX degradation term, \eqref{eq:p_simplified} is a simple exponential decay to a fixed value.
@@
 
-  The clustering of values suggests that there may be several local minima, all with similar error values.
-We further investigate by fixing the value of $\Theta$, and plotting the error as a function of $\mu$ and $F$.
-
 \begin{figure}[h]
@@
 
+ We also note that the parameter estimates come in two distinct clusters.
+We further investigate by fixing the value of $\Theta$, and plotting the error as a function of $\mu$ and $F$.
+Results are shown in Fig.
+\ref{SimplifiedModelLikelihood}.
+We find a local basin in $ \mu - F$ space, pocketed by many local minima of similar fitness values.
+The optimisation algorithm will always terminate in one of these local minima, explaining the disconnected clusters shown in Fig.
+\ref{13_9_scatterplot_simplified}.
+Thus the tightness of each peak in Fig.
+\ref{13_9_hist_simplified} is misleading - the uncertainty in parameter estimates is better reflected by the contours of Fig.
+\ref{Likelihood_profile_fine}.
+
+ \begin{figure}[h]	
+	\begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.23 ,clip = true, trim = 85 0 0 0]{Likelihood_profile_rough}
+        \caption{ Error function profile, $\Theta = 5400$}
+        \label{Likelihood_profile_rough} 
+    \end{subfigure}
+    	\begin{subfigure}[h]{0.49\textwidth}
+    \centering
+        \includegraphics[scale = 0.23 ,clip = true, trim = 85 0 0 0]{Likelihood_profile_fine}
+        \caption{ Error function profile, $\Theta = 5400$}
+        \label{Likelihood_profile_fine} 
+    \end{subfigure}
+    \caption{} 
+\label{SimplifiedModelLikelihood}   
+\end{figure}
+
 \section{Conclusions and Further work}
@@
 In this report, we have presented a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}, and attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach similar to that found in recent synthetic biology literature \cite{Hu2015}.
-We have found many of the parameters contained in the full model to be inestimable, and suggested reasons  for why this might be.
+We have found many of the parameters contained in the full model to be inestimable, suggested reasons  for why this might be.
 We have further suggested a simplified model, which provides an equally good description of the data with fewer parameters than the initial model.
@@
 
-These results suggest that, if estimates of the parameters contained in \eqref{eq:s} - \eqref{eq:c} are desired, fluorescence time series data alone may not be sufficient.
-One option, though perhaps experimentally infeasible, would be to directly observe other components of the state vector - $s$, $m$, $s:m$ and $c$.
-This would improve parameter estimates by giving direct observations of the `fast' timescale.
+These results suggest that if estimates of the parameters contained in \eqref{eq:s} - \eqref{eq:c} are desired, fluorescence time series data alone will not be sufficient, and new experiments are needed.
+If possible, direct observations of other components of the state vector - $s$, $m$, $s:m$ and $c$ - would improve parameter estimates by giving access to the `fast' timescale.
 Another option would be further investigation of the methods presented in \cite{Mclean2012,Yao2003,Mclean2012-2}.
@@
 
-On the other hand, if a good description of the fluorescence data is what is really of interest, further work may consist of reformulating the model, perhaps beginning with the simplified model in \eqref{eq:p_simplified} - \eqref{eq:z_simplified}, but improving the modelling of the translation step.
-\cite{Hu2015}, for example, uses a more detailed model than our single step description of translation.
-When constructing this revised model, a systematic approach to model comparison,
-\todo[inline]{talk about AIC, bayesian mode comparison here? https://sciencehouse.files.wordpress.com/2013/07/gordonconf13.pdf}
-We also note that $\mu$ and $\Theta$ are both parameters which can be easily estimated independent of our datasets - it may be of interest to compare those values to the ones estimated here.
+On the other hand, if a good description of the fluorescence data is what is really of interest, the simplified model may be a better starting point.
+$\mu$ and $\Theta$ are both parameters which can be easily estimated independent of our datasets - a realistic range of values could be used to bound the error function profile shown in Fig.
+\ref{}.
+Further modelling work may focus on improving the modelling of translation, the rate limiting step - \cite{Hu2015}, for example, uses a more detailed model than our single step description of translation.
+Any additional parameters introduced will likely act on the same timescale as $\mu$, and stand a better chance of being estimable.
 
@@
 Using a least squares minimisation approach has several problems, the first of which is local minima.
-Using the CMA-ES will give better results than a local minimisation algorithm, but one can never be sure that global optimum has been found - for example, \cite{Algorithms2003} compares several algorithms and finds, in the test case considered, none locate the true global optimum.
+Using the CMA-ES will give better results than a local minimisation algorithm, but one can never be sure that a global optimum has been found - for example, \cite{Algorithms2003} compares several algorithms and finds, in the test case considered, none locate the true global optimum.
+
  A related problem is that of working with point estimates of parameters rather than distributions.
-The CMA-ES will only ever find a single optimum, but in the case where the error landscape is very flat -  for example a flat basin pocketed by many local minima - the exact minimum it finds paints a misleading picture, because it gives no idea of the uncertainty in the answer given.
+The CMA-ES will only ever find a single optimum, giving a point estimate of parameter values, with no uncertainty information.
 To surmount this problem, we may rerun the algorithm many times, from many starting locations, and interpret the resulting spread of parameter values found as an uncertainty \cite{Hu2015}, as we have done in this report.
-However this methodology is rather ad hoc.
+However this methodology is rather ad hoc, and we have seen in Figs.
+\ref{13_9_scatterplot_simplified,ggds} that it can give a misleading impression of how tightly estimated parameters are.
+
+\todo[inline]{talk about AIC, bayesian mode comparison here? https://sciencehouse.files.wordpress.com/2013/07/gordonconf13.pdf}
 
@@
 \caption{Literature references, or initial rough bounds, on parameter values, with those to be estimated shown in bold}
+\label{Parameter_literature_review_table}
 \centering
@@
 This would improve parameter estimates by giving direct observations of the effects of parameters contained in \eqref{eq:s} - \eqref{eq:c}.
+
+ , and that the values of $\mu$ and $\Theta$ shown in Fig.
+\ref{13_9_hist_simplified} differ from those found in the initial fitting, Fig.
+\ref{InitialResults_13_9_hist} - in particular, $\Theta$ has increased beyond the upper bound initially specified.
+The clustering of values suggests that there may be several disconnected local minima, all with similar error values.
+Within each of these minima, $\Theta$ and $\mu$ will show tight clustering.
+
+  The initial upper bound on $\Theta$ set in table \ref{Parameter_literature_review_table} will cut off some of these minima
+
+  This may be explained with reference to the clustering of parameter estimatees \ref{13_9_scatterplot_simplified}, which shows that
+
+Finally, we note that that \ref{13_9_scatterplot_simplified}
+
+  shows two distinct clusters of parameter sets,

### 2015-09-14 MScWriteup.tex
finished writing for now, time for smaller changes

@@
 Our minimisation problem is thus
-\todo[inline]{explain why not fit each curve individually.}
 \begin{align}
@@
         \includegraphics[scale = 0.23 ,clip = true, trim = 85 0 0 0]{Likelihood_profile_rough}
-        \caption{ Error function profile, $\Theta = 5400$}
+        \caption{ Colormap scaled }
         \label{Likelihood_profile_rough} 
@@
     \end{subfigure}
-    \caption{} 
+    \caption{Error plotted as a function of $\mu$ and $F$, $\Theta = 5400$} 
 \label{SimplifiedModelLikelihood}   
@@
 In this report, we have presented a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}, and attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach similar to that found in recent synthetic biology literature \cite{Hu2015}.
-We have found many of the parameters contained in the full model to be inestimable, suggested reasons  for why this might be.
+We have found many of the parameters contained in the full model to be inestimable, and suggested reasons for why this might be.
 We have further suggested a simplified model, which provides an equally good description of the data with fewer parameters than the initial model.
@@
 $\mu$ and $\Theta$ are both parameters which can be easily estimated independent of our datasets - a realistic range of values could be used to bound the error function profile shown in Fig.
-\ref{}.
+\ref{SimplifiedModelLikelihood}.
 Further modelling work may focus on improving the modelling of translation, the rate limiting step - \cite{Hu2015}, for example, uses a more detailed model than our single step description of translation.
@@
 
-Further work may also consist of improving our methodology for parameter estimation.
+Further work may also consist of fitting each timeseries individually, rather than fitting to the experimental mean, as described in \eqref{eq:min}.
+This would allow us to detect cell to cell variability in parameter values within each dataset, something we have already suggested possible between datasets.
+
+We may also consist of changing our methodology for parameter estimation.
 Using a least squares minimisation approach has several problems, the first of which is local minima.
 Using the CMA-ES will give better results than a local minimisation algorithm, but one can never be sure that a global optimum has been found - for example, \cite{Algorithms2003} compares several algorithms and finds, in the test case considered, none locate the true global optimum.
-
 A related problem is that of working with point estimates of parameters rather than distributions.
@@
 To surmount this problem, we may rerun the algorithm many times, from many starting locations, and interpret the resulting spread of parameter values found as an uncertainty \cite{Hu2015}, as we have done in this report.
-However this methodology is rather ad hoc, and we have seen in Figs.
-\ref{13_9_scatterplot_simplified,ggds} that it can give a misleading impression of how tightly estimated parameters are.
-
-\todo[inline]{talk about AIC, bayesian mode comparison here? https://sciencehouse.files.wordpress.com/2013/07/gordonconf13.pdf}
-
-A more systematic approach might be to use Markov Chain Monte Carlo (MCMC) to perform a Bayesian parameter estimation \cite{Jitjareonchai2006,Andrieu2003}.
-This method would explicitly provide us with marginal distributions of parameters, giving a more complete and systematic picture of the best parameter sets and their uncertainties than we currently have - we would still be able to pick out a maximum likelihood point estimator, as we do now, but rather than getting an ad hoc picture of the uncertainty in the parameters by repeatedly running the minimisation algorithm, the marginals would give us this information directly.
-
-\begin{itemize}
-\item systematic mode comparison ? or subset of paramter estimation if we really want to keep this model.
-\item literature bounds?
-\end{itemize}
+However this methodology is rather ad hoc, and we have seen in Fig.
+\ref{SimplifiedModelLikelihood} that it can give a misleading impression of how tightly estimated parameters are.
+
+Another approach might be to use Markov Chain Monte Carlo (MCMC) to perform a Bayesian parameter estimation \cite{Jitjareonchai2006,Andrieu2003}.
+This method would explicitly provide us with marginal distributions of parameter probabilities, giving a more complete picture of the best parameter sets and their uncertainties than we currently have - we would still be able to pick out a maximum likelihood point estimator, as we do now, but rather than getting an ad hoc picture of the uncertainty in the parameters by repeatedly running the minimisation algorithm, the marginals would give us this information directly.
+However, this change in methodology would not have allowed us to gain estimates of the parameters in the original model, and is not necessary for the simplified one.
+It may be useful to use in place of the least squares approach if further modelling of the translation step is pursued.
 

### 2015-09-14 MScWriteup.tex
fiddling with figures

@@
 \renewcommand{\thefootnote}{\roman{footnote}}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA.
-
+\setcounter{figure}{1}
 \begin{figure*}[t!]
@@
 \end{figure*}
+\setcounter{figure}{0}
 
@@
 \end{figure}
-
-In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the UTR), upstream of the RBS \footnote{Both DNA and RNA are directional - the backbone of the molecule is not symmetric.
+\setcounter{figure}{2}
+
+In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the 5' UTR), upstream of the RBS \footnote{Both DNA and RNA are directional - the backbone of the molecule is not symmetric.
 This gives the molecule two distinct ends, denoted 5' and 3'.
@@
     \centering
-		\includegraphics[trim = 280 0 0  85 ,page=17 ,clip = true, scale=0.4]{WCPM_Warwick-talk45-220115.pdf}
+		\includegraphics[trim = 280 0 0  85 ,page=17 ,clip = true, scale=0.45]{WCPM_Warwick-talk45-220115.pdf}
         \caption{An overview of the experimental setup which allows single cell fluorescences to be recorded over time.
-Shown are the bacterial growth chambers (labelled `Microfabricated Microchemstat device'), the microscope imaging them, and the software constructing fluorescence time series.
+Shown are the bacterial growth chambers (labelled `Microfabricated Microchemstat device'), the microscope imaging them (the surrounding apparatus - lens, condenser, Dichroic mirror, CCD etc.), and the software constructing fluorescence time series.
 Image reproduced from \cite{Jaramillo}.
@@
 \begin{align}
-J(\boldsymbol{\theta}) = \sum_{i =1}^{n} (\mathbf{y}_{\mathrm{exp}}(t_{i}) - \mathbf{y}(t_{i},\boldsymbol{\theta}))^2,
+ \sum_{i =1}^{n} (\mathbf{y}_{\mathrm{exp}}(t_{i}) - \mathbf{y}(t_{i},\boldsymbol{\theta}))^2,
 \end{align}
@@
 \begin{align}
-\min_{\boldsymbol{\theta}} \sum_{i =1}^{n} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2.
+\argmin_{\boldsymbol{\theta}} \sum_{i =1}^{n} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2.
 \label{eq:min}
@@
 The next step is performing the minimisation.
-In general, the landscape defined by the error function may be rugged and contain many local minima, which a local optimisation algorithm will may get stuck in.
+In general, the landscape defined by the error function may be rugged and contain many local minima, which a local optimisation algorithm may get stuck in.
 To try and surmount this problem, \cite{Algorithms2003} suggests the use of a global optimisation algorithm, and in particular demonstrates that several Evolutionary Algorithms perform well on a test problem involving a set of ODEs modelling a biological system.
@@
     \centering
-    \includegraphics[scale = 0.26,clip = true, trim = 90 0 70 0]{13_9_bestPlot}
+    \includegraphics[scale = 0.24,clip = true, trim = 115 0 60 0]{13_9_bestPlot}
         \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset.}
@@
     \centering
-        \includegraphics[scale = 0.26,clip = true, trim = 90 0 70 0]{14_7_bestPlot}
+        \includegraphics[scale = 0.24,clip = true, trim = 100 0 70 0]{14_7_bestPlot}
         \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{14\_7} dataset.}
@@
 Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response.
-The forcing curve's height is schematic - aTc concentration is switched between off, and on a level which saturates the cell's response.}
+The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response.}
 \label{InitialResults}
@@
 
- \begin{figure}[h]	
+ \begin{figure}[h!]	
  	\begin{subfigure}[h]{0.49\textwidth}
     \centering
-        \includegraphics[scale = 0.29,clip = true, trim = 82 0 70 0]{13_9_partial_data_13_9_pred}
+        \includegraphics[scale = 0.3,clip = true, trim = 70 0 70 0]{13_9_partial_data_13_9_pred}
         \caption{Model trained on \textit{13\_9}  60/60 data only, full prediction}
@@
     \centering
-        \includegraphics[scale = 0.3,clip = true, trim = 87 0 70 0]{14_7_pred_13_9_data}
+        \includegraphics[scale = 0.3,clip = true, trim = 70 0 70 0]{14_7_pred_13_9_data}
         \caption{\textit{14\_7} prediction \textit{13\_9} data}
@@
     \centering
-        \includegraphics[scale = 0.3,clip = true, trim = 87 0 70 0]{13_9_pred_14_7_data}
+        \includegraphics[scale = 0.3,clip = true, trim = 70 0 70 0]{13_9_pred_14_7_data}
         \caption{\textit{13\_9} prediction \textit{14\_7} data}
@@
 Problems may also arise for more practical reasons (\textit{practical identifiability} \cite{Mclean2012}).
-For example, it is possible that in the experimental regime we operate in, parameter effects may be weak, or highly correlated, but in other regimes this is not true.
+For example, it is possible that in the experimental regime we operate in (eg.
+the forcing function we use, the components of the model's state we can observe), parameter effects may be weak, or highly correlated, but in other regimes this is not true.
 In this case, parameter estimates may be improved by taking data in more varied experimental conditions, or attempting to increase the number of model outputs observed - in our case, observing more of the state vector than simply $z$.
@@
 S_{ij} = \hat{\theta}_{j} \frac{\partial z}{ \partial \theta_{j}}\Bigr|_{t_{i}},
+\label{eq:Sensitivity}
 \end{align}
@@
     \end{subfigure}
-    \caption{Sensitivity coefficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset, for a single oscillation.
+    \caption{Sensitivity coefficients $S_{ij}$ \eqref{eq:Sensitivity}, evaluated about a set of estimated parameters from the \textit{13\_9} dataset, for a single oscillation.
 \ref{SensitivityMatrix_unscaled} shows them unscaled, \ref{SensitivityMatrix_scaled} shows them scaled by the norm of each column of $S$.}
@@
     \centering
-        \includegraphics[scale = 0.25,clip = true, trim = 95 0 60 0]{model_output}
+        \includegraphics[scale = 0.23,clip = true, trim = 95 0 60 0]{model_output}
         \caption{Model output for all state variables over two oscillations, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
@@
 These results suggest we may simplify our model, by replacing $\beta m +f_{s}\beta c$ in \eqref{eq:p} with the forcing function used in \eqref{eq:s}, and removing \eqref{eq:s} - \eqref{eq:c} entirely - in other words, moving the forcing term directly to the translation step.
-The simplified model is shown in \eqref{eq:p_simplified} - \eqref{eq:z_simplified}
+The simplified model is shown in \eqref{eq:p_simplified} - \eqref{eq:z_simplified}:
 
@@
 
-\begin{figure}[h]	
+\begin{figure}[h!]	
 	\begin{subfigure}[h]{0.49\textwidth}
     \centering
-        \includegraphics[scale = 0.25 ,clip = true, trim = 100 0 60 400]{13_9_hist_simplified}
+        \includegraphics[scale = 0.24 ,clip = true, trim = 100 0 60 400]{13_9_hist_simplified}
         \caption{Histograms of estimated parameter values, using the simplified model \eqref{eq:p_simplified} - \eqref{eq:z_simplified}.}
@@
     \centering
-        \includegraphics[scale =  0.25 ,clip = true, trim = 100 0 60 400]{13_9_scatterplot_simplified}
+        \includegraphics[scale =  0.24 ,clip = true, trim = 100 0 60 400]{13_9_scatterplot_simplified}
         \caption{Scatterplots of estimated parameters shown in \ref{13_9_hist_simplified}  against one another}
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 
- \begin{figure}[h]	
+ \begin{figure}[t!]	
 	\begin{subfigure}[h]{0.49\textwidth}
     \centering
-        \includegraphics[scale = 0.23 ,clip = true, trim = 85 0 0 0]{Likelihood_profile_rough}
-        \caption{ Colormap scaled }
+        \includegraphics[scale = 0.24 ,clip = true, trim = 55 0 0 0]{Likelihood_profile_rough}
+        \caption{ A basin of low error values in $\mu - F$ space. Note the colour scaling maps all values above $20$ to the maximum colour.}
         \label{Likelihood_profile_rough} 
@@
     \centering
-        \includegraphics[scale = 0.23 ,clip = true, trim = 85 0 0 0]{Likelihood_profile_fine}
-        \caption{ Error function profile, $\Theta = 5400$}
+        \includegraphics[scale = 0.24 ,clip = true, trim = 55 0 0 0]{Likelihood_profile_fine}
+        \caption{Zoomed in on the bottom of the basin in Fig.\ref{Likelihood_profile_rough}. Note the colour scaling maps all values above $5.7$ to the maximum colour.}
         \label{Likelihood_profile_fine} 
     \end{subfigure}
-    \caption{Error plotted as a function of $\mu$ and $F$, $\Theta = 5400$} 
+    \caption{Error function plotted as a function of $\mu$ and $F$, $\Theta = 5400$. Note the many disconnected local minima in the bottom of a basin of low error function values.} 
 \label{SimplifiedModelLikelihood}   
@@
 On the other hand, if a good description of the fluorescence data is what is really of interest, the simplified model may be a better starting point.
-$\mu$ and $\Theta$ are both parameters which can be easily estimated independent of our datasets - a realistic range of values could be used to bound the error function profile shown in Fig.
+On the experimental side, $\mu$ and $\Theta$ are both parameters which can be easily estimated independent of our datasets - a realistic range of values could be used to bound the error function profile shown in Fig.
 \ref{SimplifiedModelLikelihood}.
@@
 
-We may also consist of changing our methodology for parameter estimation.
+We may also consider changing our methodology for parameter estimation.
 Using a least squares minimisation approach has several problems, the first of which is local minima.
@@
         \centering
-    \includegraphics[scale = 0.24, clip = true, trim = 120 0 70 0]{13_9_heatmap.eps}
+    \includegraphics[scale = 0.235, clip = true, trim = 120 0 65 0]{13_9_heatmap.eps}
         \caption{Correlation matrix of estimated parameter sets, from the  \textit{13\_9} dataset. }
@@
     \centering
-        \includegraphics[scale = 0.24, clip = true, trim = 80 0 70 0]{14_7_heatmap.eps}
+        \includegraphics[scale = 0.235, clip = true, trim = 80 0 70 0]{14_7_heatmap.eps}
         \caption{Correlation matrix of estimated parameter sets, from the  \textit{14\_7} dataset.}

### 2015-09-15 MScWriteup.tex
finalising the report now

@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+
+\usepackage{letltxmacro}
+\LetLtxMacro{\originaleqref}{\eqref}
+\renewcommand{\eqref}{eq.~\originaleqref}
 
@@
 \caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS.
-In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$.
+In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoters, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$.
 These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E.
@@
 coli} used.
-These repressors disable the promoter regions, and so by default transcription of both RNA's is turned off, and no protein is produced.
+These repressors disable the promoters, and so by default transcription of both RNA's is turned off, and no protein is produced.
 These repressors can themselves be disabled by the presence of two chemicals, aTc and IPTG, which can be introduced externally into the cell (Fig.
@@
 Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response.
-The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response.}
+The forcing curve's height is schematic - aTc concentration is switched between off and on (where by `on' we mean a level which saturates the cell's response).}
 \label{InitialResults}
@@
 \centering
-\includegraphics[scale = 0.23,clip = true, trim = 90 0 60 0]{Sensitivity_scaled_simplified}
+\includegraphics[scale = 0.23,clip = true, trim = 80 0 60 0]{Sensitivity_scaled_simplified}
 \caption{Using the simplified model \eqref{eq:p_simplified} - \eqref{eq:z_simplified}, normalised sensitivity coefficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset, for a single oscillation.}
@@
     \end{subfigure}
-    \caption{Error function plotted as a function of $\mu$ and $F$, $\Theta = 5400$. Note the many disconnected local minima in the bottom of a basin of low error function values.} 
+    \caption{Error function plotted as a function of $\mu$ and $F$, $\Theta = 5400$ (the modal value of $\Theta$ in Fig.
+\ref{13_9_hist_simplified}).
+Note the many disconnected local minima in the bottom of a basin of low error function values.}
 \label{SimplifiedModelLikelihood}   

### 2015-09-15 MScWriteup.tex
im going to change the correlation section to dsicuss problems

@@
 This report will focus on one such sRNA system, introduced in \cite{Rodrigo2012}.
-It will extend existing understanding of it beyond the qualitative by first proposing a quantitative model of its behaviour in the form of a set of ODEs, and then fitting this ODE model to available time series data to estimate its unknown parameters.
+It will extend existing understanding of it beyond the qualitative by first introducing a quantitative model of its behaviour in the form of a set of ODEs, and then fitting this ODE model to available time series data to estimate its unknown parameters.
 
@@
 The authors then experimentally validated their methodology by testing the suggested sRNA-mRNA pairs in \textit{E.
-coli}.
+coli} bacteria.
 Further, by placing the \textit{in vivo}  concentrations of the sRNA and mRNA under the control of tuneable promoters \footnote{A promoter in a DNA sequence is a sequence of nucleotides, found upstream of where transcription of a gene begins, which can influence the transcription rate of the gene.}, they constructed a logical AND gate from one of the proposed pairs (Fig.
@@
 
-In this system, transcription of the designed sRNA and mRNA are placed under the control of promoters, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}.
+In this system, transcription of the DNA sequences which produce the designed sRNA and mRNA are placed under the control of promoters, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}.
 These are in turn controlled by two transcriptional repressors, TetR and LacI respectively, which are naturally present in the strain of \textit{E.
@@
 \ref{ANDGate}).
-So transcription of the two RNAs is indirectly controlled by the presence of two chemicals - if neither is present, sRNA and mRNA transcription is repressed, and no protein is produced.
+So transcription of the two RNAs is indirectly controlled by the presence of aTc and IPTG - if neither is present, sRNA and mRNA transcription is repressed, and no protein is produced.
 If only one is present, the AND gate remains off, either because there is no mRNA to be translated into protein, or because the mRNA is self repressed.
-But when both are present, the conformational change discussed above occurs, and protein is produced.
+But when both are present, both sRNA and mRNA are produced, the conformational change in the mRNA discussed above occurs, and protein is produced.
 
 Although a qualitative understanding of this system exists \cite{Rodrigo2012}, it is of interest to attempt a quantitative understanding of the genetic circuit involved.
-Such an understanding would allow, for example, tailoring of the system in reponse to design requirements, by altering the values of the important parameters of the model.
-By changing which sRNA-mRNA pair is used in the system, it would also allow exploration of the relationship between the thermodynamic properties of each device, and the model's rate constants.
+Such an understanding would allow, for example, tailoring of the system in response to design requirements, by altering the values of the important parameters of the model.
+By changing which sRNA-mRNA pair is used in the system, it would also allow exploration of the relationship between the thermodynamic properties of each device, and any rate constants in the proposed model.
 
@@
 mRNA concentrations in the system shown in Fig.
-\ref{ANDGate} can be indirectly observed by designing the mRNA to code for GFP \footnote{ The methodology of \cite{Rodrigo2012} only optimises the 5' UTR of the mRNA, so the actual protein being coded for is unimportant.}.
-Recent experiments have used timelapse microscopy to observe the fluoresence of bacteria which contain the above sRNA-mRNA pair, as they are periodically forced with a varying aTc or IPTG concentration \cite{Jaramillo}.
+\ref{ANDGate} can be indirectly observed by designing the mRNA to code for green fluorescent protein (GFP) \footnote{ The methodology of \cite{Rodrigo2012} only optimises the 5' UTR of the mRNA, so the actual protein being coded for is unimportant.}.
+Recent experiments have used timelapse microscopy to observe the fluoresence of \textit{E.
+coli} bacteria which contain the above sRNA-mRNA pair, as they are periodically forced with a varying aTc or IPTG concentration \cite{Jaramillo}.
 
@@
     \centering
-		\includegraphics[trim = 280 0 0  85 ,page=17 ,clip = true, scale=0.45]{WCPM_Warwick-talk45-220115.pdf}
+		\includegraphics[trim = 280 10 0  85 ,page=17 ,clip = true, scale=0.45]{WCPM_Warwick-talk45-220115.pdf}
         \caption{An overview of the experimental setup which allows single cell fluorescences to be recorded over time.
@@
             \label{ExperimentalSetup_overview}
+            \vspace{-5mm}
 \end{figure}
@@
     \label{ExperimentalSetup_chambers}
+    \vspace{-5mm}
 \end{figure}
@@
 The data we will consider consists of two sets of individual cell time series, labelled \textit{13\_9} and \textit{14\_7} \cite{Jaramilloa}.
-They correspond to different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response, and aTc concentrations were varied periodically.
+They correspond to different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response (i.e the IPTG input to the AND gate is always $1$), and aTc concentrations were varied periodically.
 Appendix \ref{Initial  Experimental Data} shows the full datasets, with their forcing functions.
@@
  Control of the system by aTc is modelled by $y(t)$ in \eqref{eq:s}.
-This function models the response of the sRNA transcription rate to a time varying aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}.
+This function models the response of the sRNA transcription rate to (time varying) aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}.
 $\alpha_{T}$ is the maximal transcription rate of the $P_{\mathrm{LtetO-1}}$ promoter, and $N$ is the copy number, which models the fact that when engineering the system, many copies of the $P_{\mathrm{LtetO-1}}$ promoter may be placed in the bacterial DNA.
-Thus the transcription rate varies as a sigmoid bounded by $N\frac{1}{f_{T}}$ and $N\frac{\alpha_{T}}{f}$.
+Thus the transcription rate varies as a sigmoid bounded by $N\frac{1}{f_{T}}$ and $N\frac{\alpha_{T}}{f}$ in response to aTc concentration.
 Identical considerations hold for $x(t)$ in \eqref{eq:m}, and IPTG concentration.
@@
 This degradation process is modelled by the final terms in \eqref{eq:p} and \eqref{eq:g}.
-Finally, \eqref{eq:z} simply represents calibration of mature GFP levels to experimentally observed fluorescence, assuming a linear response.
+Finally, \eqref{eq:z}  represents calibration of mature GFP levels to experimentally observed fluorescence, assuming a linear response.
 
@@
 \hline $g$ &  nm & Mature GFP concentration  \\
-\hline $z$ & AFU & Observed fluorescence  \\
+\hline $z$ & Arbitrary (AFU) & Observed fluorescence  \\
 \hline $y(t)$ & & Unitless aTc forcing function  \\
@@
 \hline \hline N & & Number of copies of promoter existing on plasmid DNA  \\
-\hline $z_{0}$ &  Arbitrary (AU) & Baseline experimental fluorescence  \\
+\hline $z_{0}$ &   Arbitrary (AFU) & Baseline experimental fluorescence  \\
 \hline $\alpha_{L}$ & nM/min & Maximal transcription rate of $P_\mathrm{LlacO-1}$ promoter\\
@@
 
- We begin by fitting each dataset individually, by choosing 200 sets of the parameters uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
+ We begin by fitting the two datasets individually, by choosing 200 sets of the parameters uniformly distributed over our initial parameter bounds, and running the CMA-ES starting from them.
 Results are shown in Figs.
@@
 An element of the correlation matrix $R_{ij}$ is then the Pearson correlation coefficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters}.
-These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists of positive correlation between values of $\mu$ and $\beta$.
+These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists a positive correlation between values of $\mu$ and $\beta$.
 Referring to \eqref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
@@
 \end{align}
-where $S_{ij}$ is the derivative of the observed fluorescence, $z$, with parameter $\theta_{j}$, evaluated at timepoint $t_{i}$.
+where $S_{ij}$ is the derivative of the observed fluorescence, $z(t)$, with parameter $\theta_{j}$, evaluated at timepoint $t_{i}$.
 Each column of $S$ is a time series of sensitivity co-efficients, which describe how sensitive $z$ is to perturbations in the parameter associated with that column, and at what times it is most sensitive.

### 2015-09-16 MScWriteup.tex
final day!

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Computational design of RNA-based oscillatory circuits}
+\title{Modelling RNA-based oscillatory circuits}
 
@@
 We provide estimates for a subset of parameters, but find many of them to be inestimable.
-We discuss why this is the case for our system, and suggest how this problem might be resolved.
+We suggest this is a consequence of translation being a rate limiting step in our ODE model and suggest that a simplified model, describing translation alone, may describe the data equally well.
 \end{abstract}
@@
 For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription.
-In comparison to proteins, it is relatively straightforward to predict the intra and intermolecular interactions of an RNA from its sequence using physiochemical models.
+In comparison to proteins, it is relatively straightforward to predict the intra and intermolecular interactions of RNA from its sequence using physiochemical models.
 Recently, this has been exploited to computationally design synthetic sRNAs - small RNAs which do not code for a protein, but rather have some direct regulatory function -  with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.
@@
 \ref{RBS}).
-This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA.
+This occurs at the Ribosome Binding Site (RBS) \cite{MolecularBiology}, a specific nucleotide sequence found on the mRNA.
 \setcounter{figure}{1}
@@
 This gives the molecule two distinct ends, denoted 5' and 3'.
-Translation can only occur in the 5' to 3' direction, hence the meaning of `upstream'}.
+Translation can only occur in the 5' to 3' direction.}.
 Translation may be self repressed by this `tail' folding over and binding across the RBS, forming a  loop in the mRNA and preventing the Ribosome from binding (Fig.
@@
 
-In \eqref{eq:s} - \eqref{eq:z}, we present a modified version an existing model which describes the system \cite{Jaramilloa}, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}.
+In \eqref{eq:s} - \eqref{eq:z}, we present a modified version an existing model which describes the system \cite{Jaramilloa}, consisting of a set of ODE's with mass action kinetics.
 Its state is given by the vector $(s,m,s:m,c,p,g,z)$, with all other variables representing model parameters.
@@
 The next step is performing the minimisation.
-In general, the landscape defined by the error function may be rugged and contain many local minima, which a local optimisation algorithm may get stuck in.
+In general, the landscape defined by the error function may be rugged and contain many local minima, which a simple local optimisation algorithm may get stuck in.
 To try and surmount this problem, \cite{Algorithms2003} suggests the use of a global optimisation algorithm, and in particular demonstrates that several Evolutionary Algorithms perform well on a test problem involving a set of ODEs modelling a biological system.
-We choose one of those recommended, the CMA-ES \cite{Hansen2006,Hansen2011}, which in addition.
-\todo[inline]{why use CMA-ES? Some refs}
+We choose one of those recommended, the CMA-ES \cite{Hansen2006,Hansen2011} which has been successfully used for parameter estimation in ODE models \cite{Ballesta2011,Dulong2015} .
 
@@
     \centering
-        \includegraphics[scale = 0.24, clip = true, trim = 140 0 125 0]{13_9_hist}
+        \includegraphics[scale = 0.24, clip = true, trim = 130 0 125 0]{13_9_hist}
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset.}
@@
     \centering
-        \includegraphics[scale = 0.24,clip = true, trim = 140 0 125 0]{14_7_hist}
+        \includegraphics[scale = 0.24,clip = true, trim = 130 0 125 0]{14_7_hist}
         \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{14\_7} dataset. }
@@
 Since the two sets of estimates for $\mu$ are relatively sharp, this variation may be a genuine feature of the data - one explanation for it is a variation in cell growth rate between the two experimental runs \cite{Hu2015}.
-
-Figs.
-\ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices between parameter sets\footnote{These are computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
-An element of the correlation matrix $R_{ij}$ is then the Pearson correlation coefficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters}.
-These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists a positive correlation between values of $\mu$ and $\beta$.
-Referring to \eqref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.
-The results suggest a fitness landscape relatively flat to perturbations in certain combinations of parameters, and indicate we may have difficulty obtaining unique estimates of the model parameters.
 
@@
 
+Some aspects of these initial fitting results are worth comment.
+For example, Fig.
+\ref{InitialResults_13_9_f} shows that many of the estimated parameter sets in Fig.
+\ref{InitialResults_13_9_hist} do not give the minimum error function value found - instead, the majority find a local minimum slightly above it.
+In fact, if we examine the estimated parameter sets giving the lowest error values, several have $\mu = 0.05$, the upper bound set.
+Thus, in the tightly clustered $\mu$ histogram, we must bear in mind that $\mu \approx 0.044$ is not a unique optimum - the error landscape may indeed be pocketed by many local minima, all giving a similar quality of fit.
+
+Taken together, these results suggest the following:
+
+\begin{enumerate}
+\item Simply performing a least squares fit on the available data will not give tightly constrained estimates for many of the parameters.
+\item $\mu$ appears to be an important parameter in determining the fit.
+\item Our initial bounding box may have been set incorrectly for several parameters - $\mu$, $\Theta$, for example - and further minima may lie outside it. 
+\end{enumerate}
+
+In the next section we further investigate the loose estimates for many of the parameters.
+
  \begin{figure}[h!]	
@@
 \end{figure}
-
- Taken together, these results suggest that simply performing a least squares fit on the available data will not give unique estimates for many of the parameters.
-In the following sections we investigate this further.
 
@@
 We see that many of the parameters give sensitivity curves of similar shapes, and have near linear dependence - this implies that the effects of perturbing any one of these parameters all look similar in terms of model output, and are hard to distinguish between.
-This may help to explain why some of our initial parameter estimates are very loose, and why there are correlations between the estimated parameter sets - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
+This may help to explain why some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
 As such, we should view estimates of these parameters with extreme caution.
@@
 Fig.
-\ref{model_output} shows model output for all state variables, using the parameter values giving the lowest error on the \textit{13\_9} dataset, and normalised to lie on the same scale.
+\ref{model_output} shows model output for all state variables, normalised to lie on the same scale  - it was made using the parameter values giving the lowest error on the \textit{13\_9} dataset, but the model output is very similar across all estimated parameter sets.
   We see that $s,m,s:m$ and $c$ respond rapidly to the forcing function, flipping between the two fixed points defined by the step function forcing almost instantly.
 By contrast, there is delay in the response of $p$ and $g$, on a timescale comparable to the period of the forcing.
-This result suggests that, at least in some of the parameter sets initially estimated, the system may have two timescales in it - a fast timescale in which \eqref{eq:s} - \eqref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in \eqref{eq:p}, \eqref{eq:g}, in which measured fluorescence changes in response to the forcing.
+This result suggests that the system may have two timescales in it - a fast timescale in which \eqref{eq:s} - \eqref{eq:c}, representing the hybridization of the sRNA and mRNA into a stable complex, equilibrate in response to external forcing, and a slower timescale, in \eqref{eq:p}, \eqref{eq:g}, in which measured fluorescence changes in response to the forcing.
 Biologically, this would correspond to \eqref{eq:p}, representing translation, being a rate limiting step.
@@
 This would explain the similarity between the sensitivity curves of many parameters - if the parameters in \eqref{eq:s} - \eqref{eq:c} can only act to alter the fixed point values of $p$ and $g$ which the system tends toward, they all play qualitatively identical roles on the time series of $z$.
-The fact that $\beta m +f_{s}\beta c$ is the only term in \eqref{eq:p}, \eqref{eq:g} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in \eqref{eq:s} - \eqref{eq:c} (Fig.
+The fact that $\beta m +f_{s}\beta c$ is the only term in \eqref{eq:p} to include $f_{s}$ and $\beta$ may also explain the similarity of the sensitivity curves for these parameters to those found in \eqref{eq:s} - \eqref{eq:c} (Fig.
 \ref{SensitivityMatrix_scaled}), by the same reasoning.
-This explanation also suggests that all parameter sets giving the same value of the model's fixed point will give similar model predictions - and there may be many, very different, parameter sets which all give the same fixed point.
+This explanation also suggests that all parameter sets giving the same value of the model's fixed point will give similar model predictions - and there may an enormous space of parameter sets which all give the same fixed point.
 Fig.
@@
 \ref{InitialResults}, we see relatively tightly estimated parameters, with no estimated values near the initial bounding box.
-Further, this simplified model achieves error values as low as the more complex model.
-Fig \ref{13_9_f_simplified} shows that the simplified algorithm is more consistently able to find the minimal function value identified in Fig.
-\ref{InitialResults_13_9_f}, possibly because the low dimension of the new model makes parameter space easier to search.
+Further, Fig \ref{13_9_f_simplified} shows that this simplified model achieves error values as low as the more complex model.
 
@@
 As before, the similarity of the $F$ and $\Theta$ curves suggests that any minima in $F - \Theta$ space will be locally non-unique \cite{Beck}.
-This may be explained as follows: neglecting the ClpX degradation term, \eqref{eq:p_simplified} is a simple exponential decay to a fixed value.
-Here, $F$ will determine the value the $p$ will tend to, and $\mu$ will determine its rate constant.
-$F$ is then a linear scaling factor, just as $\Theta$ is - both scale the value the exponential will tend to.
 
@@
 \ref{SimplifiedModelLikelihood}.
-We find a local basin in $ \mu - F$ space, pocketed by many local minima of similar fitness values.
+We find a basin in $ \mu - F$ space, pocketed by many local minima of similar fitness values.
 The optimisation algorithm will always terminate in one of these local minima, explaining the disconnected clusters shown in Fig.
@@
 In this report, we have presented a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}, and attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach similar to that found in recent synthetic biology literature \cite{Hu2015}.
-We have found many of the parameters contained in the full model to be inestimable, and suggested reasons for why this might be.
+We have found many of the parameters contained in the full model to be inestimable, and suggested reasons for this.
 We have further suggested a simplified model, which provides an equally good description of the data with fewer parameters than the initial model.
@@
 \ref{SimplifiedModelLikelihood}.
+If the experiment is simple to reproduce, another suggestion would be to decrease the gap between timepoints.
 Further modelling work may focus on improving the modelling of translation, the rate limiting step - \cite{Hu2015}, for example, uses a more detailed model than our single step description of translation.
@@
 
+\section*{Acknowledgments}
+I would like to thank my supervisor, Manish Kushwaha, for his help throughout the project, as well as Alfonso Jaramillo and Brian Munsky for advice on the initial ODE model used, and Shenshi Shen, Guillermo Rodrigo and Boris Kisov for the unpublished data used.
+In addition I would like to thank James Kermode and Peter Brommer for their helpful discussions on the limitations of the least squares fitting approach, and suggestion of the use of Gaussian Processes as an alternate method - I apologise for not having the time to explore it.
+Finally I would like the thank Annabelle Ballesta for her advice throughout the project.
 % references section
@@
   shows two distinct clusters of parameter sets,
+
+Figs.
+\ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices between parameter sets\footnote{These are computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
+An element of the correlation matrix $R_{ij}$ is then the Pearson correlation coefficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters}.
+These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists a positive correlation between values of $\mu$ and $\beta$.
+Referring to \eqref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.

### 2015-09-16 MScWriteup.tex
final changes

@@
 We provide estimates for a subset of parameters, but find many of them to be inestimable.
-We suggest this is a consequence of translation being a rate limiting step in our ODE model and suggest that a simplified model, describing translation alone, may describe the data equally well.
+We suggest this is a consequence of translation being a rate limiting step in our ODE model and propose a simplified model, describing translation alone, which we show to fit data as well as the original model.
 \end{abstract}
@@
 
-\cite{Rodrigo2012} proposed a computational methodology to design general genetic circuits based on RNA interactions, and as a case study of the methodology chose to design a synthetic sRNA- mRNA pair capable of acting in the manner described above.
+\cite{Rodrigo2012} proposed a computational methodology to design general genetic circuits based on RNA interactions, and as a case study chose to design a synthetic sRNA- mRNA pair capable of acting in the manner described above.
 The algorithm assumed an interaction scheme between the RNA's as shown in Fig.
@@
 `toehold' sequence of unpaired nucleotides to form an unstable transition state.
-This intermediate complex would then rapidly form a final, stable complex with the desired conformation.
+This intermediate complex would then form a final, stable complex with the desired conformation.
 By finding sRNA and mRNA sequences which optimised the energy landscape shown in Fig.
@@
 The data we will consider consists of two sets of individual cell time series, labelled \textit{13\_9} and \textit{14\_7} \cite{Jaramilloa}.
-They correspond to different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response (i.e the IPTG input to the AND gate is always $1$), and aTc concentrations were varied periodically.
+They correspond to different experimental runs of the above apparatus, for which IPTG concentration was held constant, at a value assumed large enough to saturate the cell's response (i.e the IPTG input to the AND gate is always a logical $1$), and aTc concentrations were varied periodically.
 Appendix \ref{Initial  Experimental Data} shows the full datasets, with their forcing functions.
@@
 
-In \eqref{eq:s} - \eqref{eq:z}, we present a modified version an existing model which describes the system \cite{Jaramilloa}, consisting of a set of ODE's with mass action kinetics.
+In \eqref{eq:s} - \eqref{eq:z}, we present a modified version an existing model which describes the system discussed above \cite{Jaramilloa}, consisting of a set of ODE's with mass action kinetics.
 Its state is given by the vector $(s,m,s:m,c,p,g,z)$, with all other variables representing model parameters.
@@
 \label{InitialResults}
+\vspace{-5mm}
 \end{figure*}
@@
 \label{CrossValidation}
+\vspace{-5mm}   
 \end{figure}
@@
 
-\begin{figure}[h]	
+\begin{figure}[t]	
 	\begin{subfigure}[h]{0.49\textwidth}
@@
 
-We see that many of the parameters give sensitivity curves of similar shapes, and have near linear dependence - this implies that the effects of perturbing any one of these parameters all look similar in terms of model output, and are hard to distinguish between.
+We see that many of the parameters give sensitivity curves of similar shapes, and have near linear dependence - this implies that the effects of perturbing any one of these parameters all look similar in terms of model output, and are hard to distinguish between \footnote{We show only a single parameter set, but the curves are a similar shape across all estimated sets - in fact, in the majority the similarity is more striking - see Fig.
+\ref{Sensitivity_scaled_other} for a typical set of curves.}.
 This may help to explain why some of our initial parameter estimates are very loose - this result suggests there is a family of parameters all of which, in terms of the model output we have available, cannot be resolved.
@@
  We note that \ref{13_9_scatterplot_simplified} shows a strong positive correlation between the minima found for $F$ and $\Theta$.
-This is consistent with the  normalised sensitivity curves for the \textit{13\_9} dataset, shown in Fig.
-\ref{Sensitivity_scaled_simplified}.
-As before, the similarity of the $F$ and $\Theta$ curves suggests that any minima in $F - \Theta$ space will be locally non-unique \cite{Beck}.
+This is consistent with the  normalised sensitivity curves, shown in Fig.
+\ref{Sensitivity_scaled_simplified}\footnote{We display the curves evaluated using the parameter set giving the lowest error, but they are similar in all estimated parameter sets.}.
+As before, the similarity of the $F$ and $\Theta$ curves suggests that any minima in $F - \Theta$ space will be locally non-unique \cite{Beck} - where these parameters do have distinguishable effects on model prediction is in the tail of fluorescence decay.
+We also note that the parameter estimates come in two distinct clusters - we expect this is due to  the error landscape containing many local minima of similar error value.
+
+ We may further investigate the error landscape by fixing the value of $\Theta$, and varying $\mu$ and $F$.
+Results are shown in Fig.
+\ref{SimplifiedModelLikelihood}, for $\Theta = 5400$ (the modal value of $\Theta$ in Fig.
+\ref{13_9_hist_simplified}).
+We find a basin in $ \mu - F$ space, pocketed by many local minima clustered around one of the peaks in Fig.
+\ref{13_9_hist_simplified}.
+Note that the tightness of the peak in Fig.
+\ref{13_9_hist_simplified}  does a poor job of representing the uncertainty in $\mu$ for a given $\Theta$ - the countours of Fig.
+\ref{Likelihood_profile_rough} are a better measure.
+
+ We also note consistency between minima found by the simplified model and the more complex one.
+Setting $\Theta = 1000$ (Fig.
+\ref{Likelihood_profile_theta_1000}), we find minima at $\mu \approx 0.045$, $F \approx 235$, the $\mu$ and $\beta m +f_{s}\beta c$ values shown in Figs.
+\ref{InitialResults_13_9_hist} and \ref{fixedpoint_f}.
 
@@
 \end{figure}
-
- We also note that the parameter estimates come in two distinct clusters.
-We further investigate by fixing the value of $\Theta$, and plotting the error as a function of $\mu$ and $F$.
-Results are shown in Fig.
-\ref{SimplifiedModelLikelihood}.
-We find a basin in $ \mu - F$ space, pocketed by many local minima of similar fitness values.
-The optimisation algorithm will always terminate in one of these local minima, explaining the disconnected clusters shown in Fig.
-\ref{13_9_scatterplot_simplified}.
-Thus the tightness of each peak in Fig.
-\ref{13_9_hist_simplified} is misleading - the uncertainty in parameter estimates is better reflected by the contours of Fig.
-\ref{Likelihood_profile_fine}.
 
@@
 
-In this report, we have presented a system of ODE's to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}, and attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach similar to that found in recent synthetic biology literature \cite{Hu2015}.
+In this report, we have presented a system of ODEs to model a recently proposed synthetic RNA regulatory circuit \cite{Rodrigo2012}, and attempted to estimate the model's parameters using existing time series data, with a least squares minimisation approach similar to that found in recent synthetic biology literature \cite{Hu2015}.
 We have found many of the parameters contained in the full model to be inestimable, and suggested reasons for this.
@@
 This method would explicitly provide us with marginal distributions of parameter probabilities, giving a more complete picture of the best parameter sets and their uncertainties than we currently have - we would still be able to pick out a maximum likelihood point estimator, as we do now, but rather than getting an ad hoc picture of the uncertainty in the parameters by repeatedly running the minimisation algorithm, the marginals would give us this information directly.
-However, this change in methodology would not have allowed us to gain estimates of the parameters in the original model, and is not necessary for the simplified one.
-It may be useful to use in place of the least squares approach if further modelling of the translation step is pursued.
+Having said this, a change in methodology would not have allowed us to gain estimates of the parameters in the original model - it may be worth using in place of the least squares approach if further modelling of the translation step is pursued.
 
@@
 
-\section{Additional Initial Results figures}
+\section{Initial fitting additional results}
 \label{Additional Initial Results figures}
@@
 \begin{figure}[H]
-    \begin{subfigure}[c]{0.49\textwidth}
-        \centering
-    \includegraphics[scale = 0.235, clip = true, trim = 120 0 65 0]{13_9_heatmap.eps}
-        \caption{Correlation matrix of estimated parameter sets, from the  \textit{13\_9} dataset. }
-        \label{InitialResults_13_9_correlation}
-    \end{subfigure}
-    \begin{subfigure}[c]{0.49\textwidth}
-    \centering
-        \includegraphics[scale = 0.235, clip = true, trim = 80 0 70 0]{14_7_heatmap.eps}
-        \caption{Correlation matrix of estimated parameter sets, from the  \textit{14\_7} dataset.}
-        \label{InitialResults_14_7_correlation}
-    \end{subfigure}
 
@@
     \end{subfigure}
-    \caption{Inter-parameter correlation values and error function values, found from 200 initial parameter estimates, for the \textit{13\_9} and \textit{14\_7} datasets.}     
+    \caption{Error function values, found from 200 initial parameter estimates, for the \textit{13\_9} and \textit{14\_7} datasets.}     
 \label{InitialResults_additional}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+        \includegraphics[scale = 0.35, clip = true, trim = 0 0 0 0]{Sensitivty_scaled_other}
+        \caption{A typical set of normalised sensitivity curves found from the initial 200 sets of estimated parameters, \textit{13\_9} dataset.
+Note that the all curves except those for $\Theta$ and $\mu$ lie very close to one another.}
+        \label{Sensitivity_scaled_other}
     \end{figure}
@@
         \includegraphics[scale = 0.25]{13_9_f_simplified}
-        \caption{}
+        \caption{Error function values, found from 400 parameter estimates using the simplified model \eqref{eq:p_simplified} - \eqref{eq:z_simplified}, for the \textit{13\_9} dataset.}
         \label{13_9_f_simplified} 
@@
 
+\begin{figure}[h]	
+    \centering
+        \includegraphics[scale = 0.25]{Likelihood_profile_theta_1000}
+        \caption{A basin of low error values in $\mu - F$ space, $\Theta = 1000$. Note the colour scaling maps all values above $10$ to the maximum colour.}
+        \label{Likelihood_profile_theta_1000} 
+\end{figure}
+
 \end{document}
-
-We note that it is possible that this separation of timescales is an artefact of the fitting, and not necessarily true of the system.
-We can only evaluate model predictions on the parameter sets the algorithm returns us, and it seems the case that the parameter sets found by the algorithm have this feature.
-Hypothetically, there may exist parameter sets where all state variables vary on a similar timescale - all we can say is that, in the search region specified, the algorithm has not found any which give a low error.
-
-We also note that while some parameters may be hard to get any realistic bounds on, values for others - $\mu, \Theta$ for example - may already be available.
-
-This point raises another issue - the arbitrary nature of the initial bounding box.
-Without good biological constraints on most parameter values, they have been bounded loosely and not necessarily sensibly.
-Rate constants can vary by many orders of magnitude, and it may be possible that the true optimum lies entirely outside the bounding box specified.
-Equally,  when our fitness function has turned out to be very flat in some directions, or multimodal, good biologically motivated parameter bounds may be needed to ensure our minimisation algorithm does not wander outside realistic parameter space - the algorithm may present several equally good regions of parameter space, and improved bounds will help us choose between them.
-
- The model may also have general structural issues.
-\cite{Mclean2012, Grewal1976} give general tests which can be applied to a system of ODEs to determine if all parameters are in principle estimable, and apply them successfully to models with comparable numbers of parameters to ours.
-
- Thus, further work may consist of tightening the bounds on parameter values, carrying out a structural analysis, and possibly simplifying the model.
-A much simpler phenomenological model, consisting of  just \eqref{eq:p} and \eqref{eq:g}, may explain existing data perfectly well without the need for inestimable parameters.
-Experimentally, one option  may be to use different forcing functions.
-The effect of this on parameter estimability might be anticipated by running a sensitivity analysis using this new forcing, to see if the form of Fig.
-\ref{SensitivityMatrix} (which shows model predictions for all states) changes.
-Another option, though perhaps experimentally infeasible, would be to directly observe other components of the state vector - $s$, $m$, $s:m$ and $c$.
-This would improve parameter estimates by giving direct observations of the effects of parameters contained in \eqref{eq:s} - \eqref{eq:c}.
-
- , and that the values of $\mu$ and $\Theta$ shown in Fig.
-\ref{13_9_hist_simplified} differ from those found in the initial fitting, Fig.
-\ref{InitialResults_13_9_hist} - in particular, $\Theta$ has increased beyond the upper bound initially specified.
-The clustering of values suggests that there may be several disconnected local minima, all with similar error values.
-Within each of these minima, $\Theta$ and $\mu$ will show tight clustering.
-
-  The initial upper bound on $\Theta$ set in table \ref{Parameter_literature_review_table} will cut off some of these minima
-
-  This may be explained with reference to the clustering of parameter estimatees \ref{13_9_scatterplot_simplified}, which shows that
-
-Finally, we note that that \ref{13_9_scatterplot_simplified}
-
-  shows two distinct clusters of parameter sets,
-
-Figs.
-\ref{InitialResults_13_9_correlation}, \ref{InitialResults_14_7_correlation} show correlation matrices between parameter sets\footnote{These are computed by viewing each of the estimated parameter sets as a sample from a multivariate random variable.
-An element of the correlation matrix $R_{ij}$ is then the Pearson correlation coefficient between the $i^\mathrm{th}$ and $j^\mathrm{th}$ random variable - or, in our case, the $i^\mathrm{th}$ and $j^\mathrm{th}$ parameters}.
-These correlation matrices indicate that there are spaces of parameters within which the fitness function remains approximately constant - for example, in both datasets there exists a positive correlation between values of $\mu$ and $\beta$.
-Referring to \eqref{eq:p}, this makes heuristic sense - the two parameters may have similar effects on model predictions, and may be able to co-vary in such a way as to leave the model prediction unchanged.

### 2015-09-16 MScWriteup.tex
final version?

@@
 \subsection{Single Cell Fluorescence Data}
+\label{Single_Cell_Fluorescence_Data}
 
@@
 Firstly a generic degradation rate $\delta_{g}$ is included, assumed identical for the mature and immature species, along with the dilution rate $\mu$ shared by all species.
-Secondly, in the experimental setup we can arrange for GFP molecules to be produced with a \textit{degradation tag} attached to them \cite{Hersch2004}.
+Secondly, in the experiments described in section \ref{Single_Cell_Fluorescence_Data}, GFP molecules were produced with a \textit{degradation tag} attached to them \cite{Hersch2004}.
 This tag is identified by an enzyme, ClpX, which will then degrade the molecule the tag is  attached to.
@@
 \label{InitialResults}
-\vspace{-5mm}
+\vspace{-6mm}
 \end{figure*}
@@
  There are two main reasons why a parameter may not be estimable \cite{Mclean2012,Yao2003, Beck,Jimenez-Hornero2008,Grewal1976}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others.
-
-These problems may stem from structural inadequacies in the model (often termed \textit{structural identifiability}), in which two different parameter sets can give identical model predictions \cite{Jimenez-Hornero2008,Grewal1976 }.
-If this is the case, no amount of experimental data will allow us to estimate parameters, and we must consider reformulating the model\footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product $\beta_{1}\beta_{2}$ can ever be estimated, a problem no amount of data can fix.}.
-Problems may also arise for more practical reasons (\textit{practical identifiability} \cite{Mclean2012}).
-For example, it is possible that in the experimental regime we operate in (eg.
-the forcing function we use, the components of the model's state we can observe), parameter effects may be weak, or highly correlated, but in other regimes this is not true.
-In this case, parameter estimates may be improved by taking data in more varied experimental conditions, or attempting to increase the number of model outputs observed - in our case, observing more of the state vector than simply $z$.
-
- We may begin to investigate these issues in our model by performing a local sensitivity analysis about one of the solutions found in our initial parameter estimation.
+We may begin to investigate these issues in our model by performing a local sensitivity analysis about the solutions found in our initial parameter estimation.
 We numerically estimate the sensitivity matrix, $S$:
@@
 
-We see that many of the parameters give sensitivity curves of similar shapes, and have near linear dependence - this implies that the effects of perturbing any one of these parameters all look similar in terms of model output, and are hard to distinguish between \footnote{We show only a single parameter set, but the curves are a similar shape across all estimated sets - in fact, in the majority the similarity is more striking - see Fig.
+We see that many of the parameters give sensitivity curves of similar shapes, and have near linear dependence - this implies that the effects of perturbing any one of these parameters all look similar in terms of model output, and are hard to distinguish between \footnote{Though we only show sensitivity curves evaluated at a single parameter set, their shape is similar across all estimated sets - in fact, in the majority of the 200 estimated sets, the similarity of the sensitivity curves is more striking - see Fig.
 \ref{Sensitivity_scaled_other} for a typical set of curves.}.
@@
 As such, we should view estimates of these parameters with extreme caution.
-Note that the sensitivity curve that looks least similar to the others in Fig.
-\ref{SensitivityMatrix_scaled} - $\mu$ - corresponds to a relatively tightly estimated value in Fig.
+Note that the sensitivity curve that looks least similar to the others in Figs.
+\ref{SensitivityMatrix_scaled}, \ref{Sensitivity_scaled_other} - $\mu$ - corresponds to a relatively tightly estimated value in Fig.
 \ref{InitialResults}.
-Note also that the curve for $\Theta$ is very similar to those for the remaining parameters.
+Note also that the curve for $\Theta$ is similar to those for the remaining parameters.
 This suggests that our estimated value of $\Theta$ in Fig.
@@
 \end{document}
+
+These problems may stem from structural inadequacies in the model (often termed \textit{structural identifiability}), in which two different parameter sets can give identical model predictions \cite{Jimenez-Hornero2008,Grewal1976 }.
+If this is the case, no amount of experimental data will allow us to estimate parameters, and we must consider reformulating the model\footnote{A simple example of a structurally non - identifiable model is $y = \beta_{1}\beta_{2}x$, where we are given data $(x,y)$, and asked to estimate parameters $\beta_{1}$ and $\beta_{2}$ - we can see that, in principle, only the product $\beta_{1}\beta_{2}$ can ever be estimated, a problem no amount of data can fix.}.
+Problems may also arise for more practical reasons (\textit{practical identifiability} \cite{Mclean2012}).
+For example, it is possible that in the experimental regime we operate in (eg.
+the forcing function we use, the components of the model's state we can observe), parameter effects may be weak, or highly correlated, but in other regimes this is not true.
+In this case, parameter estimates may be improved by taking data in more varied experimental conditions, or attempting to increase the number of model outputs observed - in our case, observing more of the state vector than simply $z$.

### 2015-09-16 MScWriteup.tex
final final?

@@
 \ref{InitialResults_13_9_hist} do not give the minimum error function value found - instead, the majority find a local minimum slightly above it.
-In fact, if we examine the estimated parameter sets giving the lowest error values, several have $\mu = 0.05$, the upper bound set.
+In fact, if we examine the estimated parameter sets giving the lowest error values for the \textit{13\_9} dataset, several have $\mu = 0.05$, the upper bound set.
 Thus, in the tightly clustered $\mu$ histogram, we must bear in mind that $\mu \approx 0.044$ is not a unique optimum - the error landscape may indeed be pocketed by many local minima, all giving a similar quality of fit.
@@
  There are two main reasons why a parameter may not be estimable \cite{Mclean2012,Yao2003, Beck,Jimenez-Hornero2008,Grewal1976}: Model predictions may be insensitive to the value of a particular parameter, or the effects of varying one parameter on model predictions may be highly correlated with the effects of varying several others.
-We may begin to investigate these issues in our model by performing a local sensitivity analysis about the solutions found in our initial parameter estimation.
-We numerically estimate the sensitivity matrix, $S$:
+We begin to investigate these issues in our model by performing a local sensitivity analysis about the solutions found in our initial parameter estimation.
+We numerically estimate the sensitivity matrix, $S$ \cite{Mclean2012}:
 

### 2015-09-22 Presentation/MscPresentation.tex
Initial presentation commit

% $Header: /Users/joseph/Documents/LaTeX/beamer/solutions/conference-talks/conference-ornate-20min.en.tex,v 90e850259b8b 2007/01/28 20:48:30 tantau $

\documentclass{beamer}

% This file is a solution template for:

% - Talk at a conference/colloquium.
% - Talk length is about 20min.
% - Style is ornate.

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\mode<presentation>
{
  \usetheme{Warsaw}
  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}

\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\usepackage{chemfig}
\DeclareMathOperator*{\argmin}{arg\,min}

\title % (optional, use only with long paper titles)
{Modelling RNA Oscillatory Circuits}

\subtitle
{}

\author % (optional, use only with lots of authors)
{J. Binysh\inst{1}}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[University of Warwick] % (optional, but mostly needed)
{
  \inst{1}%
 	Centre for Complexity Science\\
  University of Warwick
 }
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

%\date[CFP 2003] % (optional, should be abbreviation of conference name)
%{Conference on Fabulous Presentations, 2003}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

%\subject{Theoretical Computer Science}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
%\AtBeginSubsection[]
%{
%  \begin{frame}<beamer>{Outline}
%    \tableofcontents[currentsection,currentsubsection]
%  \end{frame}
%}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Structuring a talk is a difficult task and the following structure
% may not be suitable. Here are some rules that apply for this
% solution:

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.

% - A conference audience is likely to know very little of what you
%   are going to talk about. So *simplify*!
% - In a 20min talk, getting the main ideas across is hard
%   enough. Leave out details, even if it means being less precise than
%   you think necessary.
% - If you omit details that are vital to the proof/implementation,
%   just say so once. Everybody will be happy with that.

\section{Description of RNA Oscillatory system}

\begin{frame}{The RNA Oscillatory System}{}
  % - A title should summarize the slide in an understandable fashion
  %   for anyone how does not follow everything on the slide itself.
\includegraphics[trim = 130 600 30 30,page=4,clip = true]{../Figures/pnas1203831109.pdf}
  \begin{itemize}
    \item  mRNA produced `self repressed'  - tail folded over the RBS
    \item  sRNA binds to it - new complex has RBS uncovered.
    \end{itemize}
\end{frame}

\subsection{Recent Experimental data}

\begin{frame}{Recent Experimental data}{}
\includegraphics[scale = 0.55, trim = 0 300 0 280 ,clip = true]{../Figures/13_9}
  \begin{itemize}
    \item  Time series of individual cell fluorescence, forced with aTc.
    \end{itemize}
\end{frame}

\subsection{ODE Model for system}
\begin{frame}{ODE Model for system}{}
\begin{align*} 
\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\mu + \delta_{s})s -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m \\
\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\mu + \delta_{m})m -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m  \\
\frac{ds:m}{dt} & = k_{\mathrm{on}}sm  - (k_{\mathrm{off}}+ k_{\mathrm{hyb}})s:m  -(\mu + \delta_{sm} )s:m \\
\frac{dc}{dt} & = k_{\mathrm{hyb}}s:m  -(\mu + \delta_{c})c  \\
\end{align*}
\vspace{-3em}
  \begin{itemize}
    \item  Hybridization of sRNA($s$) and mRNA ($m$), into unstable complex ($s:m$) then stable one ($c$).
    \end{itemize}
\end{frame}

\begin{frame}{ODE Model for system}{}
\begin{align*} 
\frac{dp}{dt} & = \beta m +f_{s}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
\frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
z &= z_{0} +\frac{g}{\Theta} 
\end{align*}

  \begin{itemize}
    \item Translation of stable complex ($c$) into immature GFP ($p$). 
    \item maturation of GFP ($g$), machine calibration giving measured fluorescence ($z$).
    \end{itemize}

\end{frame}

\section{Parameter Estimation in ODE Model}

\subsection{Methodology - least squares error minimisation}

\begin{frame}{Parameter Estimation in ODE Model - Methodology}
  \begin{block}{Goal}
 Estimate unknown parameters in model using time series data.
    \end{block}

      \begin{itemize}
    \item  Least squares error minimization approach
    \begin{align*}
\argmin_{\boldsymbol{\theta}} \sum_{i =1}^{n} (z_{\mathrm{exp, mean}}(t_{i}) - z(t_{i},\boldsymbol{\theta}))^2.
\end{align*}
\item Used evolutionary algorithm, CMA-ES, to perform minimisation.
    \end{itemize}
\end{frame}

\small
\begin{frame}{Parameters to be estimated in our model}
\begin{align*} 
\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\alert{\mu} + \alert{\delta_{s}})s -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m \\
\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\mu + \alert{\delta_{m}})m -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m  \\
\frac{ds:m}{dt} & = k_{\mathrm{on}}sm  - (k_{\mathrm{off}}+ \alert{k_{\mathrm{hyb}}})s:m  -(\mu + \delta_{m} )s:m \\
\frac{dc}{dt} & = k_{\mathrm{hyb}}s:m  -(\mu + \delta_{m})c  \\
\frac{dp}{dt} & = \alert{\beta} m +\alert{f_{s}}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
\frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
z &= z_{0} +\frac{g}{\alert{\Theta}} 
\end{align*}
\end{frame}
\normalsize

\subsection{Results - Inestimable parameters, model simplifications}

\begin{frame}{Initial Minimization Results}
  \includegraphics[scale = 0.26, clip = true, trim = 50 0 0 0]{../Figures/13_9_bestPlot}
\end{frame}

\begin{frame}{Initial Minimization Results}
  \includegraphics[scale = 0.22, clip = true, trim = 0 0 0 0]{../Figures/13_9_hist}
        \begin{itemize}
    \item Many parameters poorly estimated.
            \end{itemize}
\end{frame}

\begin{frame}{Sensitivity Analysis}
  \includegraphics[scale = 0.22, clip = true, trim = 00 0 0 0]{../Figures/Sensitivty_scaled_other}
          \begin{itemize}
    \item Sensitivity analysis suggests perturbations of several parameters all give similar effects on model output - \alert{hard to resolve}
            \end{itemize}
\end{frame}

\begin{frame}{Sensitivity Analysis}
  \includegraphics[scale = 0.24, clip = true, trim = 0 0 0 0]{../Figures/fixedpoint_f}
\begin{itemize}
	\item Similar fixed point values of translation forcing term, ${\beta} m +{f_{s}}\beta c$, across estimated parameter sets.
	\item Suggests translation may be rate limiting step.
\end{itemize}
\end{frame}

\begin{frame}{Simplified Model}
\begin{align*}
\frac{dp}{dt} & = \alert{F}y(t)-(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
\frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g}  \\
z &= z_{0} +\frac{g}{\alert{\Theta}} 
\end{align*}
\begin{itemize}
\item  Model only rate limiting steps - force translation term directly.
\item 3 unknown parameters
\end{itemize}
\end{frame}

\begin{frame}{Simplified Model Results}
  \includegraphics[scale = 0.28, clip = true, trim = 100 0 0 400]{../Figures/13_9_hist_simplified}
  \begin{itemize}
\item  Error values as low as full model
\item Clearer parameter estimation results
\item Two peak structure due to local minima
\end{itemize}
\end{frame}

\begin{frame}{Simplified Model Results}
  \includegraphics[scale = 0.25, clip = true, trim = 60 0 0 00]{../Figures/Likelihood_profile_rough}
    \begin{itemize}
\item  Error landscape, $\Theta = 1000$.
\end{itemize}
\end{frame}

\subsection{Future Work}

\begin{frame}{Future Work}

  % Keep the summary *very short*.
  \begin{itemize}
  \item Fluorescence data \alert {not enough} to estimate all unknown parameters 
  \begin{itemize}
  \item Further experiments ?
  \item Bounds on unknown parameter values.
  \end{itemize} 
  \item Improve modelling of translation step
  \item Change methodology - Bayesian Methods?

  \end{itemize}

  \end{frame}

  \section*{Summary}

\end{document}

### 2015-09-22 Presentation/MscPresentation_NOTES.tex
some accompanying notes

% $Header: /Users/joseph/Documents/LaTeX/beamer/solutions/conference-talks/conference-ornate-20min.en.tex,v 90e850259b8b 2007/01/28 20:48:30 tantau $

\documentclass[8pt]{beamer}

% This file is a solution template for:

% - Talk at a conference/colloquium.
% - Talk length is about 20min.
% - Style is ornate.

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\mode<presentation>
{
  \usetheme{Warsaw}
  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}

\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\usepackage{chemfig}
\DeclareMathOperator*{\argmin}{arg\,min}

\title % (optional, use only with long paper titles)
{Modelling RNA Oscillatory Circuits}

\subtitle
{}

\author % (optional, use only with lots of authors)
{J. Binysh\inst{1}}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[University of Warwick] % (optional, but mostly needed)
{
  \inst{1}%
 	Centre for Complexity Science\\
  University of Warwick
 }
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

%\date[CFP 2003] % (optional, should be abbreviation of conference name)
%{Conference on Fabulous Presentations, 2003}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

%\subject{Theoretical Computer Science}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
%\AtBeginSubsection[]
%{
%  \begin{frame}<beamer>{Outline}
%    \tableofcontents[currentsection,currentsubsection]
%  \end{frame}
%}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
\begin{itemize}
\item My project consisted of trying to estimate unknown parameters in an ODE model, representing a recently designed genetic regulatory circuit, using some recent experimental data.
\item Begin by giving biological background, describing the system we are interested in. Then discuss recent experimental data, and the ODE model proposed to represent it.
\item Then move on to talk about how to estimate unknown parameters in this model - first discussing methodology, then results.
\item I will show that the majority of the parameters in the model cannot be uniquely estimated given the available data. I will discuss why this might be the case, and suggest the use of a simplified model, which can describe available data equally well.
\item then conclude and suggest further work.
\end{itemize}
\end{frame}

\section{Description of RNA Oscillatory system}
\begin{frame}{Description of RNA Oscillatory system}
\begin{itemize}
\item In the cell, DNA is copied into an RNA molecule, which then gets made into a protein.
\item This process is naturally regulated in many ways. One stage it can be regulated at is when the RNA - called mRNA - is on its way to get made into a protein
\item In order for the mRNA to be made into a protein, a piece of molecular machinerey called the Ribosome must attach itself to it. This occurs at a place on the mRNA molecule called the Ribsome Binding site (RBS). If the RBS is blocked, the mRNA cannot be made into a protein
\item mRNA can `self repress' - its shape is such that the molecules tail is looped over the RBS
\item This self repression can be undone by introducing another RNA - a small RNA which binds to the mRNA, and uncovers the RBS
\item Such a setup can be synthetically engineered - mRNA-sRNA pairs that act in this fashion can be designed. In addition, we can choose that the protein made be GFP, which fluoresces, so its concentration in the cell can be monitored.

\item Further, the whole setup can be placed inside a cell, and indirectly controlled by two chemcials, aTC and IPTG, whose concentrations affect the rates of mRNA and sRNA production.
\item Whole system forms a logical `and' gate.
\end{itemize}
\end{frame}

\subsection{Recent Experimental data}

\begin{frame}{Recent Experimental data}{}
  \begin{itemize}
    \item  In recent experiments, cells with the above set-up in them have been forced with a varying aTc concentration - what we see here are individual cell fluoresence time series, with aTC forcing schematically shown. 
    \end{itemize}
\end{frame}

\subsection{ODE Model for system}
\begin{frame}{ODE Model for system (1)}{}
\begin{itemize}
\item We can model this system with a set of ODE's with mass action kinetics.
\item The first set represent the production of the two RNA's discussed above, and their hybridization. This is represented as a two step process - first into an unstable, then a stable, complex.
\item aTc forcing is modelled via the y term. The various delta's are just degradation rates of the complexes involved, and mu represents dilutions of the chemicals due to cell growth.
\end{itemize} 
\end{frame}

\begin{frame}{ODE Model for system (2)}{}
\begin{itemize}
\item The next set of equations explicitly models protein production and observed fluorescence. beta m + fs beta c term describes the translation of the mRNA into an immature GFP (p). This then matures into mature GFP (g), which fluoresces (z)  
\end{itemize}
\end{frame}

\section{Parameter Estimation in ODE Model}

\subsection{Methodology - least squares error minimisation}

\begin{frame}{Parameter Estimation in ODE Model}
  \begin{itemize}
\item So we have our ODE model, and some recent time series data. Our goal is to estimate some of the unknown parameters in the described model, by fitting it to the data.
\item To do this, we  numerically simulate our ODE model, giving a predicted fluorescence time series for some set of parameters. We then define a least squares error between this simulation and our observed data, and seek to find the set of parameters which minimise this error.
\item The error landscape may have many minima - to perform the minimisation we use an Evolutionary algorithm, the CMA-ES.
\end{itemize}
\end{frame}

\begin{frame}{Parameters to be estimated in our model}
  \begin{itemize}
\item In our case, we have 9 unknown parameters, highlighted in red. 5 (name( only appear in the model before the translation step, in the initial hybridization process. Two (name) are associated with the translation process. Mu appears throughout the equations, and theta just represents machine calibration.

\end{itemize}
\end{frame}
\normalsize

\subsection{Results - Inestimable parameters, model simplifications}

\begin{frame}{Initial Estimation Results}
\begin{itemize}
    \item Prediction from a typical set of parameters found shows model capable of quantitatively capturing data
\end{itemize}
\end{frame}

\begin{frame}{Initial Estimation Results}
\begin{itemize}
    \item However, many parameters are poorly estimated. These histograms were made by randomly starting the CMA-ES at a point within an initial bounding box, and letting it run 200 times. We see little consistency in estimates for many parameters. The x axis on the histograms are the initial bounding box given, and we see some estimates spread right across the range. 
    \item However, some parmaters (mu, theta) are relatively consistently estimated.
    \item Why is this the case? 
\end{itemize}
\end{frame}

\begin{frame}{Sensitivity Analysis}
\begin{itemize}
    \item Taking one of the estimated parameter sets, and performing a local sensitivity analysis about it, suggests a possible explanation.
    \item We see that many parameters, when slightly perturbed, have similar effects on model output - thus making their effects hard to distinguish , and so hard to simultaneously estimate 
    \item Curves for Mu and theta relatively distinct.
    \item This result suggests the minimisation algorithm may be working with many fewer degrees of freedom than we gave it - only 3, rather than 9. 
\end{itemize}
\end{frame}

\begin{frame}{Model fixed point for estimated parameters}
\begin{itemize}
	\item Although individual parameter estimates are very loose, if we look at all estimated parameter sets, there is some consistency to be found. Recall that the initial hybridization steps etc were communicated to the equations modelling translation by the ${\beta} m +{f_{s}}\beta c$ term (flick back). If we solve for the models fixed point, and look at the fixed point values of ${\beta} m +{f_{s}}\beta c$, we see similarity between all parameter sets. This suggests that the algorithm is really only altering the values of ${\beta} m +{f_{s}}\beta c$, mu and theta, rather than the full 9 degrees of freedom.
	\item a biological interpretation of this fact may be that translation is a rate limiting step. 
\end{itemize}
\end{frame}

\begin{frame}{Simplified Model (1)}
\begin{itemize}
\item This suggests we can simplify by assuming the first set of equations occur instantly, and instead just force and model the translation step. We can then re-estimate our 3 parameters as before.
\end{itemize}
\end{frame}

\begin{frame}{Simplified Model Results (2)}
  \begin{itemize}
\item As before, running the CMA- ES provides us with parameter estimates. Unlike before, they are clear of the initial bounding box Further, we achieve error values as low as full model. Two peak structure likely a consequence of local minima.
\end{itemize}
\end{frame}

\begin{frame}{Simplified Model Results (3)}
  \begin{itemize}
\item The model is also now simple enough we can plot the error landscape, the contours of which offer a better indiction of uncertainity in F and theta than the histograms do.
\end{itemize}
\end{frame}

\subsection{Future Work}

\begin{frame}{Future Work}

  \end{frame}

  \section*{Summary}

\end{document}

### 2015-09-23 Presentation/MscPresentation.tex
some accompanying notes
made new figures, rewording start

@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\usepackage{color}
 
@@
 
-\begin{frame}{Outline}
-  \tableofcontents
-  % You might wish to add the option [pausesections]
-\end{frame}
-
 % Structuring a talk is a difficult task and the following structure
@@
   %   for anyone how does not follow everything on the slide itself.
-\includegraphics[trim = 130 600 30 30,page=4,clip = true]{../Figures/pnas1203831109.pdf}
+  \begin{center}
+  \includegraphics[trim = 0 0 0 30,clip = true,scale = 0.275]{Figures/schematic_initial.png}
+  \end{center}
+
   \begin{itemize}
@@
 
-\subsection{Recent Experimental data}
-
-\begin{frame}{Recent Experimental data}{}
-\includegraphics[scale = 0.55, trim = 0 300 0 280 ,clip = true]{../Figures/13_9}
+\subsection{ODE Model for system}
+\begin{frame}{ODE Model for system (1) - Complex Formation}{}
+\begin{columns}
+\column{.45\textwidth}
+\includegraphics[trim = 135 0 0 0,clip = true,scale = 0.31]{Figures/schematic_binding}
+\column{.5\textwidth}
   \begin{itemize}
-    \item  Time series of individual cell fluorescence, forced with aTc.
+    \item  Hybridization of sRNA \textcolor{blue}{($s$)} and mRNA \textcolor{blue}{($m$)}, into unstable complex \textcolor{blue}{($s:m$)} then stable one \textcolor{blue}{($c$)}.
     \end{itemize}
-\end{frame}
-
-\subsection{ODE Model for system}
-\begin{frame}{ODE Model for system}{}
+\footnotesize
 \begin{align*} 
@@
 \end{align*}
-\vspace{-3em}
-  \begin{itemize}
-    \item  Hybridization of sRNA($s$) and mRNA ($m$), into unstable complex ($s:m$) then stable one ($c$).
-    \end{itemize}
+\end{columns}
 \end{frame}
 
-\begin{frame}{ODE Model for system}{}
+\begin{frame}{ODE Model for system (2) - Translation}{}
+\begin{columns}
+\column{.45\textwidth}
+\includegraphics[trim = 0 0 0 0,clip = true,scale = 0.31]{Figures/schematic_translation}
+\column{.5\textwidth}
+  \begin{itemize}
+    \item Translation of stable complex \textcolor{blue}{($c$)} into immature GFP \textcolor{blue}{($p$)}. 
+    \item maturation of GFP \textcolor{blue}{($g$)}, machine calibration giving measured fluorescence ($z$).
+    \end{itemize}
+\footnotesize
 \begin{align*} 
@@
 \end{align*}
-
-  \begin{itemize}
-    \item Translation of stable complex ($c$) into immature GFP ($p$). 
-    \item maturation of GFP ($g$), machine calibration giving measured fluorescence ($z$).
-    \end{itemize}
-
+\end{columns}
 \end{frame}
@@
 
+\small
+\begin{frame}{Parameters to be estimated in our model}
+\begin{align*} 
+\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\alert{\mu} + \alert{\delta_{s}})s -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m \\
+\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\alert{\mu} + \alert{\delta_{m}})m -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m  \\
+\frac{ds:m}{dt} & = \alert{k_{\mathrm{on}}}sm  - (\alert{k_{\mathrm{off}}}+ \alert{k_{\mathrm{hyb}}})s:m  -(\alert{\mu} + \alert{\delta_{m}} )s:m \\
+\frac{dc}{dt} & = \alert{k_{\mathrm{hyb}}}s:m  -(\alert{\mu} + \alert{\delta_{m}})c  \\
+\frac{dp}{dt} & = \alert{\beta} m +\alert{f_{s}}\beta c -(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
+\frac{dg}{dt} & = \gamma p - (\alert{\mu} + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
+z &= z_{0} +\frac{g}{\alert{\Theta}} 
+\end{align*}
+\end{frame}
+\normalsize
+
 \subsection{Methodology - least squares error minimisation}
 
-\begin{frame}{Parameter Estimation in ODE Model - Methodology}
+\begin{frame}{Parameter Estimation in ODE Model}
   \begin{block}{Goal}
@@
 
-\small
-\begin{frame}{Parameters to be estimated in our model}
-\begin{align*} 
-\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\alert{\mu} + \alert{\delta_{s}})s -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m \\
-\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\mu + \alert{\delta_{m}})m -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m  \\
-\frac{ds:m}{dt} & = k_{\mathrm{on}}sm  - (k_{\mathrm{off}}+ \alert{k_{\mathrm{hyb}}})s:m  -(\mu + \delta_{m} )s:m \\
-\frac{dc}{dt} & = k_{\mathrm{hyb}}s:m  -(\mu + \delta_{m})c  \\
-\frac{dp}{dt} & = \alert{\beta} m +\alert{f_{s}}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
-\frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
-z &= z_{0} +\frac{g}{\alert{\Theta}} 
-\end{align*}
-\end{frame}
-\normalsize
-
 \subsection{Results - Inestimable parameters, model simplifications}
 
-\begin{frame}{Initial Minimization Results}
+\begin{frame}{Initial Estimation Results}
   \includegraphics[scale = 0.26, clip = true, trim = 50 0 0 0]{../Figures/13_9_bestPlot}
@@
 
-\begin{frame}{Initial Minimization Results}
+\begin{frame}{Initial Estimation Results}
   \includegraphics[scale = 0.22, clip = true, trim = 0 0 0 0]{../Figures/13_9_hist}
@@
 
-\begin{frame}{Sensitivity Analysis}
+\begin{frame}{Model output for all state variables}
+  \includegraphics[scale = 0.25, clip = true, trim = 30 0 0 0]{../Figures/model_output}
+\end{frame}
+
+\begin{frame}{Model fixed point for estimated parameters}
   \includegraphics[scale = 0.24, clip = true, trim = 0 0 0 0]{../Figures/fixedpoint_f}
@@
 
-\begin{frame}{Simplified Model}
+\begin{frame}{Simplified Model (1)}
 \begin{align*}
@@
 
-\begin{frame}{Simplified Model Results}
+\begin{frame}{Simplified Model Results (2)}
   \includegraphics[scale = 0.28, clip = true, trim = 100 0 0 400]{../Figures/13_9_hist_simplified}
@@
 
-\begin{frame}{Simplified Model Results}
+\begin{frame}{Simplified Model Results (3)}
   \includegraphics[scale = 0.25, clip = true, trim = 60 0 0 00]{../Figures/Likelihood_profile_rough}

### 2015-09-24 Presentation/MscPresentation.tex
reworking slides
reworked presentation

@@
 {
-  \usetheme{Warsaw}
+  \usetheme{default}
   % or ...
@@
 
-\usepackage{chemfig}
+\usepackage{epstopdf,chemfig}
+\usepackage{xcolor}
 \DeclareMathOperator*{\argmin}{arg\,min}
-\usepackage{color}
+\newenvironment{greyedout}{\color{gray}}{\ignorespacesafterend}
 
 \title % (optional, use only with long paper titles)
-{Modelling RNA Oscillatory Circuits}
+{Modelling RNA Circuits}
 
@@
 
-\section{Description of RNA Oscillatory system}
+\section{Description of RNA system}
 
-\begin{frame}{The RNA Oscillatory System}{}
+\begin{frame}{The RNA System}{}
   % - A title should summarize the slide in an understandable fashion
@@
 \small
-\begin{frame}{Parameters to be estimated in our model}
-\begin{align*} 
+\begin{frame}{Full model, with Parameters to be estimated}
+\begin{align} 
 \frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\alert{\mu} + \alert{\delta_{s}})s -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m \\
@@
 z &= z_{0} +\frac{g}{\alert{\Theta}} 
-\end{align*}
+\end{align}
 \end{frame}
@@
 
+\begin{frame}{Recent Experimental Data}{}
+\begin{columns}
+\column{.49\textwidth}
+\begin{figure}
+   \includegraphics[trim = 10 300 270 290,clip = true,scale = 0.55]{Figures/13_9}
+\end{figure}
+\column{.49\textwidth}
+\vspace{-5mm}
+\begin{figure}
+   \includegraphics[trim = 0 40 0 0 ,page=21 ,clip = true, scale=0.23]{Figures/WCPM_Warwick-talk45-220115.pdf}
+\end{figure}
+\end{columns}
+  \begin{itemize}
+    \item Single cell fluoresence time series data (Jaramillo Lab, unpublished).
+    \item  Data records $z= z_{0} +\frac{g}{\alert{\Theta}} $
+    \end{itemize}
+\end{frame}
+
 \subsection{Methodology - least squares error minimisation}
 
-\begin{frame}{Parameter Estimation in ODE Model}
+\begin{frame}{Method for Parameter Estimation in ODE Model}
   \begin{block}{Goal}
@@
 \begin{frame}{Initial Estimation Results}
-  \includegraphics[scale = 0.26, clip = true, trim = 50 0 0 0]{../Figures/13_9_bestPlot}
+  \includegraphics[scale = 0.28, clip = true, trim = 80 0 0 0]{Figures/13_9_bestPlot}
 \end{frame}
@@
 \begin{frame}{Initial Estimation Results}
-  \includegraphics[scale = 0.22, clip = true, trim = 0 0 0 0]{../Figures/13_9_hist}
+  \includegraphics[scale = 0.25, clip = true, trim = 0 0 0 0]{Figures/13_9_hist}
         \begin{itemize}
@@
 \begin{frame}{Sensitivity Analysis}
-  \includegraphics[scale = 0.22, clip = true, trim = 00 0 0 0]{../Figures/Sensitivty_scaled_other}
+  \includegraphics[scale = 0.2, clip = true, trim = 00 0 0 0]{Figures/Sensitivty_scaled_other}
+  \vspace{-2mm}
           \begin{itemize}
@@
 \begin{frame}{Model output for all state variables}
-  \includegraphics[scale = 0.25, clip = true, trim = 30 0 0 0]{../Figures/model_output}
+  \includegraphics[scale = 0.25, clip = true, trim = 30 0 0 0]{Figures/model_output}
 \end{frame}
@@
 
-\begin{frame}{Simplified Model (1)}
+\small
+\begin{frame}{Full model again}
 \begin{align*} 
+\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\alert{\mu} + \alert{\delta_{s}})s -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m \\
+\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\alert{\mu} + \alert{\delta_{m}})m -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m  \\
+\frac{ds:m}{dt} & = \alert{k_{\mathrm{on}}}sm  - (\alert{k_{\mathrm{off}}}+ \alert{k_{\mathrm{hyb}}})s:m  -(\alert{\mu} + \alert{\delta_{m}} )s:m \\
+\frac{dc}{dt} & = \alert{k_{\mathrm{hyb}}}s:m  -(\alert{\mu} + \alert{\delta_{m}})c  
+\end{align*}
+\begin{center}
+\rule{0.5\textwidth}{.4pt}
+\end{center}
+\begin{align*} 
+\frac{dp}{dt} & = \alert{\beta} m +\alert{f_{s}}\beta c -(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
+\frac{dg}{dt} & = \gamma p - (\alert{\mu} + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
+z &= z_{0} +\frac{g}{\alert{\Theta}} 
+\end{align*}
+\end{frame}
+\normalsize
+
+\small
+\begin{frame}{Simplified Model}
+\begin{greyedout}
+\begin{align*} 
+\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\mu + \delta_{s})s -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m \\
+\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\mu + \delta_{m})m -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m  \\
+\frac{ds:m}{dt} & = k_{\mathrm{on}}sm  - (k_{\mathrm{off}}+ k_{\mathrm{hyb}})s:m  -(\mu + \delta_{sm} )s:m \\
+\frac{dc}{dt} & = k_{\mathrm{hyb}}s:m  -(\mu + \delta_{c})c  
+\end{align*}
+\end{greyedout}
+\begin{center}
+\rule{0.5\textwidth}{.4pt}
+\end{center}
+\begin{align*} 
+\frac{dp}{dt} & = \alert{F}y(t)-(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
+\frac{dg}{dt} & = \gamma p - (\alert{\mu} + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g}  \\
+z &= z_{0} +\frac{g}{\alert{\Theta}}  
+\end{align*}
+\end{frame}
+\normalsize
+
+\begin{frame}{Simplified Model}
+\begin{align}
 \frac{dp}{dt} & = \alert{F}y(t)-(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
@@
 z &= z_{0} +\frac{g}{\alert{\Theta}} 
-\end{align*}
+\end{align}
 \begin{itemize}
@@
 
-\begin{frame}{Simplified Model Results (2)}
+\begin{frame}{Simplified Model Results (1)}
   \includegraphics[scale = 0.28, clip = true, trim = 100 0 0 400]{../Figures/13_9_hist_simplified}
@@
 
-\begin{frame}{Simplified Model Results (3)}
-  \includegraphics[scale = 0.25, clip = true, trim = 60 0 0 00]{../Figures/Likelihood_profile_rough}
+\begin{frame}{Simplified Model Results (2)}
+  \includegraphics[scale = 0.25, clip = true, trim = 0 0 0 0]{Figures/Likelihood_rough_presentation}
     \begin{itemize}
-\item  Error landscape, $\Theta = 1000$.
+\item  Error landscape, $\Theta = 5400$.
 \end{itemize}

### 2015-09-26 Presentation/MscPresentation.tex
further presentation work

@@
 \newenvironment{greyedout}{\color{gray}}{\ignorespacesafterend}
+\setbeamercolor{alerted text}{fg = red}
 
@@
 \begin{columns}
-\column{.45\textwidth}
-\includegraphics[trim = 135 0 0 0,clip = true,scale = 0.31]{Figures/schematic_binding}
-\column{.5\textwidth}
+\column{.4\textwidth}
+\includegraphics[trim = 135 0 0 0,clip = true,scale = 0.28]{Figures/schematic_binding}
+\column{.6\textwidth}
   \begin{itemize}
@@
 \begin{align*} 
-\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\mu + \delta_{s})s -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m \\
-\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\mu + \delta_{m})m -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m  \\
-\frac{ds:m}{dt} & = k_{\mathrm{on}}sm  - (k_{\mathrm{off}}+ k_{\mathrm{hyb}})s:m  -(\mu + \delta_{sm} )s:m \\
-\frac{dc}{dt} & = k_{\mathrm{hyb}}s:m  -(\mu + \delta_{c})c  \\
+\frac{d\textcolor{blue}{s}}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\mu + \delta_{s})s -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m \\
+\frac{d\textcolor{blue}{m}}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\mu + \delta_{m})m -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m  \\
+\frac{d\textcolor{blue}{s:m}}{dt} & = k_{\mathrm{on}}sm  - (k_{\mathrm{off}}+ k_{\mathrm{hyb}})s:m  -(\mu + \delta_{sm} )s:m \\
+\frac{d\textcolor{blue}{c}}{dt} & = k_{\mathrm{hyb}}s:m  -(\mu + \delta_{c})c  \\
 \end{align*}
@@
 \begin{columns}
-\column{.45\textwidth}
+\column{.4\textwidth}
 \includegraphics[trim = 0 0 0 0,clip = true,scale = 0.31]{Figures/schematic_translation}
-\column{.5\textwidth}
+\column{.6\textwidth}
   \begin{itemize}
@@
 \begin{align*} 
-\frac{dp}{dt} & = \beta m +f_{s}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
-\frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
+\frac{d\textcolor{blue}{p}}{dt} & = \beta m +f_{s}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
+\frac{d\textcolor{blue}{g}}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
 z &= z_{0} +\frac{g}{\Theta} 
@@
 \frac{ds:m}{dt} & = \alert{k_{\mathrm{on}}}sm  - (\alert{k_{\mathrm{off}}}+ \alert{k_{\mathrm{hyb}}})s:m  -(\alert{\mu} + \alert{\delta_{m}} )s:m \\
-\frac{dc}{dt} & = \alert{k_{\mathrm{hyb}}}s:m  -(\alert{\mu} + \alert{\delta_{m}})c  \\
+\frac{dc}{dt} & = \alert{k_{\mathrm{hyb}}}s:m  -(\alert{\mu} + \alert{\delta_{m}})c  
+\end{align}
+\begin{center}
+\rule{0.5\textwidth}{.4pt}
+\end{center}
+\begin{align}
 \frac{dp}{dt} & = \alert{\beta} m +\alert{f_{s}}\beta c -(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
@@
 \begin{frame}{Initial Estimation Results}
-  \includegraphics[scale = 0.25, clip = true, trim = 0 0 0 0]{Figures/13_9_hist}
+  \includegraphics[scale = 0.28, clip = true, trim = 130 0 0 0]{Figures/13_9_hist}
         \begin{itemize}
@@
 \begin{frame}{Sensitivity Analysis}
-  \includegraphics[scale = 0.2, clip = true, trim = 00 0 0 0]{Figures/Sensitivty_scaled_other}
+  \includegraphics[scale = 0.25, clip = true, trim = 00 0 0 0]{Figures/Sensitivty_scaled_other}
   \vspace{-2mm}
@@
 \begin{frame}{Model output for all state variables}
-  \includegraphics[scale = 0.25, clip = true, trim = 30 0 0 0]{Figures/model_output}
-\end{frame}
-
-\begin{frame}{Model fixed point for estimated parameters}
-  \includegraphics[scale = 0.24, clip = true, trim = 0 0 0 0]{../Figures/fixedpoint_f}
-\begin{itemize}
-	\item Similar fixed point values of translation forcing term, ${\beta} m +{f_{s}}\beta c$, across estimated parameter sets.
-	\item Suggests translation may be rate limiting step.
-\end{itemize}
+  \includegraphics[scale = 0.29, clip = true, trim = 0 0 0 0]{Figures/model_output}
 \end{frame}
@@
 \begin{frame}{Simplified Model Results (1)}
-  \includegraphics[scale = 0.28, clip = true, trim = 100 0 0 400]{../Figures/13_9_hist_simplified}
+  \includegraphics[scale = 0.29, clip = true, trim = 100 0 0 400]{Figures/13_9_hist_simplified}
   \begin{itemize}
@@
 \begin{frame}{Simplified Model Results (2)}
-  \includegraphics[scale = 0.25, clip = true, trim = 0 0 0 0]{Figures/Likelihood_rough_presentation}
+  \includegraphics[scale = 0.26, clip = true, trim = 40 0 0 0]{Figures/Likelihood_rough_presentation}
     \begin{itemize}
@@
 \end{document}
+
+\begin{frame}{Model fixed point for estimated parameters}
+  \includegraphics[scale = 0.24, clip = true, trim = 0 0 0 0]{../Figures/fixedpoint_f}
+\begin{itemize}
+	\item Similar fixed point values of translation forcing term, ${\beta} m +{f_{s}}\beta c$, across estimated parameter sets.
+	\item Suggests translation may be rate limiting step.
+\end{itemize}
+\end{frame}

### 2015-09-28 Presentation/MscPresentation.tex
more presentation work

@@
 
-% Structuring a talk is a difficult task and the following structure
-% may not be suitable. Here are some rules that apply for this
-% solution:
-
-% - Exactly two or three sections (other than the summary).
-% - At *most* three subsections per section.
-% - Talk about 30s to 2min per frame. So there should be between about
-%   15 and 30 frames, all told.
-
-% - A conference audience is likely to know very little of what you
-%   are going to talk about. So *simplify*!
-% - In a 20min talk, getting the main ideas across is hard
-%   enough. Leave out details, even if it means being less precise than
-%   you think necessary.
-% - If you omit details that are vital to the proof/implementation,
-%   just say so once. Everybody will be happy with that.
+\begin{frame}{Objectives}{}
+  \begin{itemize}
+    \item  Rational design of genetic circuits now possible
+      \begin{itemize}
+      \item ODE models used for prediction.
+      \item Model parameters unknown.
+      \end{itemize}
+    \item Project studied a recently designed genetic circuit, attempted parameter estimation.
+    \item Parameter estimates would inform future design/experiments.
+        \end{itemize}
+\end{frame}
 
@@
   \begin{center}
-  \includegraphics[trim = 0 0 0 30,clip = true,scale = 0.275]{Figures/schematic_initial.png}
+  \includegraphics[trim = 370 0 0 200,clip = true,scale = 0.4]{Figures/schematic_initial_with_forcing.png}
   \end{center}
@@
     \item  sRNA binds to it - new complex has RBS uncovered.
+    \item Complex translated into proteins.
     \end{itemize}
@@
 \subsection{ODE Model for system}
-\begin{frame}{ODE Model for system (1) - Complex Formation}{}
+\begin{frame}{ODE model for system (1) - Binding step}{}
 \begin{columns}
@@
   \begin{itemize}
-    \item  Hybridization of sRNA \textcolor{blue}{($s$)} and mRNA \textcolor{blue}{($m$)}, into unstable complex \textcolor{blue}{($s:m$)} then stable one \textcolor{blue}{($c$)}.
+    \item  Binding of sRNA \textcolor{blue}{($s$)} and mRNA \textcolor{blue}{($m$)}, into unstable complex \textcolor{blue}{($s:m$)} then stable one \textcolor{blue}{($c$)}.
     \end{itemize}
@@
 
-\begin{frame}{ODE Model for system (2) - Translation}{}
+\begin{frame}{ODE model for system (2) - Translation step}
 \begin{columns}
@@
 \small
-\begin{frame}{Full model, with Parameters to be estimated}
+\begin{frame}{Full model, with parameters to be estimated}
 \begin{align} 
@@
 
-\begin{frame}{Recent Experimental Data}{}
-\begin{columns}
-\column{.49\textwidth}
-\begin{figure}
-   \includegraphics[trim = 10 300 270 290,clip = true,scale = 0.55]{Figures/13_9}
-\end{figure}
-\column{.49\textwidth}
-\vspace{-5mm}
-\begin{figure}
-   \includegraphics[trim = 0 40 0 0 ,page=21 ,clip = true, scale=0.23]{Figures/WCPM_Warwick-talk45-220115.pdf}
-\end{figure}
-\end{columns}
+\begin{frame}{Recent experimental data}{}
+   \includegraphics[trim = 0 300 0 300,clip = true,scale = 0.55]{Figures/13_9}
   \begin{itemize}
@@
 
-\begin{frame}{Method for Parameter Estimation in ODE Model}
+\begin{frame}{Method for parameter estimation in ODE model}
   \begin{block}{Goal}
@@
 \end{align*}
-\item Used evolutionary algorithm, CMA-ES, to perform minimisation.
+\item Used evolutionary algorithm to perform minimisation.
     \end{itemize}
@@
 
-\begin{frame}{Initial Estimation Results}
+\begin{frame}{Initial estimation results}
   \includegraphics[scale = 0.28, clip = true, trim = 80 0 0 0]{Figures/13_9_bestPlot}
@@
 
-\begin{frame}{Initial Estimation Results}
-  \includegraphics[scale = 0.28, clip = true, trim = 130 0 0 0]{Figures/13_9_hist}
+\begin{frame}{Initial estimation results}
+  \includegraphics[scale = 0.27, clip = true, trim = 130 0 0 50]{Figures/13_9_hist}
+  \vspace{-10mm}
         \begin{itemize}
+    \item Similar error values for all results.
     \item Many parameters poorly estimated.
@@
 
-\begin{frame}{Sensitivity Analysis}
+\begin{frame}{Sensitivity analysis}
+\begin{columns}
+\column{.4\textwidth}
+ \LARGE
+\begin{align*}
+S_{ij} =  \frac{\partial z}{ \partial \theta_{j}}\Bigr|_{t_{i}}
+\end{align*}
+\normalsize
+\column{.59\textwidth}
+Perturbations of parameters all give similar effects on model output - \alert{hard to distinguish}
+\end{columns}
   \includegraphics[scale = 0.25, clip = true, trim = 00 0 0 0]{Figures/Sensitivty_scaled_other}
-  \vspace{-2mm}
-          \begin{itemize}
-    \item Sensitivity analysis suggests perturbations of several parameters all give similar effects on model output - \alert{hard to resolve}
-            \end{itemize}
-\end{frame}
-
-\begin{frame}{Model output for all state variables}
-  \includegraphics[scale = 0.29, clip = true, trim = 0 0 0 0]{Figures/model_output}
 \end{frame}
@@
 \small
-\begin{frame}{Full model again}
+\begin{frame}{Full model, with parameters to be estimated}
 \begin{align*} 
@@
 
+\begin{frame}{Model output for all state variables}
+  \includegraphics[scale = 0.26, clip = true, trim = 0 0 0 0]{Figures/model_output}
+  \begin{itemize}
+    \item $s,m,s:m,c$ \alert{(binding step)} respond instantly to forcing.
+    \item $p, z$ \alert{(translation step)} on timescale of forcing.
+\end{itemize}
+\end{frame}
+
+\footnotesize
+\begin{frame}{Translation rate limiting}
+\begin{align*} 
+\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\alert{\mu} + \alert{\delta_{s}})s -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m \\
+\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\alert{\mu} + \alert{\delta_{m}})m -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m  \\
+\frac{ds:m}{dt} & = \alert{k_{\mathrm{on}}}sm  - (\alert{k_{\mathrm{off}}}+ \alert{k_{\mathrm{hyb}}})s:m  -(\alert{\mu} + \alert{\delta_{m}} )s:m \\
+\frac{dc}{dt} & = \alert{k_{\mathrm{hyb}}}s:m  -(\alert{\mu} + \alert{\delta_{m}})c  
+\end{align*}
+\begin{center}
+\rule{0.5\textwidth}{.4pt}
+\end{center}
+\begin{align*} 
+\frac{dp}{dt} & = \alert{\beta} m +\alert{f_{s}}\beta c -(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
+\frac{dg}{dt} & = \gamma p - (\alert{\mu} + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
+z &= z_{0} +\frac{g}{\alert{\Theta}} 
+\end{align*}
+\normalsize
+\vspace{-5mm}
+  \begin{itemize}
+  \item Binding equations respond instantly.
+    \item $ \alert{\beta} m +\alert{f_{s}}\beta c $ term links binding and translation. It flips between fixed point values.
+    \item All parameters `upstream' have identical effects.
+\end{itemize}
+\end{frame}
+
 \small
-\begin{frame}{Simplified Model}
+\begin{frame}{Simplified model}
 \begin{greyedout}
@@
 
-\begin{frame}{Simplified Model}
+\begin{frame}{Simplified model}
 \begin{align}
@@
 
-\begin{frame}{Simplified Model Results (1)}
+\begin{frame}{Simplified model results (1)}
   \includegraphics[scale = 0.29, clip = true, trim = 100 0 0 400]{Figures/13_9_hist_simplified}
@@
 
-\begin{frame}{Simplified Model Results (2)}
+\begin{frame}{Simplified model results (2)}
   \includegraphics[scale = 0.26, clip = true, trim = 40 0 0 0]{Figures/Likelihood_rough_presentation}
@@
 
-\begin{frame}{Future Work}
-
-  % Keep the summary *very short*.
+\begin{frame}{Future work}
   \begin{itemize}
   \item Fluorescence data \alert {not enough} to estimate all unknown parameters 
-  \begin{itemize}
-  \item Further experiments ?
-  \item Bounds on unknown parameter values.
+  \item Further experiments needed - observe fast timescale.
+  \item Investigate methods of McAuley et al. 
+  \item Change methodology - Bayesian Methods?
   \end{itemize}
-  \item Improve modelling of translation step
-  \item Change methodology - Bayesian Methods?
-
-  \end{itemize}
-
   \end{frame}
 
-  \section*{Summary}
+  \begin{frame}{Acknowledgements}
+  \begin{itemize}
+  \item Manish Kushawaha, Jaramillo Lab - School of Life Sciences
+  \item Shenshi Shen, Boris Kisov for data.
+  \item Annabelle Ballesta - Systems Biology
+  \end{itemize}
+  \end{frame}
 
@@
 \end{frame}
+
+\begin{frame}{Recent Experimental Data}{}
+\begin{columns}
+\column{.49\textwidth}
+\begin{figure}
+   \includegraphics[trim = 10 300 270 290,clip = true,scale = 0.55]{Figures/13_9}
+\end{figure}
+\column{.49\textwidth}
+\vspace{-5mm}
+\begin{figure}
+   \includegraphics[trim = 0 40 0 0 ,page=21 ,clip = true, scale=0.23]{Figures/WCPM_Warwick-talk45-220115.pdf}
+\end{figure}
+\end{columns}
+  \begin{itemize}
+    \item Single cell fluoresence time series data (Jaramillo Lab, unpublished).
+    \item  Data records $z= z_{0} +\frac{g}{\alert{\Theta}} $
+    \end{itemize}
+\end{frame}

### 2015-09-29 Presentation/MscPresentation.tex
more presentation work
final commits

@@
   \begin{itemize}
+  \item Goal in synthetic biology - manipulate natural cellular machinery
     \item  Rational design of genetic circuits now possible
@@
       \end{itemize}
-    \item Project studied a recently designed genetic circuit, attempted parameter estimation.
+    \item Project studied a recently designed genetic circuit 
+    \item \textcolor{blue}{Aim}: use \textcolor{blue}{ODE model} with \textcolor{blue}{recent data} to perform \textcolor{blue}{parameter estimation}.
     \item Parameter estimates would inform future design/experiments.
@@
     \item Translation of stable complex \textcolor{blue}{($c$)} into immature GFP \textcolor{blue}{($p$)}. 
-    \item maturation of GFP \textcolor{blue}{($g$)}, machine calibration giving measured fluorescence ($z$).
+    \item maturation of GFP \textcolor{blue}{($g$)}, machine calibration giving measured fluorescence \textcolor{blue}{($z$)}.
     \end{itemize}
@@
   \begin{itemize}
+      \item  Data records $z= z_{0} +\frac{g}{\alert{\Theta}} $
     \item Single cell fluoresence time series data (Jaramillo Lab, unpublished).
-    \item  Data records $z= z_{0} +\frac{g}{\alert{\Theta}} $
     \end{itemize}
@@
 \begin{frame}{Initial estimation results}
-  \includegraphics[scale = 0.27, clip = true, trim = 130 0 0 50]{Figures/13_9_hist}
+  \includegraphics[scale = 0.27, clip = true, trim = 100 0 0 40]{Figures/13_9_hist}
   \vspace{-10mm}
         \begin{itemize}
+            \item Many parameters poorly estimated.
     \item Similar error values for all results.
-    \item Many parameters poorly estimated.
             \end{itemize}
@@
 \small
-\begin{frame}{Full model, with parameters to be estimated}
+\begin{frame}{Plotting all states of model output}
 \begin{align*} 
@@
 \begin{frame}{Simplified model results (1)}
-  \includegraphics[scale = 0.29, clip = true, trim = 100 0 0 400]{Figures/13_9_hist_simplified}
+  \includegraphics[scale = 0.29, clip = true, trim = 70 0 0 400]{Figures/13_9_hist_simplified}
   \begin{itemize}
@@
 
-\begin{frame}{Future work}
+\begin{frame}{Conclusions and further work}
   \begin{itemize}
   \item Fluorescence data \alert {not enough} to estimate all unknown parameters 
+  \begin{itemize}
+  \item Simplified model not helpful in estimating original parameters.
+  \end{itemize}
+  \end{itemize}
+To get these parameters ...
+  \begin{itemize}
   \item Further experiments needed - observe fast timescale.
-  \item Investigate methods of McAuley et al. 
-  \item Change methodology - Bayesian Methods?
+  \item Improve methodology
+    \begin{itemize}
+    \item Algorithms for dealing with complex chemical models \& limited data.
+    \item Bayesian Methods
+      \end{itemize}
   \end{itemize}
@@
 
-\end{document}
+  \begin{frame}{Additional Slides}
+  \Huge
+	Additional Slides
+	\normalsize
+  \end{frame}
 
@@
 \end{frame}
+
+\end{document}
 
